\chapter{Remembering His teaching in our past history}


\chapter{Pamiętając Jego nauczanie w naszej przeszłej historii}


We are bringing our journey in this study to its end. We started this journey with a strong proclamation of what constitutes the foundation of our faith. We became acquainted with the history and experience of our pioneers in establishing the Seventh-day Adventist Church. We have seen the mission and purpose God bestowed upon them in proclaiming the three angels’ messages to the whole world. These messages are interwoven with all important doctrines of the Bible. These doctrines are what our pioneers called the \emcap{fundamental principles}, or the pillars of our faith. These doctrines represent the foundation of our faith.


Kończymy naszą podróż w tym studium. Rozpoczęliśmy tę podróż od silnego ogłoszenia tego, co stanowi fundament naszej wiary. Zapoznaliśmy się z historią i doświadczeniem naszych pionierów w ustanawianiu Kościoła Adwentystów Dnia Siódmego. Zobaczyliśmy misję i cel, który Bóg im powierzył w głoszeniu poselstw trzech aniołów całemu światu. Te poselstwa są powiązane ze wszystkimi ważnymi doktrynami Biblii. Te doktryny to to, co nasi pionierzy nazywali \emcap{fundamentalnymi zasadami} lub filarami naszej wiary. Te doktryny stanowią fundament naszej wiary.


Once acquainted with the \emcap{Fundamental Principles} that we had in the beginning, we recognize their difference to our current Fundamental Beliefs, particularly to the question “\textit{who is God}”? Additionally, the present doctrine of God lacks understanding of His personality. In other words, it lacks the understanding of \textit{the quality or state} of God \textit{being a person}. This change in our doctrine becomes important in light of the first angel’s message, which pertains to the God we worship. Is the God we worship a triune God, or is He one God, the Father, the Ancient of Days, a personal and spiritual Being?


Po zapoznaniu się z \emcap{Fundamentalnymi Zasadami}, które mieliśmy na początku, dostrzegamy ich różnicę w stosunku do naszych obecnych Fundamentalnych Wierzeń, szczególnie w kwestii “\textit{kim jest Bóg}”? Dodatkowo, obecna doktryna o Bogu nie uwzględnia zrozumienia Jego osobowości. Innymi słowy, brakuje w niej zrozumienia \textit{właściwości lub stanu} Boga \textit{jako osoby}. Ta zmiana w naszej doktrynie staje się ważna w świetle poselstwa pierwszego anioła, które dotyczy Boga, którego czcimy. Czy Bóg, którego czcimy, jest Bogiem w trójcy, czy jest jednym Bogiem, Ojcem, Przedwiecznym, osobową i duchową Istotą?


The change in our understanding of who God is for Seventh-day Adventists was not instantaneous; it took years of controversy to arrive at our present standing. In these studies, we did not delve into history beyond the life of Ellen White. We saw change begin to take place in her time, when Dr. Kellogg insinuated the sentiment over the \emcap{personality of God}, which would \egwinline{lead astray the minds of those who are not thoroughly established on the foundation principles of present truth}[SpTB02 51.3; 1904][https://egwwritings.org/?ref=en\_SpTB02.51.3&para=417.262]. He insinuated doubt in the clear revelations of the \emcap{personality of God} and of His Son, Jesus Christ. His sentiments were met by fierce rebukes from Sister White and strong warnings for the church, to avoid the path of doubt in the simple truth expressed in the \emcap{Fundamental Principles}—that one God is a personal, spiritual being, and Christ is His Son, \egwinline{begotten in the express image of the Father’s person}[ST May 30, 1895, par. 3; 1895][https://egwwritings.org/?ref=en\_ST.May.30.1895.par.3&para=820.12891]. Thusly, Sister White firmly defended the first two points of the \emcap{Fundamental Principles}.


Zmiana w naszym zrozumieniu, kim jest Bóg dla Adwentystów Dnia Siódmego, nie była natychmiastowa; zajęło lata sporów, aby dojść do naszego obecnego stanowiska. W tych studiach nie zagłębialiśmy się w historię poza życiem Ellen White. Widzieliśmy, jak zmiany zaczęły zachodzić za jej życia, gdy dr Kellogg insynuował poglądy na temat \emcap{osobowości Boga}, które \egwinline{sprowadzą na manowce umysły tych, którzy nie są gruntownie utwierdzeni na fundamentalnych zasadach obecnej prawdy}[SpTB02 51.3; 1904][https://egwwritings.org/?ref=en\_SpTB02.51.3&para=417.262]. Insynuował wątpliwości w jasne objawienia \emcap{osobowości Boga} i Jego Syna, Jezusa Chrystusa. Jego poglądy spotkały się z ostrymi naganami ze strony Siostry White i silnymi ostrzeżeniami dla kościoła, aby unikać ścieżki wątpliwości w prostej prawdzie wyrażonej w \emcap{Fundamentalnych Zasadach} - że jeden Bóg jest osobową, duchową istotą, a Chrystus jest Jego Synem, \egwinline{zrodzonym na wyraz Jego istoty}[ST May 30, 1895, par. 3; 1895][https://egwwritings.org/?ref=en\_ST.May.30.1895.par.3&para=820.12891]. W ten sposób Siostra White stanowczo broniła pierwszych dwóch punktów \emcap{Fundamentalnych Zasad}.


Just as in the time of Dr. Kellogg, when many brethren were departing from the simplicity of Christ’s teaching, so it is today. Sister White prophesied that this change over the understanding of the \emcap{personality of God} would occur in our church, and re-establishment of our old foundation of faith would be necessary. Will the \emcap{Fundamental Principles} be re-established in our midst? The outcome of it lies completely on every individual that makes up the body of the Seventh-day Adventist Church. Today, in this day and age, is when this re-establishment will take place. The warnings and rebukes uttered by the pen of the Spirit of Prophecy have never been more relevant than today. God has put the final outcome into your hands. If these warnings touch the core of your soul, God is calling you to stand firm on the platform of eternal Truth. He calls you to hold firmly to the \emcap{Fundamental Principles}, which are based on unquestionable authority.


Tak jak w czasach dr. Kelloga, gdy wielu braci odchodziło od prostoty nauczania Chrystusa, tak jest i dzisiaj. Siostra White przepowiedziała, że ta zmiana w zrozumieniu \emcap{osobowości Boga} nastąpi w naszym kościele i konieczne będzie ponowne ustanowienie naszego starego fundamentu wiary. Czy \emcap{Fundamentalne Zasady} zostaną ponownie ustanowione wśród nas? Wynik tego zależy całkowicie od każdej jednostki, która tworzy ciało Kościoła Adwentystów Dnia Siódmego. Dziś, w tej epoce, jest czas, kiedy to ponowne ustanowienie nastąpi. Ostrzeżenia i nagany wypowiedziane przez pióro Ducha Proroctwa nigdy nie były bardziej aktualne niż dziś. Bóg złożył ostateczny wynik w twoje ręce. Jeśli te ostrzeżenia dotykają sedna twojej duszy, Bóg wzywa cię do stanowczego stania na platformie wiecznej Prawdy. Wzywa cię do mocnego trzymania się \emcap{Fundamentalnych Zasad}, które opierają się na niepodważalnym autorytecie.


Below, we present a portion of the letter from Sister White to Dr. Kellogg, in which we find solemn warning for us today in re-establishing the foundation of our faith. When we are acquainted with the truth on the \emcap{personality of God} and its controversy over the Trinity doctrine, the following letter shines in a new light, with messages and principles that are valuable for us today, that we may know how to behave in our current state of affairs.


Poniżej przedstawiamy fragment listu Siostry White do dr. Kelloga, w którym znajdujemy uroczyste ostrzeżenie dla nas dzisiaj w ponownym ustanawianiu fundamentu naszej wiary. Kiedy zapoznamy się z prawdą o \emcap{osobowości Boga} i jej sporze z doktryną o Trójcy, poniższy list jaśnieje w nowym świetle, z przesłaniami i zasadami, które są cenne dla nas dzisiaj, abyśmy wiedzieli, jak postępować w naszym obecnym stanie rzeczy.


\egw{\textbf{The Foundation of Our Faith}}


\egw{\textbf{Fundament Naszej Wiary}}


\egw{In regard to the book Living Temple, \textbf{I have been instructed by the heavenly messenger} that \textbf{some of the reasoning in this book is untrue, and that this reasoning would lead astray the minds of those who are not thoroughly established \underline{on the foundation principles} of present truth.} \textbf{It introduces that which is naught but speculation in regard to \underline{the personality of God} and \underline{where His presence is}}. No one on this earth has a right to speculate on this question. ‘The secret things belong unto the Lord our God, but those things which are revealed belong unto us, and to our children forever.’}[Lt232-1903.39; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.39&para=10197.48]


\egw{Odnośnie książki Living Temple, \textbf{zostałam pouczona przez niebiańskiego posłańca}, że \textbf{niektóre rozumowanie w tej książce jest nieprawdziwe i że to rozumowanie sprowadziłoby na manowce umysły tych, którzy nie są gruntownie utwierdzeni \underline{na fundamentalnych zasadach} obecnej prawdy.} \textbf{Wprowadza to, co jest niczym innym jak spekulacją odnośnie \underline{osobowości Boga} i \underline{gdzie jest Jego obecność}}. Nikt na tej ziemi nie ma prawa spekulować na ten temat. ‘Rzeczy tajemne należą do Pana, Boga naszego, a rzeczy objawione należą do nas i do naszych dzieci na wieki.’}[Lt232-1903.39; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.39&para=10197.48]


\egwnogap{\textbf{I am authorized by the Lord to say, The sentiments contained in Living Temple in regard to the personality of God \underline{are opposed to the truth that God has given us}}. \textbf{The truth for this time is now to be brought before the people.} Our brethren and sisters in every church and in every place are to guard carefully against allowing their minds to be engrossed with matters that draw them away from eternal things. The enemy will use some of the statements made in Living Temple to tempt some as he tempted Adam and Eve in Eden. \textbf{I warn our brethren not to enter into controversy over the presence and personality of God. \underline{The statements made in Living Temple in regard to this point are incorrect}.} The Scripture used to substantiate the doctrine there set forth is Scripture misapplied.}[Lt232-1903.40; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.40&para=10197.49]


\egwnogap{\textbf{Jestem upoważniona przez Pana, aby powiedzieć, że poglądy zawarte w Living Temple odnośnie osobowości Boga \underline{są przeciwne prawdzie, którą dał nam Bóg}}. \textbf{Prawda na ten czas ma być teraz przedstawiona ludziom.} Nasi bracia i siostry w każdym kościele i w każdym miejscu muszą uważnie strzec się, aby ich umysły nie zostały pochłonięte sprawami, które odciągają ich od rzeczy wiecznych. Wróg użyje niektórych stwierdzeń zawartych w Living Temple, aby kusić niektórych, tak jak kusił Adama i Ewę w Edenie. \textbf{Ostrzegam naszych braci, aby nie wchodzili w spór dotyczący obecności i osobowości Boga. \underline{Stwierdzenia zawarte w Living Temple odnośnie tego punktu są niepoprawne}.} Pismo Święte użyte do poparcia doktryny tam przedstawionej jest błędnie zastosowanym Pismem.}[Lt232-1903.40; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.40&para=10197.49]


\egwnogap{I was cautioned not to enter into controversy regarding the question that will come up over these things, \textbf{because controversy might lead men to resort to \underline{subterfuges}, and their minds would be led away from the truth of the Word of God \underline{to assumption and guesswork}. The more that fanciful theories are discussed, the less men will know of God and of the truth that sanctifies the soul}.}[Lt232-1903.41; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.41&para=10197.50]


\egwnogap{Zostałam przestrzeżona, aby nie wchodzić w spór dotyczący kwestii, które pojawią się w związku z tymi sprawami, \textbf{ponieważ spór mógłby prowadzić ludzi do uciekania się do \underline{wybiegów}, a ich umysły zostałyby odciągnięte od prawdy Słowa Bożego \underline{do założeń i domysłów}. Im więcej dyskutuje się o fantazyjnych teoriach, tym mniej ludzie będą wiedzieć o Bogu i o prawdzie, która uświęca duszę}.}[Lt232-1903.41; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.41&para=10197.50]


\egwnogap{We are God’s commandment-keeping people. \textbf{For the last fifty years every phase of heresy has been brought to bear upon us, to tear down \underline{the foundation principles of our faith}}. Messages of every order and kind have been urged upon Seventh-day Adventists to take the place of the truth which \textbf{point by point} has been testified to by the miracle-working power of the Lord. \textbf{But the waymarks which have made us what we are are to be preserved, and they \underline{will be preserved}, as God has signified through His Word and the testimony of His Spirit. From the great system of truth as it has been presented by God’s messengers, \underline{not a pin is to be removed}}.}[Lt232-1903.42; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.42&para=10197.51]


\egwnogap{Jesteśmy ludem przestrzegającym przykazań Bożych. \textbf{Przez ostatnie pięćdziesiąt lat każda forma herezji była kierowana przeciwko nam, aby zniszczyć \underline{fundamentalne zasady naszej wiary}}. Wiadomości wszelkiego rodzaju i typu były narzucane Adwentystom Dnia Siódmego, aby zastąpić prawdę, która \textbf{punkt po punkcie} została potwierdzona przez cudowną moc Pana. \textbf{Ale znaki, które uczyniły nas tym, kim jesteśmy, mają być zachowane i \underline{będą zachowane}, jak Bóg oznajmił przez Swoje Słowo i świadectwo Swojego Ducha. Z wielkiego systemu prawdy, jaki został przedstawiony przez posłańców Bożych, \underline{nie wolno usunąć ani jednego pinu}}.}[Lt232-1903.42; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.42&para=10197.51]


\egwnogap{\textbf{I am called upon by God to stand in defense of the truth that has been given us as we have followed the leading of Him who is the way, the truth, and the life. Let every pioneer in the work adhere firmly to this truth. \underline{The peculiarities of our faith are to be held fast with the grip of faith}.}}[Lt232-1903.43; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.43&para=10197.52]


\egwnogap{\textbf{Jestem powołana przez Boga, aby stanąć w obronie prawdy, która została nam dana, gdy podążaliśmy za przewodnictwem Tego, który jest drogą, prawdą i życiem. Niech każdy pionier w tym dziele mocno trzyma się tej prawdy. \underline{Szczególne cechy naszej wiary muszą być mocno trzymane uściskiem wiary}.}}[Lt232-1903.43; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.43&para=10197.52]


\egwnogap{The fables that at the present time are being framed by some medical missionary workers are not to be regarded as truth. \textbf{\underline{Their true origin will ere long be revealed.}} \textbf{It will be seen that they were formed under the subtle power of the great apostate, who works as an angel of light, controlling minds by deceptions so concealed that he seeks by them to deceive if possible the very elect}.}[Lt232-1903.44; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.44&para=10197.53]


\egwnogap{Bajki, które obecnie są tworzone przez niektórych pracowników misji medycznej, nie powinny być uważane za prawdę. \textbf{\underline{Ich prawdziwe pochodzenie zostanie wkrótce ujawnione.}} \textbf{Okaże się, że zostały one utworzone pod subtelną mocą wielkiego odstępcy, który działa jako anioł światłości, kontrolując umysły przez tak ukryte oszustwa, że stara się przez nie zwieść, jeśli to możliwe, nawet wybranych}.}[Lt232-1903.44; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.44&para=10197.53]


\egwnogap{What influence but that of the deceiver could lead men at \textbf{this stage of our history to work in an underhand, powerful way to tear down \underline{the foundations of our faith}—\underline{the foundations which were laid at the beginning of our work} by prayerful study of the Word and by revelation}. \textbf{\underline{Upon these foundations we have been building for the last fifty years}}. Shall a new foundation be built up by men to whom God has not granted the special experience He has granted to the men whom He ordained to establish the foundations of our faith? \textbf{The men who are striving to build up this false foundation may suppose that they have found a new way, and that they can lay a stronger foundation than that which has been laid. \underline{But this is a great deception}. \underline{Other foundation can no man lay than that which has been laid}.}}[Lt232-1903.45; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.45&para=10197.54]


\egwnogap{Jaki wpływ, jeśli nie zwodziciela, mógłby prowadzić ludzi na \textbf{tym etapie naszej historii do działania w ukryty, potężny sposób, aby zniszczyć \underline{fundamenty naszej wiary}—\underline{fundamenty, które zostały położone na początku naszego dzieła} przez modlitewne studiowanie Słowa i przez objawienie}. \textbf{\underline{Na tych fundamentach budujemy od pięćdziesięciu lat}}. Czy nowy fundament ma być zbudowany przez ludzi, którym Bóg nie udzielił szczególnego doświadczenia, jakie dał ludziom, których wyznaczył do ustanowienia fundamentów naszej wiary? \textbf{Ludzie, którzy starają się zbudować ten fałszywy fundament, mogą przypuszczać, że znaleźli nową drogę i że mogą położyć silniejszy fundament niż ten, który został położony. \underline{Ale to jest wielkie oszustwo}. \underline{Innego fundamentu nikt nie może położyć niż ten, który został położony}.}}[Lt232-1903.45; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.45&para=10197.54]


\egwnogap{I am instructed to say to our people that in the past many have undertaken the building of a new faith, the establishment of new principles. But how long did their building stand? It soon fell to pieces; \textbf{for it was not founded upon the Rock}.}[Lt232-1903.46; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.46&para=10197.55]


\egwnogap{Jestem pouczona, aby powiedzieć naszemu ludowi, że w przeszłości wielu podejmowało się budowania nowej wiary, ustanawiania nowych zasad. Ale jak długo stała ich budowla? Wkrótce rozpadła się na kawałki; \textbf{ponieważ nie była zbudowana na Skale}.}[Lt232-1903.46; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.46&para=10197.55]


\egwnogap{Did not the first disciples have to meet the sayings of men? Did they not have to listen to false theories and then stand firm, having done all, to stand, saying, ‘Other foundation can no man lay than that which is laid’? One class after another arose with false doctrines, because men were so little acquainted with God.}[Lt232-1903.47; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.47&para=10197.56]


\egwnogap{Czy pierwsi uczniowie nie musieli stawić czoła wypowiedziom ludzi? Czy nie musieli słuchać fałszywych teorii, a potem stać mocno, uczyniwszy wszystko, aby stać, mówiąc: ‘Innego fundamentu nikt nie może położyć niż ten, który jest położony’? Jedna klasa po drugiej powstawała z fałszywymi doktrynami, ponieważ ludzie tak mało znali Boga.}[Lt232-1903.47; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.47&para=10197.56]


\egwnogap{\textbf{My brethren and sisters, study the thirteenth, fourteenth, fifteenth, sixteenth, and seventeenth chapters of John. The words of these chapters explain themselves. ‘This is life eternal,’ Christ declared, ‘that they might know Thee the only true God, and Jesus Christ, whom Thou hast sent.’ \underline{In these words the personality of God and of His Son is clearly spoken of.} \underline{The personality of the one does not do away with the necessity for the personality of the other}.}}[Lt232-1903.48; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.48&para=10197.57]


\egwnogap{\textbf{Moi bracia i siostry, studiujcie rozdziały trzynasty, czternasty, piętnasty, szesnasty i siedemnasty Ewangelii Jana. Słowa tych rozdziałów same się wyjaśniają. ‘A to jest życie wieczne,’ oświadczył Chrystus, ‘aby poznali ciebie, jedynego prawdziwego Boga i Jezusa Chrystusa, którego posłałeś.’ \underline{W tych słowach wyraźnie mówi się o osobowości Boga i Jego Syna.} \underline{Osobowość jednego nie eliminuje konieczności osobowości drugiego}.}}[Lt232-1903.48; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.48&para=10197.57]


\egwnogap{God is never to be understood by any human being. His ways and His works are past finding out. In regard to the revelations \textbf{that He has made of Himself in His Word, we may talk}. \textbf{But when it comes to talking or writing of God’s person and presence, let us say, ‘Thou art God, and Thy ways are past finding out.’}}[Lt232-1903.49; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.49&para=10197.58]


\egwnogap{Bóg nigdy nie może być w pełni zrozumiany przez żadną ludzką istotę. Jego drogi i Jego dzieła są niezbadane. W odniesieniu do objawień \textbf{które dał o sobie w swoim Słowie, możemy mówić}. \textbf{Ale gdy przychodzi do mówienia lub pisania o osobie i obecności Boga, powiedzmy: ‘Ty jesteś Bogiem, a Twoje drogi są niezbadane.’}}[Lt232-1903.49; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.49&para=10197.58]


\egwnogap{It is sacrilegious to put into the minds of young or old the \textbf{seeds of \underline{speculation} regarding this subject}. Such seeds, planted and left to grow, will spring up and \textbf{bring forth a harvest of \underline{infidel sentiments}}. I give this warning to all. \textbf{We want no such sophistry as that presented in Living Temple}. There are excellent things in the book. But there are also tares among the wheat. The book contains many correct ideas, but it contains also statements that will do harm. Those who accept the chaff for the wheat will find themselves losing their sense of God’s greatness and bringing Him into cheap commonness. This is the work of the great deceiver. \textbf{Our brethren are not to be called from their work to study the question of where God is and what He is. We are not to dare to engage in this discussion, lest we be destroyed.} When the ark of God was being taken from the land of the Philistines to the camp of Israel, curiosity led the men of Bethshemesh to look into it. God was displeased, and many were smitten with death.}[Lt232-1903.50; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.50&para=10197.59]


\egwnogap{Jest świętokradztwem umieszczać w umysłach młodych lub starszych \textbf{nasiona \underline{spekulacji} na ten temat}. Takie nasiona, zasiane i pozostawione do wzrostu, wykiełkują i \textbf{przyniosą żniwo \underline{niewiernych poglądów}}. Daję to ostrzeżenie wszystkim. \textbf{Nie chcemy takiej sofistyki, jaka została przedstawiona w Living Temple}. W książce są doskonałe rzeczy. Ale są tam również chwasty wśród pszenicy. Ci, którzy przyjmują plewy zamiast pszenicy, odkryją, że tracą poczucie wielkości Boga i sprowadzają Go do taniej pospolitości. To jest dzieło wielkiego zwodziciela. \textbf{Nasi bracia nie mają być odwoływani od swojej pracy, aby studiować kwestię, gdzie jest Bóg i czym On jest. Nie wolno nam odważyć się angażować w tę dyskusję, abyśmy nie zostali zniszczeni.} Kiedy arka Boża była przenoszona z ziemi Filistynów do obozu Izraela, ciekawość skłoniła mężczyzn z Bet-Szemesz do zajrzenia do niej. Bóg był niezadowolony i wielu zostało porażonych śmiercią.}[Lt232-1903.50; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.50&para=10197.59]


\egwnogap{\textbf{Let us talk of Christ, His preexistence, His humble ministry, His mighty power, His prospective personal glory in the heavenly courts. The Son of God restores to life whom He will. ‘All that the Father hath is Mine,’ He says. ‘I and My Father are one.’ He has greatness, present and prospective, that baffles human conception. He encircles the race with His long human arm, while with His divine arm He grasps the throne of the Infinite}.}[Lt232-1903.51; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.51&para=10197.60]


\egwnogap{\textbf{Mówmy o Chrystusie, Jego preegzystencji, Jego pokornej służbie, Jego potężnej mocy, Jego przyszłej osobistej chwale w niebiańskich dworach. Syn Boży przywraca do życia, kogo chce. ‘Wszystko, co ma Ojciec, jest moje,’ mówi. ‘Ja i Ojciec jedno jesteśmy.’ Ma wielkość, obecną i przyszłą, która przekracza ludzkie pojęcie. Obejmuje rasę ludzką swoim długim ludzkim ramieniem, podczas gdy swoim boskim ramieniem chwyta tron Nieskończonego}.}[Lt232-1903.51; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.51&para=10197.60]


\egwnogap{\textbf{There is a knowledge of God and of Christ which all who are saved must have. ‘\underline{This is life eternal},’ Christ says, ‘\underline{that they might know Thee, the only true God, and Jesus Christ whom Thou hast sent}.’ And He says again, ‘If any man will come after Me, let him deny himself, and take up his cross, and follow Me.’ To all who receive Him as their Redeemer, He gives power to become the sons of God. Every one who truly believes in Him will be inspired by faith and raised by the arm of Omnipotence.}}[Lt232-1903.52; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.52&para=10197.61]


\egwnogap{\textbf{Istnieje wiedza o Bogu i o Chrystusie, którą wszyscy, którzy mają być zbawieni, muszą posiadać. ‘\underline{To jest życie wieczne},’ mówi Chrystus, ‘\underline{aby poznali Ciebie, jedynego prawdziwego Boga, i Jezusa Chrystusa, którego posłałeś}.’ I mówi ponownie: ‘Jeśli kto chce pójść za Mną, niech się zaprze samego siebie, niech weźmie swój krzyż i niech Mnie naśladuje.’ Wszystkim, którzy przyjmują Go jako swojego Odkupiciela, daje moc, aby stali się synami Bożymi. Każdy, kto prawdziwie w Niego wierzy, będzie natchniony wiarą i podniesiony ramieniem Wszechmocy.}}[Lt232-1903.52; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.52&para=10197.61]


\egwnogap{\textbf{Those who do not receive in faith God’s plan for redeeming the race} do despite to the Spirit of grace, and at the last great day their sentence will be, ‘Depart from Me.’ They have hated righteousness and fostered iniquity, and they must be banished forever from the presence of God, exiled from happiness to death—eternal death.}[Lt232-1903.53; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.53&para=10197.62]


\egwnogap{\textbf{Ci, którzy nie przyjmują z wiarą Bożego planu odkupienia ludzkości}, znieważają Ducha łaski, a w dniu ostatecznego sądu usłyszą wyrok: ‘Odejdźcie ode Mnie.’ Nienawidzili sprawiedliwości i pielęgnowali nieprawość, dlatego muszą być na zawsze wygnani sprzed oblicza Boga, wygnani ze szczęścia ku śmierci – wiecznej śmierci.}[Lt232-1903.53; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.53&para=10197.62]


\egwnogap{Those who in this life love God and cherish the thought of Him will employ their faculties and their talents as faithful stewards, making the very best use of them, but not claiming any reward as their due. \textbf{As they deny self and follow Jesus, lifting the cross, they will find that the cross is light, and that it is a pledge, as they bear it, that they will one day be given a crown of everlasting life}. What will be the glory and the gain and the enjoyment of that eternal life that is to be given to those only for whom it has been prepared? The great joy of the overcomer will be that he is in the presence of Christ. ‘Where I am, there shall also My servant be,’ He declared. And He prayed, ‘Father, I will that they also whom Thou hast given Me be with Me where I am; that they may behold My glory.’ Christ is speaking of the glory of His Father’s presence and His Father’s house. \textbf{The glory that is to be revealed to all who are saved is the glory which Christ had with His Father before the world was—\underline{the unapproachable splendor of their converse together}}. \textbf{The angels were not admitted to the interviews between the Father and the Son when the plan of salvation was laid.} Those human beings who seek to intrude into the secrets of the Most High, who inhabiteth eternity, show their ignorance of spiritual and eternal things. Far better might they, while mercy’s voice is still heard, humble themselves in the dust and plead with God to teach them His ways.}[Lt232-1903.54; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.54&para=10197.63]


\egwnogap{Ci, którzy w tym życiu kochają Boga i pielęgnują myśl o Nim, będą wykorzystywać swoje zdolności i talenty jako wierni szafarze, wykorzystując je najlepiej jak potrafią, ale nie domagając się żadnej nagrody jako należnej im. \textbf{Gdy zapierają się siebie i naśladują Jezusa, dźwigając krzyż, odkrywają, że krzyż jest lekki i jest obietnicą, że gdy go niosą, pewnego dnia otrzymają koronę życia wiecznego}. Czym będzie chwała, zysk i radość tego wiecznego życia, które ma być dane tylko tym, dla których zostało przygotowane? Wielką radością zwycięzcy będzie to, że jest w obecności Chrystusa. “Gdzie Ja jestem, tam będzie i mój sługa”, oświadczył. I modlił się: “Ojcze, chcę, aby ci, których Mi dałeś, byli ze Mną tam, gdzie Ja jestem, aby oglądali moją chwałę”. Chrystus mówi o chwale obecności Jego Ojca i domu Jego Ojca. \textbf{Chwała, która ma być objawiona wszystkim zbawionym, to chwała, którą Chrystus miał ze swoim Ojcem, zanim powstał świat – \underline{niedostępny splendor ich wspólnych rozmów}}. \textbf{Aniołowie nie byli dopuszczeni do rozmów między Ojcem a Synem, gdy układany był plan zbawienia.} Ci ludzie, którzy próbują wdzierać się w tajemnice Najwyższego, który zamieszkuje wieczność, pokazują swoją niewiedzę o sprawach duchowych i wiecznych. O wiele lepiej byłoby dla nich, dopóki głos miłosierdzia jest jeszcze słyszalny, ukorzyć się w prochu i błagać Boga, aby nauczył ich swoich dróg.}[Lt232-1903.54; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.54&para=10197.63]


\egw{A Timely Warning}


\egw{Aktualne Ostrzeżenie}


\egw{There are those who have been seeking to carry out their own selfish designs, without regard to the influence that this would have upon the cause and work of God. It is time that such ones felt the inward work of grace upon their hearts, that the medical missionary work may not be grossly misrepresented. Let not our medical missionary workers become so like the world in habit and practice that worldlings will turn away from them with scorn, saying, ‘I am just as good as they are.’ There are instances where the medical missionary work has been so conducted that the name ‘medical missionary’ might better be dropped; for it has been badly misrepresented, and God has been dishonored.}[Lt232-1903.55; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.55&para=10197.65]


\egw{Są tacy, którzy starali się realizować swoje własne samolubne plany, nie zważając na wpływ, jaki miałoby to na sprawę i dzieło Boże. Nadszedł czas, aby tacy odczuli wewnętrzne działanie łaski w swoich sercach, aby praca misyjna medyczna nie była rażąco przedstawiana w złym świetle. Niech nasi pracownicy misji medycznej nie stają się tak podobni do świata w zwyczajach i praktykach, że ludzie ze świata odwrócą się od nich z pogardą, mówiąc: ‘Jestem tak samo dobry jak oni’. Są przypadki, gdy praca misyjna medyczna była prowadzona w taki sposób, że nazwa “misjonarz medyczny” mogłaby być lepiej pominięta; ponieważ została źle przedstawiona, a Bóg został zniesławiony.}[Lt232-1903.55; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.55&para=10197.65]


\egwnogap{\textbf{We are living amidst the perils of the last days. Our people must now arouse to the work before them. We are to lift up the standard and proclaim the last message of warning to a perishing world. Those who have a knowledge of the truth for this time are now to hold firmly aloft the banner bearing the inscription, ‘The commandments of God and the faith of Jesus.’ }[Revelation 14:12.]}[Lt232-1903.56; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.56&para=10197.66]


\egwnogap{\textbf{Żyjemy wśród niebezpieczeństw ostatnich dni. Nasz lud musi teraz obudzić się do pracy, która jest przed nimi. Mamy podnieść sztandar i głosić ostatnie poselstwo ostrzeżenia ginącemu światu. Ci, którzy mają znajomość prawdy na ten czas, muszą teraz mocno trzymać sztandar z napisem: ‘Przykazania Boże i wiara Jezusa.’ }[Objawienie 14:12.]}[Lt232-1903.56; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.56&para=10197.66]


\egwnogap{I ask my ministering brethren to examine themselves, to see whether they are in the faith or not. \textbf{If they accept the spiritualistic representations made in Living Temple, their feet will soon be treading in forbidden paths. These representations are the Alpha of doctrines that would lead far away from the truth as we have received it from the Word of God}. \textbf{\underline{The acceptation of these sentiments will result in a weak, wavering faith}}. \textbf{If this is the teaching that is to be given in the medical missionary work, it will not be long before we have no foundation upon which to plant our feet}. \textbf{I am bidden to say that these erroneous sentiments are the sentiments of the wily foe} and should not be presented to any of our youth who are seeking to gain an education in medical missionary lines. For the sake of these youth, I speak decidedly.}[Lt232-1903.57; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.57&para=10197.67]


\egwnogap{Proszę moich braci w służbie, aby zbadali samych siebie, czy są w wierze, czy nie. \textbf{Jeśli przyjmą spirytualistyczne przedstawienia zawarte w Living Temple, ich stopy wkrótce będą stąpać po zakazanych ścieżkach. Te przedstawienia są Alfą doktryn, które prowadziłyby daleko od prawdy, jaką otrzymaliśmy ze Słowa Bożego}. \textbf{\underline{Akceptacja tych poglądów doprowadzi do słabej, chwiejnej wiary}}. \textbf{Jeśli to jest nauczanie, które ma być przekazywane w pracy misyjnej medycznej, to nie minie dużo czasu, a nie będziemy mieli fundamentu, na którym moglibyśmy postawić nasze stopy}. \textbf{Zostałam wezwana, aby powiedzieć, że te błędne poglądy są poglądami przebiegłego wroga} i nie powinny być przedstawiane żadnemu z naszych młodych ludzi, którzy starają się zdobyć wykształcenie w zakresie misji medycznej. Ze względu na tych młodych ludzi, mówię stanowczo.}[Lt232-1903.57; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.57&para=10197.67]


\egwnogap{\textbf{The expiring faith of the people of God \underline{must have a resurrection}}. \textbf{The exaltation of human reason has begun its work among us and has gone altogether too far}. Human reason is placed where divine, sanctifying truth should be. \textbf{Christ is waiting to kindle faith and love in the hearts of His people}. \textbf{Let not erroneous theories receive countenance from the people who ought to be standing firm on the platform of eternal truth.} \textbf{God calls upon us to hold firmly to \underline{the fundamental principles that are based upon unquestionable authority}}. He calls upon us to study the words and works of Christ, the greatest missionary that this world has ever known.}[Lt232-1903.58; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.58&para=10197.68]


\egwnogap{\textbf{Gasnąca wiara ludu Bożego \underline{musi mieć zmartwychwstanie}}. \textbf{Wywyższanie ludzkiego rozumu rozpoczęło swoją pracę wśród nas i zaszło zdecydowanie za daleko}. Ludzki rozum jest umieszczany tam, gdzie powinna być boska, uświęcająca prawda. \textbf{Chrystus czeka, aby rozpalić wiarę i miłość w sercach swojego ludu}. \textbf{Niech błędne teorie nie otrzymują poparcia od ludzi, którzy powinni stać mocno na platformie wiecznej prawdy.} \textbf{Bóg wzywa nas, abyśmy mocno trzymali się \underline{fundamentalnych zasad, które są oparte na niepodważalnym autorytecie}}. On wzywa nas do studiowania słów i dzieł Chrystusa, największego misjonarza, jakiego ten świat kiedykolwiek poznał.}[Lt232-1903.58; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.58&para=10197.68]


\egwnogap{\textbf{When the mind of a teacher of truth becomes in any way divorced from plain, self-denying gospel truth, he is prepared to receive fanciful sentiments called truth. Arrayed in the garments of light, these sentiments are presented to others, and too often they find favor. I am instructed to say to the members of our churches, Keep away from spiritualistic ideas}. \textbf{We are not dealing in fables}. God forbid that fables in the disguise of truth shall be presented to our people. \textbf{God forbid that any among us shall build upon the sand}.}[Lt232-1903.59; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.59&para=10197.69]


\egwnogap{\textbf{Kiedy umysł nauczyciela prawdy staje się w jakikolwiek sposób oddzielony od prostej, wymagającej zaparcia się siebie ewangelicznej prawdy, jest przygotowany do przyjęcia fantazyjnych poglądów nazywanych prawdą. Ubrane w szaty światła, te poglądy są przedstawiane innym i zbyt często znajdują uznanie. Zostałam pouczona, aby powiedzieć członkom naszych kościołów: Trzymajcie się z dala od spirytualistycznych idei}. \textbf{Nie zajmujemy się baśniami}. Nie daj Boże, aby baśnie w przebraniu prawdy były przedstawiane naszemu ludowi. \textbf{Nie daj Boże, aby ktokolwiek z nas budował na piasku}.}[Lt232-1903.59; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.59&para=10197.69]


\egwnogap{\textbf{The Lord has given us a clear, distinct message of truth for this time}. Let us proclaim this message. \textbf{Let us study the teaching of Christ}, and present what He has commanded us to present. \textbf{He who launches out in his own wisdom to preach strange things, which God has not given him, finds minds ready to be leavened with the new ideas that he has to present}. \textbf{Satan follows up the work that he does, and the efforts of the true servants of God are made much harder}. \textbf{The advancement of His cause is hindered, and His Spirit is grieved}.}[Lt232-1903.60; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.60&para=10197.70]


\egwnogap{\textbf{Pan dał nam jasne, wyraźne poselstwo prawdy na ten czas}. Głośmy to poselstwo. \textbf{Studiujmy nauczanie Chrystusa} i przedstawiajmy to, co On nakazał nam przedstawiać. \textbf{Ten, kto wyrusza w swojej własnej mądrości, aby głosić dziwne rzeczy, których Bóg mu nie dał, znajduje umysły gotowe do zakwaszenia nowymi ideami, które ma do przedstawienia}. \textbf{Szatan podąża za pracą, którą on wykonuje, a wysiłki prawdziwych sług Bożych stają się znacznie trudniejsze}. \textbf{Postęp Jego sprawy jest hamowany, a Jego Duch jest zasmucony}.}[Lt232-1903.60; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.60&para=10197.70]


We pray that God will speak decisively clear into everyone's heart reading these warnings, to keep from stepping off the foundation of our faith. God calls upon us to hold firmly to the \emcap{Fundamental Principles} that are based upon unquestionable authority.


Modlimy się, aby Bóg przemówił zdecydowanie i wyraźnie do serca każdego, kto czyta te ostrzeżenia, aby nie zszedł z fundamentu naszej wiary. Bóg wzywa nas, abyśmy mocno trzymali się \emcap{Fundamentalnych Zasad}, które opierają się na niepodważalnym autorytecie.


\input{lang/en/poems/chapter29}

% \input{lang/pl/poems/chapter29}