\chapter{The final call}


\chapter{Ostateczne wezwanie}


When we are acquainted with the \emcap{Fundamental Principles}, especially with the first point dealing with the \emcap{personality of God} and where His presence is, Sister White’s warnings suddenly shine in a new light. In light of the \emcap{Fundamental Principles}, and as a conclusion to these studies, we would like to present the following warnings:


Kiedy zapoznamy się z \emcap{Fundamentalnymi Zasadami}, szczególnie z pierwszym punktem dotyczącym \emcap{osobowości Boga} i tego, gdzie jest Jego obecność, ostrzeżenia Siostry White nagle jaśnieją w nowym świetle. W świetle \emcap{Fundamentalnych Zasad} i jako podsumowanie tych studiów, chcielibyśmy przedstawić następujące ostrzeżenia:


\egw{Satan presents his theories cautiously at first, and if he sees that his efforts are successful, he brings in theories that are still more misleading, \textbf{seeking to lead men and women away from \underline{the foundation principles that God designs shall be the safeguards of His people}}.}[Ms132-1903.40; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.40&para=9056.54]


\egw{Szatan przedstawia swoje teorie początkowo ostrożnie, a jeśli widzi, że jego wysiłki odnoszą sukces, wprowadza teorie, które są jeszcze bardziej zwodnicze, \textbf{starając się odwieść mężczyzn i kobiety od \underline{podstawowych zasad, które Bóg zamierzył jako zabezpieczenie dla swojego ludu}}.}[Ms132-1903.40; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.40&para=9056.54]


\egwnogap{\textbf{Let not our medical missionary workers accept theories that God has not given to any one}. \textbf{God will not excuse men for teaching theories that Christ has not taught}. He calls upon His army of workers to fall into line, taking their stand under the banner of truth. \textbf{He warns them to beware of occupying their time in the discussion of matters that God has not authorized any human being to discuss}.}[Ms132-1903.41; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.41&para=9056.55]


\egwnogap{\textbf{Niech nasi pracownicy misji medycznej nie przyjmują teorii, których Bóg nikomu nie dał}. \textbf{Bóg nie usprawiedliwi ludzi za nauczanie teorii, których Chrystus nie nauczał}. Wzywa swoją armię pracowników, aby ustawiła się w szeregu, zajmując stanowisko pod sztandarem prawdy. \textbf{Ostrzega ich, aby strzegli się poświęcania czasu na dyskusję o sprawach, których Bóg nie upoważnił żadnej ludzkiej istoty do omawiania}.}[Ms132-1903.41; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.41&para=9056.55]


\egwnogap{Let us put on every piece of the Christian armor and steadfastly resist the enemy. We shall have to meet fallen angels and the prince of the powers of darkness. Satan is by no means asleep; he is wide-awake and is playing the game of life for the souls of the people of God. He will come to them with flattery of all kinds, in the hope of leading them to swerve from their allegiance. \textbf{He desires to call their attention from the real issues to false theories}.}[Ms132-1903.42, 1903][https://egwwritings.org/?ref=en\_Ms132-1903.42&para=9056.56]


\egwnogap{Załóżmy każdą część chrześcijańskiej zbroi i stanowczo opierajmy się wrogowi. Będziemy musieli zmierzyć się z upadłymi aniołami i księciem mocy ciemności. Szatan wcale nie śpi; jest w pełni przebudzony i rozgrywa grę o życie dusz ludu Bożego. Przyjdzie do nich z wszelkiego rodzaju pochlebstwami, mając nadzieję, że doprowadzi ich do odchylenia się od wierności. \textbf{Pragnie odwrócić ich uwagę od prawdziwych kwestii ku fałszywym teoriom}.}[Ms132-1903.42, 1903][https://egwwritings.org/?ref=en\_Ms132-1903.42&para=9056.56]


\egwnogap{Ministers and physicians, sound an alarm. Call upon the people of God to be true and faithful. Be on your guard. Remember that as you co-operate with God, you have as your helpers angels that excel in strength. \textbf{Accept not the theories advanced by those who are not standing on the \underline{true foundation}, those who are charmed with that [of] which they do not know the true meaning}.}[Ms132-1903.43; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.43&para=9056.57]


\egwnogap{Kaznodzieje i lekarze, bijcie na alarm. Wzywajcie lud Boży, aby był prawdziwy i wierny. Bądźcie czujni. Pamiętajcie, że współpracując z Bogiem, macie jako pomocników aniołów, którzy przewyższają siłą. \textbf{Nie przyjmujcie teorii przedstawianych przez tych, którzy nie stoją na \underline{prawdziwym fundamencie}, tych, którzy są oczarowani tym, czego prawdziwego znaczenia nie znają}.}[Ms132-1903.43; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.43&para=9056.57]


\egwnogap{\textbf{\underline{Wake up}, my brethren, wake up and lift the danger signal}. \textbf{\underline{Sound the warning}}. \textbf{Let no man persuade you to accept theories that are opposed to the truths of God’s Word}. The servants of God have a solemn message to bear to this fallen, sin-cursed world. They are to hold aloft the banner on which is inscribed the words, ‘\textbf{The commandments of God and the faith of Jesus}.’ [Revelation 14:12.] \textbf{Those who are working in harmony with God will be of one heart and one mind}. With unflagging zeal they will proclaim the message, ‘Prepare to meet thy God.’ [Amos 4:12.] \textbf{They will not unite with worldlings, but will take their stand firmly in defense of \underline{the principles of truth}}.}[Ms132-1903.44; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.44&para=9056.58]


\egwnogap{\textbf{\underline{Obudźcie się}, moi bracia, obudźcie się i podnieście sygnał niebezpieczeństwa}. \textbf{\underline{Ogłoście ostrzeżenie}}. \textbf{Niech nikt nie przekona was do przyjęcia teorii, które są sprzeczne z prawdami Słowa Bożego}. Słudzy Boży mają poważne przesłanie do przekazania temu upadłemu, przeklętemu przez grzech światu. Mają trzymać wysoko sztandar, na którym wypisane są słowa: ‘\textbf{Przykazania Boże i wiara Jezusa}.’ [Objawienie 14:12.] \textbf{Ci, którzy działają w harmonii z Bogiem, będą jednego serca i jednego umysłu}. Z nieustającym zapałem będą głosić przesłanie: ‘Przygotuj się na spotkanie z twoim Bogiem.’ [Amos 4:12.] \textbf{Nie zjednoczą się ze światowcami, ale będą stanowczo bronić \underline{zasad prawdy}}.}[Ms132-1903.44; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.44&para=9056.58]


\egwnogap{Things must now be called by their right name. \textbf{Backsliding leaders are not to be honored as men who are holding their confidence firm}. God is in earnest with us. We are to sound the note of warning.}[Ms132-1903.45; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.45&para=9056.59]


\egwnogap{Rzeczy muszą być teraz nazywane po imieniu. \textbf{Odstępujący przywódcy nie powinni być honorowani jako ludzie, którzy trwają mocno w swojej wierze}. Bóg traktuje nas poważnie. Musimy ogłosić ton ostrzeżenia.}[Ms132-1903.45; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.45&para=9056.59]


\egwnogap{\textbf{Wake up, for Christ’s sake, wake up}. May God give success to those who are trying to arouse the sleepy watchmen. \textbf{Of many of those who profess to be shepherds of the flock, God says, ‘They are unfaithful}. \textbf{They have left their \underline{first love}}. \textbf{Unless they repent, I will come suddenly, and will remove their candlestick out of his place}.’ [See Revelation 2:4, 5.]}[Ms132-1903.46; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.47&para=9056.61]


\egwnogap{\textbf{Obudźcie się, dla Chrystusa, obudźcie się}. Niech Bóg da powodzenie tym, którzy próbują obudzić śpiących stróżów. \textbf{O wielu z tych, którzy twierdzą, że są pasterzami stada, Bóg mówi: ‘Są niewierni}. \textbf{Opuścili swoją \underline{pierwszą miłość}}. \textbf{Jeśli nie pokutują, przyjdę nagle i usunę ich świecznik z jego miejsca}.’ [Zobacz Objawienie 2:4, 5.]}[Ms132-1903.46; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.47&para=9056.61]


\egwnogap{Go to work now, without delay. How many judgment calls must the Lord make before His people cease to provoke Him to His face? Should He deal with them according to their backsliding, according to their worldliness and \textbf{to the way in which they have called darkness light and light darkness}, they would have no further calls to repentance, no more evidence or light to trifle with.}[Ms132-1903.47; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.47&para=9056.61]


\egwnogap{Zabierzcie się do pracy teraz, bez zwłoki. Ile wezwań sądowych musi Pan wydać, zanim Jego lud przestanie prowokować Go w twarz? Gdyby postępował z nimi zgodnie z ich odstępstwem, zgodnie z ich światowością i \textbf{ze sposobem, w jaki nazywali ciemność światłem, a światło ciemnością}, nie mieliby więcej wezwań do pokuty, nie mieliby więcej dowodów ani światła, którymi mogliby się bawić.}[Ms132-1903.47; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.47&para=9056.61]


\egwnogap{God’s people provoke Him by their idolatry and by their union with worldlings. He says, ‘My Spirit shall not always strive with men. I will not always bear with the perversity of those who lead souls from the narrow way into paths of uncertainty and falsehood. Unless those who have been often reproved make a decided change, they will be left to follow their own way.’ \textbf{His blessings will be taken from those who choose darkness rather than light, those who choose false guides rather than true. To those who disregard the evidence given them, making no difference between truth and error, the light bestowed will become darkness, and how great will be that darkness}.}[Ms132-1903.48; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.48&para=9056.62]


\egwnogap{Lud Boży prowokuje Go swoim bałwochwalstwem i swoim zjednoczeniem ze światowcami. On mówi: ‘Mój Duch nie zawsze będzie się spierał z ludźmi. Nie zawsze będę znosił przewrotność tych, którzy prowadzą dusze z wąskiej drogi na ścieżki niepewności i fałszu. Jeśli ci, którzy byli często upominani, nie dokonają zdecydowanej zmiany, zostaną pozostawieni, aby podążali własną drogą.’ \textbf{Jego błogosławieństwa zostaną zabrane tym, którzy wybierają ciemność zamiast światła, tym, którzy wybierają fałszywych przewodników zamiast prawdziwych. Dla tych, którzy lekceważą dane im dowody, nie czyniąc różnicy między prawdą a błędem, udzielone światło stanie się ciemnością, a jakże wielka będzie ta ciemność}.}[Ms132-1903.48; 1903][https://egwwritings.org/?ref=en\_Ms132-1903.48&para=9056.62]


May the Lord help us to wake up to His callings. Let us choose light rather than darkness. Let us plant our feet upon the platform of eternal Truth. Let us believe and obey the truths we first embraced regarding the personality of the Father and of the Son, and the blessings will follow.


Niech Pan pomoże nam przebudzić się na Jego wezwania. Wybierzmy światło zamiast ciemności. Postawmy nasze stopy na platformie wiecznej Prawdy. Wierzmy i bądźmy posłuszni prawdom, które najpierw przyjęliśmy odnośnie osobowości Ojca i Syna, a błogosławieństwa będą następować.


\egw{A liar is one that \textbf{presents false theories and doctrines}. \textbf{He who \underline{denies the personality of God and of His Son Jesus Christ} is \underline{denying God and Christ}}. ‘If that which ye have heard \textbf{\underline{from the beginning}} shall \textbf{remain in you}, ye \textbf{also shall continue in the Son and in the Father}.’ \textbf{If you continue to believe and obey \underline{the truths you first embraced regarding the personality of the Father and the Son}, you will be joined together with them in love}. \textbf{There will be seen that union for which Christ prayed just before His trial and crucifixion}.}[Ms23-1906.20; 1906][https://egwwritings.org/?ref=en\_Ms23-1906.20&para=7556.27]


\egw{Kłamcą jest ten, kto \textbf{przedstawia fałszywe teorie i doktryny}. \textbf{Ten, kto \underline{zaprzecza osobowości Boga i Jego Syna Jezusa Chrystusa} \underline{zaprzecza Bogu i Chrystusowi}}. ‘Jeśli to, co słyszeliście \textbf{\underline{od początku}} \textbf{pozostanie w was}, \textbf{wy również będziecie trwać w Synu i w Ojcu}.’ \textbf{Jeśli będziecie nadal wierzyć i być posłuszni \underline{prawdom, które najpierw przyjęliście odnośnie osobowości Ojca i Syna}, będziecie złączeni z nimi w miłości}. \textbf{Będzie widoczna ta jedność, o którą modlił się Chrystus tuż przed swoim procesem i ukrzyżowaniem}.}[Ms23-1906.20; 1906][https://egwwritings.org/?ref=en\_Ms23-1906.20&para=7556.27]


If we continue in the truth we first embraced regarding the personality of the Father and Son, we will be joined together with them in love. What a promise! We need this vital connection in our Christian life. Without exercising this union Christ prayed for, we are not able to fulfill the mission God has bestowed upon us as a prophetic movement.


Jeśli trwamy w prawdzie, którą najpierw przyjęliśmy odnośnie osobowości Ojca i Syna, będziemy złączeni z nimi w miłości. Cóż za obietnica! Potrzebujemy tego żywotnego połączenia w naszym chrześcijańskim życiu. Bez praktykowania tej jedności, o którą modlił się Chrystus, nie jesteśmy w stanie wypełnić misji, którą Bóg powierzył nam jako ruchowi proroczemu.


\egw{\textbf{What possibilities are opened up to the youth who lay hold of the divine assurances of God’s Word}! \textbf{Scarcely can the human mind comprehend what is the breadth and depth and height of the spiritual attainments that can be reached by becoming partakers of the divine nature}. The human agent who <daily> yields obedience to God, who becomes a partaker of the divine nature, finds pleasure <daily> in keeping the commandments of God, \textbf{for he is one with God}. \textbf{<It is essential that> he \underline{holds as vital a relation with God as does the Son to the Father}}. \textbf{He understands the oneness that Christ prayed might exist between the Father and the Son}.}[Lt43-1895.18; 1895][https://egwwritings.org/?ref=en\_Lt43-1895.18&para=5554.24]


\egw{\textbf{Jakie możliwości otwierają się przed młodzieżą, która chwyta się boskich zapewnień Słowa Bożego}! \textbf{Ludzki umysł z trudem może pojąć, jaka jest szerokość, głębokość i wysokość duchowych osiągnięć, które można zdobyć, stając się uczestnikami boskiej natury}. Ludzki przedstawiciel, który <codziennie> okazuje posłuszeństwo Bogu, który staje się uczestnikiem boskiej natury, znajduje przyjemność <codziennie> w przestrzeganiu przykazań Bożych, \textbf{ponieważ jest jedno z Bogiem}. \textbf{<Jest istotne, aby> \underline{utrzymywał tak żywotną relację z Bogiem, jak Syn z Ojcem}}. \textbf{Rozumie jedność, o którą modlił się Chrystus, aby mogła istnieć między Ojcem a Synem}.}[Lt43-1895.18; 1895][https://egwwritings.org/?ref=en\_Lt43-1895.18&para=5554.24]


God has laid before us our church’s current state of affairs. The renewal of the old foundation will be accomplished. The \emcap{Fundamental Principles} will be restored and the three angels’ messages will be proclaimed in their true light and power. The only decision we are to make is a decision of faithfulness to Him, regardless of the cost. Do you and I want to be part of it? The blessings will follow as we hold fast unto the hand of the all-knowing God and His leadings. \bible{But those things, which God before had shewed by the mouth of all his prophets,} He will fulfill! \bible{Repent ye therefore, and be converted, that your sins may be blotted out, \textbf{when the times of refreshing shall come from the presence of the Lord; And he shall send Jesus Christ}, which before was preached unto you.}[Acts 3:18-20]. God will soon pour out the Spirit of Christ, which was preached unto us. He will send the Spirit of His Son into our hearts \bible{crying, Abba, Father}[Galatians 4:6].  Then, enabled with the power on High, His people will proclaim, in a loud cry, the three angels’ messages. \bible{And this gospel of the kingdom shall be preached in all the world for a witness unto all nations; and then shall the end come}[Matthew 24:14]. Christ will descend from High on the clouds of Heaven, escorted by thousands upon thousands of His Hosts, and He will take us into His Father’s presence, that we may be there where He is, and we may behold His glory, the glory which He had with God before the world was (\textit{John 17: 24, 5}). He will take us in the presence of His Father, and the glory will be revealed unto us, \egwinline{the unapproachable splendor of their converse together}[Lt232-1903.54; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.54&para=10197.63]. Then God shall wipe away all tears from our eyes; \bible{and there shall be no more death, neither sorrow, nor crying, neither shall there be any more pain: for the former things are passed away.}[Revelation 21:4]. \bible{And there shall be no more curse: but \textbf{the throne of God and of the Lamb} shall be in it; and his servants shall serve him.}[Revelation 22:3] And then from the core of our being, in gratitude toward God for His unspeakable Gift, all saved \bible{and every creature which is in heaven, and on the earth, and under the earth, and such as are in the sea, and all that are in them,} shall proclaim \bible{Blessing, and honour, and glory, and power, be unto \textbf{Him that sitteth upon the throne, and unto the Lamb} for ever and ever.}[Revelation 5:13]


Bóg przedstawił nam obecny stan spraw naszego kościoła. Odnowienie starego fundamentu zostanie dokonane. \emcap{Fundamentalne Zasady} zostaną przywrócone, a poselstwa trzech aniołów będą głoszone w ich prawdziwym świetle i mocy. Jedyną decyzją, którą mamy podjąć, jest decyzja wierności Jemu, bez względu na koszty. Czy ty i ja chcemy być tego częścią? Błogosławieństwa będą następować, gdy będziemy mocno trzymać się ręki wszechwiedzącego Boga i Jego prowadzenia. \bible{Ale to, co Bóg wcześniej zapowiedział przez usta wszystkich swoich proroków,} On wypełni! \bible{Dlatego pokutujcie i nawróćcie się, aby wasze grzechy były zgładzone, \textbf{gdy nadejdą czasy ochłody od obecności Pana; I pośle Jezusa Chrystusa}, który był wam wcześniej głoszony.}[Dzieje 3:18-20]. Bóg wkrótce wyleje Ducha Chrystusa, który był nam głoszony. Pośle Ducha swego Syna do naszych serc \bible{wołającego: Abba, Ojcze}[Galacjan 4:6]. Wtedy, uzdolniony mocą z Wysokości, Jego lud będzie głosił, donośnym głosem, poselstwa trzech aniołów. \bible{A ta ewangelia królestwa będzie głoszona po całym świecie na świadectwo wszystkim narodom. I wtedy nadejdzie koniec}[Mateusz 24:14]. Chrystus zstąpi z Wysokości na obłokach nieba, eskortowany przez tysiące tysięcy Jego Zastępów, i zabierze nas do obecności swego Ojca, abyśmy byli tam, gdzie On jest, i abyśmy oglądali Jego chwałę, chwałę, którą miał u Boga, zanim powstał świat (\textit{Jan 17: 24, 5}). Zabierze nas do obecności swego Ojca, a chwała zostanie nam objawiona, \egwinline{niedostępny splendor ich rozmowy ze sobą}[Lt232-1903.54; 1903][https://egwwritings.org/?ref=en\_Lt232-1903.54&para=10197.63]. Wtedy Bóg otrze wszelką łzę z naszych oczu; \bible{i nie będzie już śmierci ani smutku, ani płaczu, ani bólu nie będzie, bo pierwsze rzeczy przeminęły.}[Objawienie 21:4]. \bible{I nie będzie już żadnego przekleństwa. \textbf{Tron Boga i Baranka} będzie w nim, a jego słudzy będą mu służyć.}[Objawienie 22:3] A wtedy z głębi naszej istoty, w wdzięczności wobec Boga za Jego niewypowiedziany Dar, wszyscy zbawieni \bible{i wszelkie stworzenie, które jest w niebie i na ziemi, i pod ziemią, i w morzu, i wszystko, co w nich jest,} ogłoszą \bible{Błogosławieństwo, cześć, chwała i moc \textbf{Temu, który siedzi na tronie, i Barankowi} na wieki wieków.}[Objawienie 5:13]


Amen!


Amen!


\input{lang/en/poems/chapter30}

% \input{lang/pl/poems/chapter30}