\chapter{The great apostasy is soon to be realized} \label{chap:apostasy}


\chapter{Wielkie odstępstwo wkrótce się urzeczywistni} \label{chap:apostasy}


In 1903, when the Living Temple was published and instigated the controversy over the \emcap{personality of God}, Sister White was faithfully obeying the command of the Great Commander. She was called by the words “\textit{Meet it!}” She faced this controversy by writing numerous letters to many people in the field. In these letters, we trace the prophetic insight of the future of the Seventh-day Adventist Church.


W 1903 roku, gdy opublikowano The Living Temple i wywołano spór o \emcap{osobowość Boga}, Siostra White wiernie wypełniała polecenie Wielkiego Dowódcy. Została wezwana słowami “\textit{Staw temu czoła!}” Stawiła czoła temu sporowi, pisząc liczne listy do wielu osób w terenie. W tych listach odnajdujemy prorocze spojrzenie na przyszłość Kościoła Adwentystów Dnia Siódmego.


One example is the correspondence between Sister White and her son William White. On November 26, 1905, there was a great Health Conference in College View Nebraska, where many medical missionary workers met together. William White was there and he had a short, 30-minute public talk. Afterwards, he wrote a letter to his mother regarding his impressions from the conference. Here is part of that letter:


Jednym z przykładów jest korespondencja między Siostrą White a jej synem Williamem White. 26 listopada 1905 roku odbyła się wielka Konferencja Zdrowia w College View w Nebrasce, gdzie spotkało się wielu pracowników misji medycznej. William White był tam i wygłosił krótkie, 30-minutowe przemówienie publiczne. Później napisał list do swojej matki dotyczący jego wrażeń z konferencji. Oto fragment tego listu:


\others{College View, Ne. – Tuesday, November 28, 1905; Author: William C. White} \\
\others{Nov. 28, 1905.} \\
\others{Mrs. E. G. White, Sanitarium, Cala.}


\others{College View, Ne. – Wtorek, 28 listopada 1905; Autor: William C. White} \\
\others{28 listopada 1905.} \\
\others{Pani E. G. White, Sanitarium, Cala.}


\othersnogap{...Sabbath morning I had opportunity to speak about thirty minutes. In my remarks I refered to the history of the Christian church. They began with pure principles, but through the attacks of Satan they became backslidden and departed from those principles. \textbf{I pointed out that the only hope for the S. D. A. church was to \underline{adhear to first principles}}. \textbf{I then referred to the order in which the enemy is attacking our work. His first effort was to destroy union and establish separation. His next work was to weaken our reverence for the Sabbath, then to weaken our faith in the Sanctuary service, then \underline{to break our confidence in the Spirit of Prophecy}, then to \underline{confuse our conception regarding a personal God}}.}[Letter from W. C. White to E. G. White, November 28, 1905.][http://ellenwhite.org/content/correspondence/incoming/43292pdf]


\othersnogap{...W sobotni poranek miałem okazję mówić przez około trzydzieści minut. W moich uwagach nawiązałem do historii kościoła chrześcijańskiego. Zaczęli od czystych zasad, ale przez ataki Szatana stali się odstępcami i odeszli od tych zasad. \textbf{Wskazałem, że jedyną nadzieją dla kościoła A.D.S. jest \underline{trzymanie się pierwszych zasad}}. \textbf{Następnie odniosłem się do kolejności, w jakiej wróg atakuje naszą pracę. Jego pierwszym wysiłkiem było zniszczenie jedności i ustanowienie podziału. Jego następnym dziełem było osłabienie naszego szacunku dla Sabatu, następnie osłabienie naszej wiary w służbę Świątynną, potem \underline{złamanie naszego zaufania do Ducha Proroctwa}, a następnie \underline{wprowadzenie zamieszania w nasze pojmowanie osobowego Boga}}.}[List od W. C. White'a do E. G. White, 28 listopada 1905.][http://ellenwhite.org/content/correspondence/incoming/43292pdf]


According to William White, our only hope as Seventh-day Adventists is to adhere to first principles. These principles, as we know, are the \emcap{Fundamental Principles}. Then, he referred to the order in which the enemy is attacking our work. The attack begins with our disunity, then aims to weaken our reverence for the Sabbath and the Sanctuary service, targets our confidence in the Spirit of Prophecy, and finally focuses on confusing our conceptions regarding personal God.


Według Williama White'a, naszą jedyną nadzieją jako Adwentystów Dnia Siódmego jest trzymanie się pierwszych zasad. Te zasady, jak wiemy, to \emcap{Fundamentalne Zasady}. Następnie odniósł się do kolejności, w jakiej wróg atakuje naszą pracę. Atak zaczyna się od naszego braku jedności, następnie dąży do osłabienia naszego szacunku dla Sabatu i służby Świątynnej, celuje w nasze zaufanie do Ducha Proroctwa, a w końcu skupia się na wprowadzeniu zamieszania w nasze pojmowanie osobowego Boga.


Sister White’s response to William White is of a startling nature. She hints to us that the great apostasy is soon to be realized, and that our hope is to adhere to the first principles of our faith—the \emcap{Fundamental Principles}.


Odpowiedź Siostry White do Williama White'a ma zaskakujący charakter. Sugeruje nam, że wielkie odstępstwo wkrótce się urzeczywistni, a naszą nadzieją jest trzymanie się pierwszych zasad naszej wiary—\emcap{Fundamentalnych Zasad}.


\egw{Elmshaven, St. Helena, California} \\
\egw{December 4, 1905} \\
\egw{W. C. White} \\
\egw{My dear son - }


\egw{Elmshaven, St. Helena, Kalifornia} \\
\egw{4 grudnia 1905} \\
\egw{W. C. White} \\
\egw{Mój drogi synu - }


\egw{...}


\egw{...}


\egw{“\textbf{One thing it is certain is soon to be realized—\underline{the great apostasy}, which is developing and increasing and waxing stronger and \underline{will continue} to do so until the Lord shall descend from heaven with a shout. \underline{We are to hold fast the first principles of our denominated faith} and go forward from strength to increased faith. \underline{Ever} we are to keep the faith that has been substantiated by the Holy Spirit of God \underline{from the earlier events of our experience until the present time}.} We need now larger breadth and deeper, more earnest, unwavering faith in the leadings of the Holy Spirit. \textbf{If we needed the manifest proof of the Holy Spirit’s power to confirm truth \underline{in the beginning}, after the passing of the time, \underline{we need today all the evidence in the confirmation of the truth}, when souls are departing from the faith and giving heed to seducing spirits and doctrines of devils.} There must not be any languishing of soul now. If ever there was a period of time when we needed the Holy Spirit’s power in our discourses, in our prayers, in every action proposed, it is now. \textbf{We are not to stop at the first experience, but while we bear \underline{the same message} to the people, \underline{this message is to be strengthened and enlarged}}. \textbf{We are to see and realize the importance of the message made certain by its divine origin}. We are to follow on to know the Lord, that we may know that His going forth is prepared as the morning. Our souls need the quickening from the Source of all power. \textbf{We may be strengthened and confirmed in the past experience \underline{that holds us to the essential points of truth which have made us what we are—Seventh-day Adventists}.}“}[Lt326-1905.2; 1905][https://egwwritings.org/?ref=en\_Lt326-1905.2&para=7678.8]


\egw{“\textbf{Jedna rzecz jest pewna, że wkrótce się urzeczywistni—\underline{wielkie odstępstwo}, które rozwija się, zwiększa i staje się silniejsze i \underline{będzie nadal} tak czynić, aż Pan zstąpi z nieba z okrzykiem. \underline{Musimy trzymać się pierwszych zasad naszej denominacyjnej wiary} i iść naprzód od siły do zwiększonej wiary. \underline{Zawsze} musimy zachować wiarę, która została potwierdzona przez Ducha Świętego Bożego \underline{od wcześniejszych wydarzeń naszego doświadczenia aż do obecnego czasu}.} Potrzebujemy teraz szerszego zakresu i głębszej, bardziej żarliwej, niezachwianej wiary w prowadzenie Ducha Świętego. \textbf{Jeśli potrzebowaliśmy wyraźnego dowodu mocy Ducha Świętego, aby potwierdzić prawdę \underline{na początku}, po upływie czasu, \underline{potrzebujemy dziś wszystkich dowodów w potwierdzeniu prawdy}, gdy dusze odchodzą od wiary i dają posłuch zwodniczym duchom i naukom demonów.} Nie może być teraz żadnego osłabienia duszy. Jeśli kiedykolwiek był okres, kiedy potrzebowaliśmy mocy Ducha Świętego w naszych przemówieniach, w naszych modlitwach, w każdym proponowanym działaniu, to jest to teraz. \textbf{Nie możemy zatrzymać się na pierwszym doświadczeniu, ale głosząc \underline{to samo poselstwo} ludziom, \underline{poselstwo to ma być wzmocnione i poszerzone}}. \textbf{Musimy dostrzec i zrozumieć znaczenie poselstwa, które stało się pewne dzięki jego boskiemu pochodzeniu}. Musimy dążyć do poznania Pana, abyśmy wiedzieli, że Jego wyjście jest przygotowane jak poranek. Nasze dusze potrzebują ożywienia ze Źródła wszelkiej mocy. \textbf{Możemy być wzmocnieni i utwierdzeni w przeszłym doświadczeniu, \underline{które trzyma nas przy istotnych punktach prawdy, które uczyniły nas tym, czym jesteśmy—Adwentystami Dnia Siódmego}.}”}[Lt326-1905.2; 1905][https://egwwritings.org/?ref=en\_Lt326-1905.2&para=7678.8]


\egwnogap{\textbf{The past fifty years have not dimmed one jot or principle of our faith as we received the great and wonderful evidences that were made certain to us in 1844, after the passing of the time.} The languishing souls are to be confirmed and quickened according to His Word. And many of the ministers of the gospel and the Lord’s physicians will have their languishing souls quickened according to the Word. \textbf{\underline{Not a word is changed or denied}.} \textbf{That which the Holy Spirit testified to as truth after the passing of the time, in our great disappointment, is \underline{the solid foundation of truth}. Pillars of truth were revealed, and we accepted \underline{the foundation principles} that have made us what we are—Seventh-day Adventists, keeping the commandments of God and having the faith of Jesus.}}[Lt326-1905.3; 1905][https://egwwritings.org/?ref=en\_Lt326-1905.3&para=7678.9]


\egwnogap{\textbf{Ostatnie pięćdziesiąt lat nie przyćmiło ani jednej joty czy zasady naszej wiary, którą przyjęliśmy wraz z wielkimi i wspaniałymi dowodami, które stały się dla nas pewne w 1844 roku, po upływie czasu.} Osłabione dusze mają być wzmocnione i ożywione zgodnie z Jego Słowem. I wielu kaznodziejów ewangelii oraz lekarzy Pańskich będzie miało swoje osłabione dusze ożywione zgodnie ze Słowem. \textbf{\underline{Ani jedno słowo nie zostało zmienione ani zaprzeczone}.} \textbf{To, co Duch Święty poświadczył jako prawdę po upływie czasu, w naszym wielkim rozczarowaniu, jest \underline{solidnym fundamentem prawdy}. Filary prawdy zostały objawione i przyjęliśmy \underline{podstawowe zasady}, które uczyniły nas tym, kim jesteśmy - Adwentystami Dnia Siódmego, zachowującymi przykazania Boże i mającymi wiarę Jezusa.}}[Lt326-1905.3; 1905][https://egwwritings.org/?ref=en\_Lt326-1905.3&para=7678.9]


This letter is startling because it is an answer to the order of how the enemy is attacking our work. Sister White is well aware of these attacks and she presented the problem in its correct light, also showing us what we shall do to prevent Satan's attacks on us. The enemy wants to \others{confuse our conception regarding a personal God}. This is the very point of great apostasy that \egwinline{is soon to be realized}, and has been \egwinline{developing and increasing and waxing stronger and will continue to do so until the Lord shall descend from heaven with a shout}. This is the apostasy we are experiencing today. What is our hope against this deception and great apostasy? \egwinline{\textbf{\underline{We are to hold fast the first principles of our denominated faith} and go forward from strength to increased faith. \underline{Ever} we are to keep the faith that has been substantiated by the Holy Spirit of God from the earlier events of our experience until the present time.}} \egwinline{...\textbf{\underline{this message is to be strengthened and enlarged}}...} \egwinline{...\textbf{\underline{we need today all the evidence in the confirmation of the truth}}...} \egwinline{\textbf{We may be strengthened and confirmed in the past experience that holds us to the essential points of truth which have made us what we are—Seventh-day Adventists}}. These essential points of truth, which have made us Seventh-day Adventists, are the \emcap{Fundamental Principles}, born in the beginning of our work. In 1905, she wrote, \egwinline{\textbf{The past fifty years have not dimmed one jot or principle of our faith as we received the great and wonderful evidences that were made certain to us in 1844, after the passing of the time.}} \egwinline{\textbf{Not a word is changed or denied.} \textbf{That which the Holy Spirit testified to as truth after the passing of the time, in our great disappointment, is \underline{the solid foundation of truth}. Pillars of truth were revealed, and we accepted \underline{the foundation principles} that have made us what we are—Seventh-day Adventists, keeping the commandments of God and having the faith of Jesus.}}


Ten list jest zaskakujący, ponieważ jest odpowiedzią na sposób, w jaki wróg atakuje naszą pracę. Siostra White jest dobrze świadoma tych ataków i przedstawiła problem we właściwym świetle, pokazując nam również, co powinniśmy zrobić, aby zapobiec atakom Szatana na nas. Wróg chce \others{zaciemnić nasze pojęcie o osobowym Bogu}. To jest właśnie punkt wielkiego odstępstwa, które \egwinline{wkrótce się urzeczywistni}, i które \egwinline{rozwija się, zwiększa i rośnie w siłę, i będzie to kontynuować, aż Pan zstąpi z nieba z okrzykiem}. To jest odstępstwo, którego doświadczamy dzisiaj. Jaka jest nasza nadzieja przeciwko temu zwiedzeniu i wielkiemu odstępstwu? \egwinline{\textbf{\underline{Mamy mocno trzymać się pierwszych zasad naszej denominacyjnej wiary} i iść naprzód od siły do zwiększonej wiary. \underline{Zawsze} musimy zachować wiarę, która została potwierdzona przez Ducha Świętego Bożego od wcześniejszych wydarzeń naszego doświadczenia aż do obecnego czasu.}} \egwinline{...\textbf{\underline{to poselstwo ma być wzmocnione i poszerzone}}...} \egwinline{...\textbf{\underline{potrzebujemy dziś wszystkich dowodów w potwierdzeniu prawdy}}...} \egwinline{\textbf{Możemy być wzmocnieni i utwierdzeni w przeszłym doświadczeniu, które trzyma nas przy istotnych punktach prawdy, które uczyniły nas tym, kim jesteśmy - Adwentystami Dnia Siódmego}}. Te istotne punkty prawdy, które uczyniły nas Adwentystami Dnia Siódmego, to \emcap{Fundamentalne Zasady}, zrodzone na początku naszej pracy. W 1905 roku napisała, \egwinline{\textbf{Ostatnie pięćdziesiąt lat nie przyćmiło ani jednej joty czy zasady naszej wiary, którą przyjęliśmy wraz z wielkimi i wspaniałymi dowodami, które stały się dla nas pewne w 1844 roku, po upływie czasu.}} \egwinline{\textbf{Ani jedno słowo nie zostało zmienione ani zaprzeczone.} \textbf{To, co Duch Święty poświadczył jako prawdę po upływie czasu, w naszym wielkim rozczarowaniu, jest \underline{solidnym fundamentem prawdy}. Filary prawdy zostały objawione i przyjęliśmy \underline{podstawowe zasady}, które uczyniły nas tym, kim jesteśmy - Adwentystami Dnia Siódmego, zachowującymi przykazania Boże i mającymi wiarę Jezusa.}}


God calls us to be steadfast in the \emcap{Fundamental Principles}, especially over the \others{conception regarding a personal God}. This is the first point of the \emcap{Fundamental Principles}.


Bóg wzywa nas do wytrwałości w \emcap{Fundamentalnych Zasadach}, szczególnie w kwestii \others{pojęcia o osobowym Bogu}. Jest to pierwszy punkt \emcap{Fundamentalnych Zasad}.


Sister White foretold that a great apostasy is developing in our church regarding the understanding of the \emcap{personality of God}. The true understanding of the \emcap{personality of God} is presented in the \emcap{Fundamental Principles}. She clearly warned us of Satan’s attack on these principles. She calls us to \egw{\textbf{hold fast the first principles of our denominated faith} and go forward from strength to increased faith}.


Siostra White przepowiedziała, że w naszym kościele rozwija się wielkie odstępstwo dotyczące zrozumienia \emcap{osobowości Boga}. Prawdziwe zrozumienie \emcap{osobowości Boga} jest przedstawione w \emcap{Fundamentalnych Zasadach}. Wyraźnie ostrzegała nas przed atakiem Szatana na te zasady. Wzywa nas, abyśmy \egw{\textbf{mocno trzymali się pierwszych zasad naszej denominacyjnej wiary} i szli naprzód od siły do zwiększonej wiary}.


\egw{\textbf{“After the passing of the time, God entrusted to His faithful followers the precious \underline{principles of present truth}. These principles were not given to those who had had no part in the giving of the first and second angels’ messages. They were given to the workers who had had a part in the cause from the beginning}.”}[Ms129-1905.5; 1905][https://egwwritings.org/?ref=en\_Ms129-1905.5&para=9797.12]


\egw{\textbf{“Po upływie czasu, Bóg powierzył swoim wiernym naśladowcom cenne \underline{zasady teraźniejszej prawdy}. Te zasady nie zostały dane tym, którzy nie mieli udziału w głoszeniu poselstw pierwszego i drugiego anioła. Zostały one dane pracownikom, którzy mieli udział w tej sprawie od początku}.”}[Ms129-1905.5; 1905][https://egwwritings.org/?ref=en\_Ms129-1905.5&para=9797.12]


\egwnogap{\textbf{Those who passed through these experiences are to be as \underline{firm as a rock to the principles} that have made us Seventh-day Adventists}. They are to be workers together with God, binding up the testimony and sealing the law among His disciples. Those who took part in the establishment of our work upon the foundation of Bible truth; \textbf{those who know the waymarks that have pointed out the right path} are to be regarded as workers of the highest value. They can speak from personal experience, regarding the truths entrusted to them. These men are not to permit their faith to be changed to infidelity; they are not to permit the banner of the third angel to be taken from their hands. They are to hold the beginning of their confidence firm unto the end. \textbf{\underline{The Lord has declared that the history of the past shall be rehearsed as we enter upon the closing work}. Every truth that He has given for these last days is to be proclaimed to the world. \underline{Every pillar} that He has established \underline{is to be strengthened}. We cannot now step off the foundation that God has established. We cannot now enter into any new organization; for this would mean apostasy from the truth}.}[Ms129-1905.6; 1905][https://egwwritings.org/?ref=en\_Ms129-1905.6&para=9797.13]


\egwnogap{\textbf{Ci, którzy przeszli przez te doświadczenia, mają być \underline{mocni jak skała wobec zasad}, które uczyniły nas Adwentystami Dnia Siódmego}. Mają być współpracownikami Boga, wiążąc świadectwo i pieczętując prawo wśród Jego uczniów. Ci, którzy brali udział w ustanawianiu naszej pracy na fundamencie biblijnej prawdy; \textbf{ci, którzy znają znaki, które wskazały właściwą ścieżkę}, mają być uważani za pracowników najwyższej wartości. Mogą mówić z osobistego doświadczenia o powierzonych im prawdach. Ci ludzie nie powinni pozwolić, aby ich wiara zmieniła się w niewiarę; nie powinni pozwolić, aby sztandar trzeciego anioła został zabrany z ich rąk. Mają trzymać początek swojej ufności mocno aż do końca. \textbf{\underline{Pan oświadczył, że historia przeszłości zostanie powtórzona, gdy wkraczamy w końcowe dzieło}. Każda prawda, którą dał na te ostatnie dni, ma być ogłoszona światu. \underline{Każdy filar}, który ustanowił, \underline{ma być wzmocniony}. Nie możemy teraz zejść z fundamentu, który Bóg ustanowił. Nie możemy teraz wejść w żadną nową organizację; oznaczałoby to odstępstwo od prawdy}.}[Ms129-1905.6; 1905][https://egwwritings.org/?ref=en\_Ms129-1905.6&para=9797.13]


Stepping off of the foundation that God has established means entering into new organization; this is apostasy from the truth. Comparing the \emcap{Fundamental Principles} of the past with current trinitarian Fundamental Beliefs, it’s evident we’re in a state of apostasy.  Ellen White prophesied that this apostasy will be \egwinline{\textbf{developing and increasing and waxing stronger and \underline{will continue} to do so until the Lord shall descend from heaven with a shout}}[Lt326-1905.2; 1905][https://egwwritings.org/?ref=en\_Lt326-1905.2&para=7678.8].


Zejście z fundamentu, który Bóg ustanowił, oznacza wejście w nową organizację; to jest odstępstwo od prawdy. Porównując \emcap{Fundamentalne Zasady} z przeszłości z obecnymi trynitarnymi Fundamentalnymi Wierzeniami, widać wyraźnie, że znajdujemy się w stanie odstępstwa. Ellen White przepowiedziała, że to odstępstwo będzie \egwinline{\textbf{rozwijać się, zwiększać i rosnąć w siłę, i \underline{będzie kontynuować} to aż Pan zstąpi z nieba z okrzykiem}}[Lt326-1905.2; 1905][https://egwwritings.org/?ref=en\_Lt326-1905.2&para=7678.8].


\input{lang/en/poems/chapter23}

% \input{lang/pl/poems/chapter23}