\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter9}{The proper solution}


\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter9}{La solución adecuada}


God has given Sister White presentations regarding the deceptions that Satan is bringing into our church. She was instructed that we should read to and understand the works of our pioneers in order to combat these deceptions.


Dios ha dado a la hermana White presentaciones sobre los engaños que Satanás está introduciendo en nuestra iglesia. Se le instruyó que deberíamos leer y entender las obras de nuestros pioneros a fin de combatir estos engaños.


\egw{\textbf{I have had presentations regarding the deceptions that Satan is \underline{bringing in at this time}}. \textbf{I have been instructed} that \textbf{we should make prominent the testimony of some of the \underline{old workers} who are now dead}. \textbf{Let them continue to speak through their articles as found \underline{in the early numbers of our papers}}. \textbf{These articles should now be reprinted, that there may be \underline{a living voice from the Lord’s witnesses}}. \textbf{\underline{The history of the early experiences in the message will be a power to withstand the masterly ingenuity of Satan’s deceptions}.} This instruction has been repeated recently. I must present before the people \textbf{the testimonies of Bible truth and repeat the decided messages given years ago}...}[Lt99-1905.7; 1905][https://egwwritings.org/read?panels=p9997.13]


\egw{\textbf{He tenido presentaciones con respecto a los engaños que Satanás está \underline{trayendo en este tiempo}}. \textbf{Se me ha instruido} que \textbf{debemos hacer prominente el testimonio de algunos de los \underline{antiguos obreros} que ahora están muertos}. \textbf{Dejemos que continúen hablando a través de sus artículos como se encuentran \underline{en los primeros números de nuestros periódicos}}. \textbf{Estos artículos deben ser reimpresos ahora, para que haya \underline{una voz viva de los testigos del Señor}}. \textbf{\underline{La historia de las primeras experiencias en el mensaje será un poder para resistir el ingenio magistral de los engaños de Satanás}.} Esta instrucción se ha repetido recientemente. Debo presentar ante el pueblo \textbf{los testimonios de la verdad bíblica y repetir los mensajes decididos dados hace años}...}[Lt99-1905.7; 1905][https://egwwritings.org/read?panels=p9997.13]


\egw{\textbf{When men come in who would move one pin or pillar from the foundation} \textbf{which God has established by His Holy Spirit}, \textbf{let the} \textbf{\underline{aged men who were pioneers} in our work \underline{speak plainly}, and let those who are dead \underline{speak also by reprinting of their articles in our periodicals}}. Gather up the rays of divine light that God has given \textbf{as He has led His people on step by step in the way of truth}. \textbf{This truth will stand the test of time and trial}.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


\egw{\textbf{Cuando vengan hombres que quieran mover una clavija o pilar del fundamento} \textbf{que Dios ha establecido por su Espíritu Santo}, \textbf{que} \textbf{\underline{los ancianos que fueron pioneros} en nuestra obra \underline{hablen claramente}, y que hablen también los que han muerto \underline{reimprimiendo sus artículos en nuestros periódicos}}. Recojan los rayos de la luz divina que Dios ha dado \textbf{al guiar a su pueblo paso a paso en el camino de la verdad}. \textbf{Esta verdad resistirá la prueba del tiempo y del juicio}.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


We have seen that the Seventh-day Adventist Church has experienced the change in our foundation regarding the \emcap{personality of God}. We have been told that the first truth received by our pioneers will stand the test of time and trial. Has the truth on the \emcap{personality of God}, which we’ve received first, failed the test of time?


Hemos visto que la Iglesia Adventista del Séptimo Día ha experimentado el cambio en nuestro fundamento con respecto a la \emcap{personalidad de Dios}. Se nos ha dicho que la primera verdad recibida por nuestros pioneros resistirá la prueba del tiempo y del juicio. ¿Ha fallado la verdad sobre la \emcap{personalidad de Dios}, que hemos recibido primero, la prueba del tiempo?


\egw{\textbf{When the power of God testifies to what is truth, the truth is to stand forever as the truth}. \textbf{No after suppositions, contrary to the light God has given are to be entertained}. Men will arise with interpretations of Scripture which are to them truth, but which are not truth. \textbf{The truth for this time, God has given us as \underline{a foundation for our faith}.} \textbf{He Himself has taught us what is truth.} One will arise and still another with new light which \textbf{contradicts }the light that God has given under the demonstration of His Holy Spirit. \textbf{A few are still alive who passed through the experience gained in the establishment of this truth}. God has graciously spared their lives\textbf{ to repeat and repeat} till the close of their lives, \textbf{the experience through which they passed} even as did John the apostle till the very close of his life. \textbf{And the standard bearers who have fallen in death, are to \underline{speak through the reprinting of their writings}. \underline{I am instructed} that thus [their] voices are to be heard}. \textbf{\underline{They are to bear their testimony as to what constitutes the truth for this time.}} \textbf{We are not to receive the words of those who come with a message \underline{that contradicts the special points of our faith}}. They gather together a mass of Scripture, and pile it as proof around their asserted theories. This has been done over and over again during the past fifty years. And while the Scriptures are God’s Word, and are to be respected, the application of them,\textbf{ if such application moves one pillar of the foundation that God has sustained these fifty years, is a great mistake}. He who makes such an application knows not the wonderful demonstration of the Holy Spirit that gave power and force to the past messages that have come to the people of God.}[PH020 14.2; 1905][https://egwwritings.org/read?panels=p276.89]


\egw{\textbf{Cuando el poder de Dios da testimonio de lo que es verdad, la verdad debe permanecer para siempre como la verdad}. \textbf{No se deben sostener suposiciones posteriores, contrarias a la luz que Dios ha dado}. Los hombres se levantarán con interpretaciones de la Escritura que para ellos son la verdad, pero que no son la verdad. \textbf{La verdad para este tiempo, Dios nos la ha dado como \underline{fundamento de nuestra fe}.} \textbf{Él mismo nos ha enseñado lo que es la verdad.} Se levantará uno y otro con nueva luz que \textbf{contradice }la luz que Dios ha dado bajo la demostración de su Espíritu Santo. \textbf{Todavía viven algunos que pasaron por la experiencia adquirida en el establecimiento de esta verdad}. Dios les ha perdonado graciosamente la vida\textbf{ para que repitan y repitan} hasta el final de sus vidas, \textbf{la experiencia por la que pasaron} como lo hizo el apóstol Juan hasta el final de su vida. \textbf{Y los abanderados que han caído en la muerte, deben \underline{hablar a través de la reimpresión de sus escritos}. \underline{Se me ha instruido} que así [sus] voces se escuchen}. \textbf{\underline{Ellos deben dar su testimonio sobre lo que constituye la verdad para este tiempo.}} \textbf{No debemos recibir las palabras de aquellos que vienen con un mensaje \underline{que contradice los puntos especiales de nuestra fe}}. Ellos reúnen una masa de Escritura, y la apilan como prueba alrededor de sus teorías afirmadas. Esto se ha hecho una y otra vez durante los últimos cincuenta años. Y aunque las Escrituras son la Palabra de Dios, y deben ser respetadas, la aplicación de las mismas,\textbf{ si tal aplicación mueve un pilar del fundamento que Dios ha sostenido estos cincuenta años, es un gran error}. El que hace tal aplicación no conoce la maravillosa demostración del Espíritu Santo que dio poder y fuerza a los mensajes pasados que han llegado al pueblo de Dios.}[PH020 14.2; 1905][https://egwwritings.org/read?panels=p276.89]


Let us become acquainted with the writing of our pioneers regarding the first point of the \emcap{Fundamental Principles}. Let’s become acquainted with the truth the pioneers wrote about the \emcap{personality of God} and where His presence is. Let us, \egwinline{gather up the rays of divine light that God has given as He has led His people on step by step in the way of truth.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


Conozcamos los escritos de nuestros pioneros con respecto al primer punto de los \emcap{Principios Fundamentales}. Conozcamos la verdad que los pioneros escribieron sobre la \emcap{personalidad de Dios} y dónde está su presencia. \egwinline{Recojamos los rayos de luz divina que Dios ha dado al guiar a su pueblo paso a paso en el camino de la verdad.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


\input{lang/en/poems/chapter9}


\input{lang/en/poems/chapter9}
