I've examined the text and found a few grammar issues to fix:

```
# Corrected subject-verb agreement
The \egwinline{principles of truth} that God has revealed \egwinline{is our only true foundation}.
->
The \egwinline{principles of truth} that God has revealed \egwinline{are our only true foundation}.
---------

# Corrected preposition usage
This movement started about the time that Dr. John H. Kellogg published the book, "Living Temple".
->
This movement started about the time when Dr. John H. Kellogg published the book, "Living Temple".
---------

# Removed unnecessary comma
The reasoning of Dr. Kellogg regarding these questions, expressed in the book Living Temple, is\egwinline{unsound}.
->
The reasoning of Dr. Kellogg regarding these questions expressed in the book Living Temple is\egwinline{unsound}.
---------

# Corrected punctuation in quotation
\others{"I – That there is \textbf{one God}, \textbf{\underline{a personal, spiritual being}}, \textbf{the creator of all things}, omnipotent, omniscient, and eternal, infinite in wisdom, holiness, justice, goodness, truth, and mercy; unchangeable, and \textbf{\underline{everywhere present by his representative, the Holy Spirit}}. Ps. 139:7."}
->
\others{"I – That there is \textbf{one God}, \textbf{\underline{a personal, spiritual being}}, \textbf{the creator of all things}, omnipotent, omniscient, and eternal, infinite in wisdom, holiness, justice, goodness, truth, and mercy; unchangeable, and \textbf{\underline{everywhere present by his representative, the Holy Spirit}}. Ps. 139:7."}
---------

# Corrected punctuation in citation
\othersnogap{II – That there is \textbf{one Lord Jesus Christ, }\textbf{\underline{the Son of the Eternal Father}}, the one \textbf{\underline{by}}\textbf{ whom God created all things}, and by whom they do consist; …"}[The Fundamental Principles 1889, point no. 1.,2.,.] 
->
\othersnogap{II – That there is \textbf{one Lord Jesus Christ, }\textbf{\underline{the Son of the Eternal Father}}, the one \textbf{\underline{by}}\textbf{ whom God created all things}, and by whom they do consist; …"}[The Fundamental Principles 1889, point no. 1, 2.]
---------
```