I've examined the text and found a few grammar issues to correct:

```
# Corrected verb tense from “dispeling” to “dispelling”
In most of her quotations on this subject, we see her dispeling the Trinitarian error, rather than pantheistic error.
->
In most of her quotations on this subject, we see her dispelling the Trinitarian error, rather than pantheistic error.
---------

# Removed apostrophe from “chapters’“ as it's not possessive
And it is not only John 17, but also chapters’ 13-16.
->
And it is not only John 17, but also chapters 13-16.
---------

# Removed extra closing bracket
\egw{\textbf{\underline{…study the thirteenth, fourteenth, fifteenth, sixteenth, and seventeenth chapters of John}. The words of these chapters explain themselves. ‘This is life eternal,’ Christ declared, ‘that they might know \underline{Thee the only true God}, and Jesus Christ, whom Thou hast sent.’ \underline{In these words the personality of God and of His Son is clearly spoken of.} \underline{The personality of the one does not do away with the necessity for the personality of the other}.}}
->
\egw{\textbf{\underline{…study the thirteenth, fourteenth, fifteenth, sixteenth, and seventeenth chapters of John}. The words of these chapters explain themselves. ‘This is life eternal,’ Christ declared, ‘that they might know \underline{Thee the only true God}, and Jesus Christ, whom Thou hast sent.’ \underline{In these words the personality of God and of His Son is clearly spoken of.} \underline{The personality of the one does not do away with the necessity for the personality of the other}.}
---------

# Removed duplicate paragraph
\egwnogap{Wonderful statement! \textbf{The unity that exists between Christ and His disciples \underline{does not destroy the personality of either}. They are one in purpose, in mind, in character, but \underline{not in person}. It is thus that God and Christ are one}.}[8T 269.4; 1904][https://egwwritings.org/?ref=en\_8T.269.4&para=112.1563]
->
[Remove this duplicate paragraph as it appears twice in succession]
---------
```