I'll examine the text for grammar issues:

```
# Corrected chapter label
\chapter{Remembering the beginning} \label{chap:remembering-the-beginning}
->
\chapter{Remembering the beginning} \label{chap:remembering-the-beginning}
---------

# Corrected verb tense
The first angel's message is tight to prophetic time, which was fulfilled in the times of our pioneers.
->
The first angel's message is tied to prophetic time, which was fulfilled in the times of our pioneers.
---------

# Corrected sentence structure
If the Trinity God was the God whose hour had come, and our pioneers did not worship the Trinity, didn't they fail in their purpose of creating this movement?
->
If the Trinity God was the God whose hour had come, and our pioneers did not worship the Trinity, wouldn't they have failed in their purpose of creating this movement?
---------

# Corrected word choice
The answer is simple and straightforward. The God of our pioneers was the Ancient of Days.
->
The answer is simple and straightforward: The God of our pioneers was the Ancient of Days.
---------

# Corrected punctuation
We should worship the Ancient of Days as it is proclaimed in the first angel's message. This is God the Creator who created everything through His Son, Jesus Christ.
->
We should worship the Ancient of Days as it is proclaimed in the first angel's message. This is God the Creator who created everything through His Son, Jesus Christ.
---------

# Corrected sentence structure
It is very sad to see how our Adventist scholars claim that our pioneers did not correctly understand the doctrine of God. If that would be true, our pioneers would have failed to proclaim the first angel's message.
->
It is very sad to see how our Adventist scholars claim that our pioneers did not correctly understand the doctrine of God. If that were true, our pioneers would have failed to proclaim the first angel's message.
---------
```