I've examined the text and found a few grammar issues to correct:

```
# Changed “In Sister White's reply” to “In Sister White's reply,”
In Sister White's reply to Dr. Kellogg's belief on the Trinity doctrine and his attempts to \textit{patch up} the Living Temple, we see that she viewed the Trinity doctrine as contradicting the light given her regarding \emcap{the personality of God}.
->
In Sister White's reply, to Dr. Kellogg's belief on the Trinity doctrine and his attempts to \textit{patch up} the Living Temple, we see that she viewed the Trinity doctrine as contradicting the light given her regarding \emcap{the personality of God}.
---------

# Changed “Currently, they do not represent our beliefs.” to “Currently, these do not represent our beliefs.”
According to Sister White's testimony, the foundation of our faith was the \emcap{Fundamental Principles}. Currently, they do not represent our beliefs.
->
According to Sister White's testimony, the foundation of our faith was the \emcap{Fundamental Principles}. Currently, these do not represent our beliefs.
---------

# Changed “what He is, robbed” to “what He is, has robbed”
From the light and the experiences of how God established the first point of the \emcap{Fundamental Principles}, does the newly formed doctrine about who God is and what He is, robbed the people of God of their past experience?
->
From the light and the experiences of how God established the first point of the \emcap{Fundamental Principles}, does the newly formed doctrine about who God is and what He is, has robbed the people of God of their past experience?
---------

# Changed “Could it be that by adherence” to “Could it be that by adhering”
Could it be that by adherence to the Trinity doctrine we have torn down this particular pillar of our faith?
->
Could it be that by adhering to the Trinity doctrine we have torn down this particular pillar of our faith?
---------
```