\chapter{Reply to Kellogg’s trinitarian sentiments}

If we look at the Kellogg crisis through the perspective of the \emcap{personality of God} and the \emcap{Fundamental Principles}, Sister White’s quotations inevitably shine in a new light. In this light we see the conflict between the truth we have received in the beginning, on the \emcap{personality of God}, and the Trinity doctrine. In order to avoid discrepancy, in the interest of defending the Trinity doctrine, scholars always overemphasize the pantheistic side of the problem.

We would like to challenge this tendency to overemphasize the pantheistic side of Kellogg’s controversy. Sister White generally wrote proactive truth; she approached the error by uplifting the truth. This is why she wrote so much about the \emcap{personality of God}. In most of her quotations on this subject, we see her dispelling the Trinitarian error, rather than pantheistic error. We read one such example where she establishes the truth on the \emcap{personality of God} referencing the seventeenth chapter of John.

\egw{\textbf{The personality of the Father and the Son, also the unity that exists between Them, are presented in the seventeenth chapter of John}, in the prayer of Christ for His disciples:}[MH 421.7; 1905][https://egwwritings.org/?ref=en\_MH.421.7&para=135.2173]

There are many cases where Sister White quotes John 17 in regard to Kellogg’s crisis. Those who assert that Kellogg’s crisis was solely about pantheism should inquire how John 17 addresses God in nature. And it is not only John 17, but also chapters 13-16. In her letter to Kellogg, she wrote:

\egw{\textbf{\underline{…study the thirteenth, fourteenth, fifteenth, sixteenth, and seventeenth chapters of John}. The words of these chapters explain themselves. ‘This is life eternal,’ Christ declared, ‘that they might know \underline{Thee the only true God}, and Jesus Christ, whom Thou hast sent.’ \underline{In these words the personality of God and of His Son is clearly spoken of.} \underline{The personality of the one does not do away with the necessity for the personality of the other}.}}[Lt232-1903.48, 1903][https://egwwritings.org/?ref=en\_Lt232-1903.48&para=10197.57]

In the aforementioned chapters of John, John did not reference anything pertaining to God in nature. The content of those chapters covers who is the only true God, how the Father and the Son are one, their true relation, and how Jesus can be everywhere present yet will ascend to the Father.

\egw{Jesus said to the Jews: ‘My Father worketh hitherto, and I work.... The Son can do nothing of Himself, but what He seeth the Father do: for what things soever He doeth, these also doeth the Son likewise. For the Father loveth the Son, and showeth Him all things that Himself doeth.’ John 5:17-20.}[8T 268.4, 1904][https://egwwritings.org/?ref=en\_8T.268.4&para=112.1557]

\egwnogap{\textbf{Here again is brought to view the \underline{personality of the Father and the Son}, showing the unity that exists between them}.}[8T 269.1; 1904][https://egwwritings.org/?ref=en\_8T.269.1&para=112.1560]

\egwnogap{\textbf{This unity is expressed also in \underline{the seventeenth chapter of John}}, in the prayer of Christ for His disciples:}[8T 269.2; 1904][https://egwwritings.org/?ref=en\_8T.269.2&para=112.1561]

\egwnogap{‘Neither pray I for these alone, but for them also which shall believe on Me through their word; that they all may be one; \textbf{as Thou, Father, art in Me, and I in Thee, that they also may be one in Us}: that the world may believe that Thou hast sent Me. And \textbf{the glory which Thou gavest Me} I have given them; \textbf{that they may be one, even as We are one: I in them, and Thou in Me, that they may be made perfect in one}; and that the world may know that Thou hast sent Me, and hast loved them, as Thou hast loved Me.’ John 17:20-23.}[8T 269.3; 1904][https://egwwritings.org/?ref=en\_8T.269.3&para=112.1562]

\egwnogap{Wonderful statement! \textbf{The unity that exists between Christ and His disciples \underline{does not destroy the personality of either}. They are one in purpose, in mind, in character, but \underline{not in person}. It is thus that God and Christ are one}.}[8T 269.4; 1904][https://egwwritings.org/?ref=en\_8T.269.4&para=112.1563]

\egwnogap{\textbf{The relation between the Father and the Son, and the personality of both, are made plain in this scripture also}:}[8T 269.5; 1904][https://egwwritings.org/?ref=en\_8T.269.5&para=112.1564]

\egwnogap{Thus speaketh \textbf{Jehovah of hosts}, saying,} \\
\egw{Behold, \textbf{the man} whose name is\textbf{ the Branch}:} \\
\egw{And He shall grow up out of His place;} \\
\egw{\textbf{And He shall build the temple of Jehovah;... }} \\
\egw{\textbf{And He shall bear the glory,}} \\
\egw{\textbf{And shall sit and rule upon His throne;}} \\
\egw{\textbf{And He shall be a priest upon His throne;}} \\
\egw{\textbf{And \underline{the counsel of peace shall be between Them both}}.’}[8T 269.6; 1904][https://egwwritings.org/?ref=en\_8T.269.6&para=112.1565]

The aforementioned chapters of the Gospel of John deal with the \emcap{personality of God}, which had been expressed in the first two points of the \emcap{Fundamental Principles}. What error did Sister White combat when she referenced verses on how the Father was the only true God, and how the Father and the Son are not one in person? Pantheism? Certainly not; but most probably the trinitarian sentiments, or belief in a one-in-three, or three-in-one God.

Brother J. N. Loughborough, one of the first brethren who wrote on the \emcap{personality of God}, wrote the following comment on John chapter 17:

\others{\textbf{\underline{The seventeenth chapter of John is alone sufficient to refute the doctrine of the Trinity}}. \textbf{...\underline{Read the seventeenth chapter of John, and see if it does not completely upset the doctrine of the Trinity}}.}[John N. Loughborough, The Adventist Review, and Sabbath Herald, November 5, 1861, p. 184.10][https://egwwritings.org/?ref=en\_ARSH.November.5.1861.p.184.1&para=1685.6615]

Sister White’s proactive writing in support of the truth on the \emcap{personality of God} and His presence is the same as other Adventist pioneers. If Adventist pioneers were debunking the Trinity doctrine by exalting the truth on the \emcap{personality of God} and God’s presence, what makes us think Ellen White was not doing the same, when the theological side of the question of the Trinity was raised? By stating this, we do not deny the pantheistic side of Kellogg’s controversy, but by overemphasizing it, it falls short of accurately describing its real issue. The correct understanding of the Kellogg controversy can only be accomplished by focusing primarily on the truth Sister White uplifted, rather than focusing on error, whether pantheism or Trinity. This truth that Sister White uplifted was the truth on the \emcap{personality of God} and where His presence is. This is expressed in the first point of the \emcap{Fundamental Principles}, which were the official synopsis and representation of Seventh-day Adventist beliefs in the time of Ellen White; the truth which we, as a church, \egwinline{have received and heard and advocated}[Ms124-1905.12; 1905][https://egwwritings.org/?ref=en\_Ms124-1905.12&para=9099.18] in the beginning.

\egw{\textbf{I entreat every one to be clear and firm regarding the certain truths that we have received and heard and advocated. The statements of God’s Word are plain. Plant your feet firmly on \underline{the platform of eternal truth}. \underline{Reject every phase of error}, even \underline{though it be covered with a semblance of reality, which denies the personality of God or of Christ}}.}[Ms124-1905.12; 1905][https://egwwritings.org/?ref=en\_Ms124-1905.12&para=9099.18]

The warning from the previous quotations did not lessen in the course of time. Today it is even more relevant. We should \egwinline{reject every phase of error, even though it be covered with a semblance of reality, which denies the personality of God or of Christ}.