\chapter{The prophetic call for a renewal of the old pillars}

Today, God is in the business of renewing the \emcap{Fundamental Principles}. We have the promise that the old pillars of our faith will be preserved because God calls us to renew these pillars. Let us harken unto the will of God!

\egw{\textbf{\underline{Our people need to understand the reasons of our faith and our past experiences}. How sad it is that so many of them apparently place unlimited confidence in men who present theories tending to uproot our past experiences and to remove the old landmarks!} Those who can so easily be led by a false spirit show that they have been following the wrong captain for some time--so long that they do not discern that they are departing from the faith, or that they are not building upon the true foundation. We need to urge all to put on their spiritual eyeglasses, to have their eyes anointed that they may see clearly and \textbf{discern the true pillars of the faith}. Then they will know that ‘the foundation of God standeth sure, having this seal, The Lord knoweth them that are his’‘ (2 Tim. 2:19). \textbf{\underline{We need to revive the old evidences of the faith once delivered to the saints}}.}[SW April 5, 1904, Art. B, par. 1; 1904][https://egwwritings.org/?ref=en\_SW.April.5.1904.Art.B.par.1&para=489.857]

As we read, Satan’s biggest effort is to change our conception of the \emcap{personality of God}. In order to save us from Satan’s deception, God wants us to revive the old evidences of faith given to our pioneers. We need to understand the biblical evidence on why “\textit{one God}” is the Father, and that He is a personal, spiritual being. Studying this subject, we are rehearsing the history of our pioneers.

\egw{\textbf{The Lord has declared that \underline{the history of the past shall be rehearsed} as we enter upon the closing work. \underline{Every truth} that He has given for these last days is to be proclaimed to the world. \underline{Every pillar} that He has established \underline{is to be strengthened}. We cannot now step off the foundation that God has established... \underline{There is need now to rehearse the experience of the men} who acted a part in the establishment of our work \underline{at the beginning}}.}[Ms129-1905.7; 1905][https://egwwritings.org/?ref=en\_Ms129-1905.7&para=9797.14]

We need to study all the pillars of our faith, including the \emcap{personality of God}! God is in the business of reviving His truth, as well as His church. This will not happen without a shaking among God’s people. We have a special testimony about what will cause the shaking of the church:

\egw{I asked \textbf{the meaning of the shaking} I had seen, and was shown that \textbf{it would be caused by \underline{the straight testimony} called forth by the counsel of the True Witness to the Laodiceans}. \textbf{This will have its effect upon the heart of the receiver, and will lead him to \underline{exalt the standard and pour forth the straight truth}. Some will not bear this straight testimony. \underline{They will rise up against it, and this will cause a shaking among God’s people}}.}[T04 34.4; 1857][https://egwwritings.org/?ref=en\_T04.34.4&para=12677.185]

The following quotation gives us details about what message will contain the straight testimony.

\egw{\textbf{\underline{The Lord calls for a renewal of the straight testimony borne in years past}.} \textbf{He calls for a renewal of spiritual life. The spiritual energies of His people have long been torpid, but there is to be a resurrection from apparent death}.}[8T 297.5; 1904][https://egwwritings.org/?ref=en\_8T.297.5&para=112.1796]

\egwnogap{\textbf{\underline{By prayer and confession of sin we must clear the King’s highway}. As we do this, the power of the Spirit will come to us. We need the Pentecostal energy. \underline{This will come}, for the Lord has promised to send His Spirit as the all-conquering power}.}[8T 297.6; 1904][https://egwwritings.org/?ref=en\_8T.297.6&para=112.1797]

\egwnogap{\textbf{Perilous times are before us. Everyone who has a knowledge of the truth should awake and place himself, body, soul, and spirit, \underline{under the discipline of God}}. \textbf{The enemy is on our track. We must be wide awake, on our guard against him}. \textbf{We must put on the whole armor of God}. \textbf{\underline{We must follow the directions given through the spirit of prophecy}. \underline{We must love and obey the truth for this time}. \underline{This will save us from accepting strong delusions}. God has spoken to us through His word. He has spoken to us through the testimonies to the church and through the books that have helped to make plain our present duty and the position that we should now occupy. The warnings that have been given, line upon line, precept upon precept, should be heeded. If we disregard them, what excuse can we offer?}}[8T 298.1; 1904][https://egwwritings.org/?ref=en\_8T.298.1&para=112.1800]

\egwnogap{“\textbf{I beseech those who are laboring for God not to accept the spurious for the genuine. Let not human reason be placed where divine, sanctifying truth should be}. \textbf{Christ is waiting to kindle faith and love in the hearts of His people}. \textbf{Let not erroneous theories receive countenance from the people who ought to be standing firm on the platform of eternal truth. \underline{God calls upon us to hold firmly to the fundamental principles that are based upon unquestionable authority}}”.}[8T 298.2; 1904][https://egwwritings.org/?ref=en\_8T.298.2&para=112.1801]

The straight testimony that will cause the shaking is the testimony borne in years past. This testimony is the message contained in the \emcap{Fundamental Principles} mingled with the council of the True Witness to the church of Laodicea.

The final outcome of the shaking will be God’s revival of the first powerful experiences of our pioneers, which they had after the great disappointment. Sister White confirms this several times. One instance is found in her diary, dated November 27, 1902.

\egw{\textbf{I have been deeply impressed by the Spirit of God that we are to pass through severe trials. Everyone’s faith will be tested. \underline{We must study carefully the old waymarks}.} \textbf{\underline{These experiences in the past are to be revived}}. Daniel is to stand out conspicuously with the Revelation given to John on the Isle of Patmos.}[Ms223-1902.11; 1902][https://egwwritings.org/?ref=en\_Ms223-1902.11&para=9124.26]

\egwnogap{\textbf{In our experience in these last days we shall meet every conceivable thing that Satan can invent to make \underline{of none effect the established points of our faith} that have been, in the providence of God, so greatly blessed.} \textbf{\underline{These foundation principles} are to be held fast unto the end. Read the Word of God}.}[Ms223-1902.13; 1902][https://egwwritings.org/?ref=en\_Ms223-1902.13&para=9124.28]

Again, God calls us to hold fast to the \emcap{Fundamental Principles} unto the end.

\egw{“We are God’s commandment-keeping people. For the past fifty years every phase of heresy has been brought to bear upon us, to becloud our minds regarding the teaching of the word,—especially concerning the ministration of Christ in the heavenly sanctuary, and the message of heaven for these last days, as given by the angels of the fourteenth chapter of Revelation. \textbf{Messages of every order and kind have been urged upon Seventh-day Adventists, to take the place of the truth which, \underline{point by point}, has been sought out by prayerful study, and testified to by the miracle-working power of the Lord}. \textbf{But the way-marks which have made us what we are, \underline{are to be preserved, and they will be preserved}, as God has signified through His word and the testimony of His Spirit. \underline{He calls upon us to hold firmly}, with the grip of faith, \underline{to the fundamental principles that are based upon unquestionable authority}}.“}[SpTB02 59.1; 1904][https://egwwritings.org/?ref=en\_SpTB02.59.1&para=417.299]

Let us make careful study of the \emcap{Fundamental Principles}.

\egw{“\textbf{We are living in a time when every wind of doctrine is blowing and when those who think they stand are liable to fall}. We are living in a time when Satan is striving to implant seeds of skepticism and infidelity in every mind. \textbf{We are living in a time when error is taught so insidiously that the faith of many is being rapidly undermined}.“}[Ms143-1907.17; 1907][https://egwwritings.org/?ref=en\_Ms143-1907.17&para=7851.23]

\egw{Oh, how much we lose by neglecting the privilege of partaking freely of the bread of life! Shall we not resolutely refuse to be ensnared by the enemy of our souls? Shall we not place beyond our reach everything that turns the mind away from the truths that God desires us to learn? \textbf{Let us seek to become familiar with the books that clearly outline \underline{the truths for this time}}. \textbf{\underline{Let us make a careful study of the fundamental principles} of the message that is being proclaimed by God’s children throughout the world}. \textbf{Let us keep informed regarding the progress of this message}. A most solemn work is now in progress—the work of warning an impenitent world of the judgment day and of the soon coming of our Saviour in the clouds of heaven. God desires that every child of His shall have a part to act in this great work. Let us come up to the help of the Lord, to the help of the Lord against the mighty.}[Ms143-1907.18; 1907][https://egwwritings.org/?ref=en\_Ms143-1907.18&para=7851.24]

In this study we have had the opportunity to become acquainted with \egwinline{the books that clearly outline the truths for this time}. We have looked unto the writings of our pioneers regarding the \emcap{personality of God}. We have seen \egwinline{the old evidences of the faith once delivered to the saints}. In examining the evidence regarding the \emcap{personality of God}, we have also seen their evidence for not accepting the Trinity doctrine. Unfortunately, we have forgotten the pillar of our faith regarding the \emcap{personality of God}, and because we have forgotten, it’s necessary for us to remember the \egwinline{\textbf{way the Lord has led us, and \underline{His teaching} in our past history.}}[LS 196.2; 1915][https://egwwritings.org/?ref=en\_LS.196.2] We should make careful study of the \emcap{Fundamental Principles}. This is the purpose behind “\textit{The Forgotten Pillar Project}”. We encourage you to carefully study the first and second points of the \emcap{Fundamental Principles}, which deal with the \emcap{personality of God} and where His presence is. For this cause, we have made an in-depth study of the \emcap{personality of God} in contrast to the present understanding of the Trinity doctrine. We invite you to read and study “\textit{Rediscovering the Pillar}”\footnote{You can find the book on The Forgotten Pillar website: \href{http://forgottenpillar.com}{forgottenpillar.com}}, which is the sequel to this book.