\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter20}{Dr. Kellogg and Ellen White writings}


\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter20}{د. كيلوغ وكتابات إلين وايت}


Dr. Kellogg asserted that in the Living Temple he represented the same sentiments advocated by Sister White. Likewise, today many claim that Sister White was trinitarian and was responsible for the church's acceptance of the Trinity doctrine\footnote{William Johnsson, Adventist Review, January 6th, 1994, ‘\textit{Present Truth –Walking in God’s Light}’}. Sister White, herself, declared such claims to be false.


ادعى د. كيلوغ أنه في كتاب ذا ليفينغ تمبل قدم نفس الآراء التي دافعت عنها الأخت وايت. وبالمثل، اليوم يدعي الكثيرون أن الأخت وايت كانت تؤمن بالثالوث وكانت مسؤولة عن قبول الكنيسة لعقيدة الثالوث\footnote{William Johnsson, Adventist Review, January 6th, 1994, ‘\textit{Present Truth –Walking in God's Light}’}. أعلنت الأخت وايت نفسها أن مثل هذه الادعاءات كاذبة.


\egw{\textbf{The enemy is seeking to \underline{bring in} among the people of God spiritualistic theories, which \underline{if accepted, would undermine the foundation of the faith} that has made us what we are}. He leads men to present fables clothed with Scripture. \textbf{There are those who assert that Sister White’s writings are in harmony with these teachings}.\textbf{ \underline{I declare this to be false}. Men may misapply Scripture; they may misinterpret my words; but God understands their devising}. How thankful I am for this! When the enemy comes in like a flood, \textbf{the Spirit of the Lord will lift up a standard for us against him}.}[Ms137-1903.21; 1903][https://egwwritings.org/read?panels=p9939.30]


\egw{\textbf{العدو يسعى \underline{لإدخال} نظريات روحانية بين شعب الله، والتي \underline{إذا قُبلت، ستقوض أساس الإيمان} الذي جعلنا ما نحن عليه}. إنه يقود الرجال لتقديم خرافات مغلفة بالكتاب المقدس. \textbf{هناك من يؤكدون أن كتابات الأخت وايت تتوافق مع هذه التعاليم}.\textbf{ \underline{أعلن أن هذا كذب}. قد يسيء الرجال تطبيق الكتاب المقدس؛ قد يسيئون تفسير كلماتي؛ لكن الله يفهم تدبيرهم}. كم أنا ممتنة لهذا! عندما يأتي العدو كسيل، \textbf{فإن روح الرب سيرفع راية ضده من أجلنا}.}[Ms137-1903.21; 1903][https://egwwritings.org/read?panels=p9939.30]


Dr. Kellogg advocated the theories that, if accepted, would undermine the foundation of our faith. It is crucial to correctly understand what constitutes the foundation of our faith, which Sister White referred to. We have seen that it refers to the \emcap{Fundamental Principles}. Looking at her writings, and the writings of our pioneers, we see that the Trinity doctrine contradicts the \emcap{personality of God} and the truth about God’s presence. Today, with the Trinity doctrine as part of our belief, we recognize that we have moved away from the \emcap{Fundamental Principles} and formed another foundation. Sister White was not responsible for this transition. It is purely a misinterpretation of her works. Her writings do not undermine the foundation of the faith that has made us what we are. Her later work is completely in harmony with the truth given in the beginning.


دافع د. كيلوغ عن النظريات التي، إذا قُبلت، ستقوض أساس إيماننا. من الضروري فهم ما يشكل أساس إيماننا بشكل صحيح، والذي أشارت إليه الأخت وايت. لقد رأينا أنه يشير إلى \emcap{المبادئ الجوهرية}. بالنظر إلى كتاباتها، وكتابات روادنا، نرى أن عقيدة الثالوث تتعارض مع \emcap{شخصانية الله} والحقيقة حول حضور الله. اليوم، مع وجود عقيدة الثالوث كجزء من معتقدنا، ندرك أننا ابتعدنا عن \emcap{المبادئ الجوهرية} وشكلنا أساسًا آخر. لم تكن الأخت وايت مسؤولة عن هذا التحول. إنه مجرد سوء تفسير لأعمالها. كتاباتها لا تقوض أساس الإيمان الذي جعلنا ما نحن عليه. عملها اللاحق متوافق تمامًا مع الحقيقة المعطاة في البداية.


\egw{\textbf{The past fifty years have not dimmed one jot or principle of our faith as we received the great and wonderful evidences that were made certain to us in 1844, after the passing of the time.} ... \textbf{\underline{Not a word is changed or denied}. That which the Holy Spirit testified to as truth after the passing of the time, in our great disappointment, \underline{is the solid foundation of truth}. \underline{Pillars of truth were revealed}, and we accepted \underline{the foundation principles} that have made us what we are—Seventh-day Adventists, keeping the commandments of God and having the faith of Jesus.}}[Lt326-1905.3; 1905][https://egwwritings.org/read?panels=p7678.9]


\egw{\textbf{الخمسون سنة الماضية لم تُعتم ذرة أو مبدأ من إيماننا كما تلقينا الأدلة العظيمة والرائعة التي تأكدت لنا في عام 1844، بعد مرور الوقت.} ... \textbf{\underline{لم تتغير أو تُنكر كلمة واحدة}. ما شهد به الروح القدس كحقيقة بعد مرور الوقت، في خيبة أملنا العظيمة، \underline{هو أساس الحقيقة الصلب}. \underline{كُشفت أعمدة الحقيقة}، وقبلنا \underline{المبادئ الأساسية} التي جعلتنا ما نحن عليه - أدفنتست سبتيين، نحفظ وصايا الله ولنا إيمان يسوع.}}[Lt326-1905.3; 1905][https://egwwritings.org/read?panels=p7678.9]


\section*{Misrepresentation of the church standpoint}


\section*{سوء تمثيل موقف الكنيسة}


By misrepresenting Sister White’s writings, Dr. Kellogg did not only misrepresent her work, but also the church’s official standpoint expressed in the \emcap{Fundamental Principles}. Ellen White rebuked Kellogg for misrepresenting the church’s standpoint. As we read this rebuke, let us keep in mind the church’s current standpoint on the \emcap{personality of God} as it compares to the first point of the \emcap{Fundamental Principles}.


من خلال تحريف كتابات الأخت وايت، لم يسئ د. كيلوغ تمثيل عملها فحسب، بل أيضًا موقف الكنيسة الرسمي المعبر عنه في \emcap{المبادئ الجوهرية}. وبخت إلين وايت كيلوغ لسوء تمثيله لموقف الكنيسة. بينما نقرأ هذا التوبيخ، دعونا نضع في اعتبارنا موقف الكنيسة الحالي حول \emcap{شخصانية الله} مقارنة بالنقطة الأولى من \emcap{المبادئ الجوهرية}.


\egw{You \textbf{are not sound in the truth}. Your statements made to believers and unbelievers \textbf{misrepresent us as a people who have not changed the truth for error}. They detract from the influence \textbf{God would have us possess before the world in revealing in plain, unmistakable language that we are \underline{true to the principles of our faith} and that we hold the beginning of our confidence firm unto the end}. We are strictly denominational. \textbf{We believe in 1903 the same truths we did believe when we established the Sanitarium and the College in Battle Creek, and \underline{we know that we had no ifs or ands about this matter}}.}[Lt300-1903.4; 1903][https://egwwritings.org/read?panels=p7705.10]


\egw{أنت \textbf{لست راسخًا في الحق}. تصريحاتك التي قدمتها للمؤمنين وغير المؤمنين \textbf{تسيء تمثيلنا كشعب لم يغير الحق بالخطأ}. إنها تنتقص من التأثير \textbf{الذي يريد الله أن نمتلكه أمام العالم في إظهار بلغة واضحة لا لبس فيها أننا \underline{أمناء لمبادئ إيماننا} وأننا نتمسك ببداية ثقتنا ثابتة إلى النهاية}. نحن طائفيون بشكل صارم. \textbf{نؤمن في عام 1903 بنفس الحقائق التي آمنا بها عندما أسسنا المصح والكلية في باتل كريك، \underline{ونعلم أنه لم يكن لدينا أي شكوك أو تحفظات حول هذا الأمر}}.}[Lt300-1903.4; 1903][https://egwwritings.org/read?panels=p7705.10]


\egwnogap{While you have told the things that you have and made the statements you have before unbelievers, my heart has been sad indeed. \textbf{You have evidenced that you have departed from the faith}. The very statements you have made before worldly men of influence, as the papers have reported your words, have been presented to me distinctly from your lips as you have spoken them. We cannot labor to give you influence as one whom we can trust with the sacred work connected with our institutions, for you need first to be converted and led.}[Lt300-1903.5; 1903][https://egwwritings.org/read?panels=p7705.11]


\egwnogap{بينما كنت تخبر بالأشياء التي ذكرتها وتدلي بالتصريحات التي أدليت بها أمام غير المؤمنين، كان قلبي حزينًا حقًا. \textbf{لقد أظهرت أنك انحرفت عن الإيمان}. نفس التصريحات التي أدليت بها أمام رجال العالم ذوي النفوذ، كما نقلت الصحف كلماتك، قد قُدمت لي بوضوح من شفتيك كما تحدثت بها. لا يمكننا العمل لمنحك تأثيرًا كشخص يمكننا أن نثق به في العمل المقدس المرتبط بمؤسساتنا، لأنك تحتاج أولاً إلى أن تتحول وتُقاد.}[Lt300-1903.5; 1903][https://egwwritings.org/read?panels=p7705.11]


\egwnogap{You are not sound in the faith. I have stated this in my diary months ago. \textbf{You have certainly placed the people of God, whom the Lord has led step by step in the ways of truth and placed upon \underline{a solid foundation}, in a false showing before unbelievers. Some have departed from the faith and \underline{will continue to misrepresent the work God has given me}}.}[Lt300-1903.6; 1903][https://egwwritings.org/read?panels=p7705.12]


\egwnogap{أنت لست راسخًا في الإيمان. لقد ذكرت هذا في مذكراتي منذ شهور. \textbf{لقد وضعت بالتأكيد شعب الله، الذين قادهم الرب خطوة بخطوة في طرق الحق ووضعهم على \underline{أساس صلب}، في صورة زائفة أمام غير المؤمنين. البعض انحرف عن الإيمان و\underline{سيستمر في إساءة تمثيل العمل الذي أعطاني الله إياه}}.}[Lt300-1903.6; 1903][https://egwwritings.org/read?panels=p7705.12]


\egwnogap{\textbf{The sanctuary question is a clear and definite doctrine as we have held it as a people. \underline{You are not definitely clear on the personality of God, which is everything to us as a people}. \underline{You have virtually destroyed the Lord God Himself}}.}[Lt300-1903.7; 1903][https://egwwritings.org/read?panels=p7705.13]


\egwnogap{\textbf{إن مسألة المقدس هي عقيدة واضحة ومحددة كما تمسكنا بها كشعب. \underline{أنت لست واضحًا تمامًا بشأن شخصانية الله، التي هي كل شيء بالنسبة لنا كشعب}. \underline{لقد دمرت فعليًا الرب الإله نفسه}}.}[Lt300-1903.7; 1903][https://egwwritings.org/read?panels=p7705.13]


\egwnogap{Why should you take the liberty to make the statements which you have made, as though you had authority for thus stating, when they are falsehoods? \textbf{You have made the facts of our faith of none effect before unbelievers,} \textbf{and the truth which should ever be kept prominent and exalted with this people you have virtually denied and ignored in your many statements. How dared you to do this?} \textbf{It necessitates us now to present our true position which constitutes us Seventh-day Adventists}. Whatever influence God has given you in the past has been in mercy to you, letting the light shine upon you.}[Lt300-1903.8; 1903][https://egwwritings.org/read?panels=p7705.14]


\egwnogap{لماذا تأخذ لنفسك الحرية في الإدلاء بالتصريحات التي أدليت بها، كما لو كان لديك سلطة للتصريح بذلك، بينما هي أكاذيب؟ \textbf{لقد جعلت حقائق إيماننا بلا تأثير أمام غير المؤمنين،} \textbf{والحقيقة التي يجب أن تُبقى دائمًا بارزة ومرتفعة مع هذا الشعب قد أنكرتها وتجاهلتها فعليًا في تصريحاتك العديدة. كيف تجرأت على فعل ذلك؟} \textbf{هذا يستلزم منا الآن أن نقدم موقفنا الحقيقي الذي يجعلنا أدفنتست سبتيين}. أي تأثير منحك الله إياه في الماضي كان برحمة منه، تاركًا النور يشرق عليك.}[Lt300-1903.8; 1903][https://egwwritings.org/read?panels=p7705.14]


\egwnogap{\textbf{We cannot for a moment have any misrepresentation upon these solemn and important subjects of truth which have been the faith of our people since 1844. This means much to us.} The Lord would have me say to you that the enemy has, through his specious deceptions, placed his unbelief in your mind, and you have been working it out. \textbf{\underline{All who receive your presentations will enter upon strange paths if they connect with you}}. \textbf{You are \underline{bringing in} strange, common fire}, \textbf{but not the fire of God’s own kindling}; and now \textbf{I must speak plainly to our people that the Lord has led us step by step and shown us clear light upon the heavenly sanctuary in the most holy of holies where \underline{God revealed Himself} to His appointed ones.}}[Lt300-1903.9; 1903][https://egwwritings.org/read?panels=p7705.15]


\egwnogap{\textbf{لا يمكننا للحظة أن نسمح بأي تحريف لهذه المواضيع الخطيرة والمهمة من الحق التي كانت إيمان شعبنا منذ عام 1844. هذا يعني الكثير بالنسبة لنا.} الرب يريدني أن أقول لك أن العدو، من خلال خداعه الماكر، قد وضع عدم إيمانه في عقلك، وأنت تعمل على تحقيقه. \textbf{\underline{كل من يتلقى عروضك سيدخل في مسارات غريبة إذا ارتبط بك}}. \textbf{أنت \underline{تُدخل} نارًا غريبة عادية}، \textbf{وليست نار الله التي أشعلها بنفسه}؛ والآن \textbf{يجب أن أتحدث بوضوح إلى شعبنا أن الرب قادنا خطوة بخطوة وأظهر لنا نورًا واضحًا على المقدس السماوي في قدس الأقداس حيث \underline{كشف الله عن نفسه} لمختاريه.}}[Lt300-1903.9; 1903][https://egwwritings.org/read?panels=p7705.15]


Dr. Kellogg misrepresented the truth that constituted the foundation of our faith; most specifically, he misrepresented the truth on the \emcap{personality of God}, which was everything to us as people. If in 1903, it necessitated \egwinline{\textbf{to present our true position which constitutes us Seventh-day Adventists}}, how much more important is it for us today? Sister White did her part in upholding the foundation of our faith in the beginning, but it seems like we have forgotten.


لقد أساء الدكتور كيلوغ تمثيل الحق الذي شكّل أساس إيماننا؛ وبشكل أكثر تحديدًا، أساء تمثيل الحق حول \emcap{شخصانية الله}، التي كانت كل شيء بالنسبة لنا كشعب. إذا كان في عام 1903، استلزم الأمر \egwinline{\textbf{تقديم موقفنا الحقيقي الذي يجعلنا أدفنتست سبتيين}}، فكم هو أكثر أهمية بالنسبة لنا اليوم؟ لقد قامت الأخت وايت بدورها في الحفاظ على أساس إيماننا في البداية، ولكن يبدو أننا قد نسينا.


\input{lang/en/poems/chapter20.tex}