\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter30}{The final call}


\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter30}{النداء الأخير}


When we are acquainted with the \emcap{Fundamental Principles}, especially with the first point dealing with the \emcap{personality of God} and where His presence is, Sister White’s warnings suddenly shine in a new light. In light of the \emcap{Fundamental Principles}, and as a conclusion to these studies, we would like to present the following warnings:


عندما نكون على دراية بـ \emcap{المبادئ الجوهرية}، وخاصة النقطة الأولى التي تتناول \emcap{شخصانية الله} وأين يوجد حضوره، فجأة تشع تحذيرات الأخت وايت بنور جديد. في ضوء \emcap{المبادئ الجوهرية}، وكخاتمة لهذه الدراسات، نود أن نقدم التحذيرات التالية:


\egw{Satan presents his theories cautiously at first, and if he sees that his efforts are successful, he brings in theories that are still more misleading, \textbf{seeking to lead men and women away from \underline{the foundation principles that God designs shall be the safeguards of His people}}.}[Ms132-1903.40; 1903][https://egwwritings.org/read?panels=p9056.54]


\egw{يقدم الشيطان نظرياته بحذر في البداية، وإذا رأى أن جهوده ناجحة، فإنه يقدم نظريات أكثر تضليلاً، \textbf{ساعياً لإبعاد الرجال والنساء عن \underline{المبادئ الأساسية التي يريد الله أن تكون حماية لشعبه}}.}[Ms132-1903.40; 1903][https://egwwritings.org/read?panels=p9056.54]


\egwnogap{\textbf{Let not our medical missionary workers accept theories that God has not given to any one}. \textbf{God will not excuse men for teaching theories that Christ has not taught}. He calls upon His army of workers to fall into line, taking their stand under the banner of truth. \textbf{He warns them to beware of occupying their time in the discussion of matters that God has not authorized any human being to discuss}.}[Ms132-1903.41; 1903][https://egwwritings.org/read?panels=p9056.55]


\egwnogap{\textbf{لا تدعوا العاملين في الإرسالية الطبية يقبلون نظريات لم يعطها الله لأي شخص}. \textbf{الله لن يعذر الرجال الذين يعلّمون نظريات لم يعلّمها المسيح}. إنه يدعو جيشه من العاملين للانضمام إلى الصف، واتخاذ موقفهم تحت راية الحق. \textbf{إنه يحذرهم من قضاء وقتهم في مناقشة أمور لم يأذن الله لأي إنسان بمناقشتها}.}[Ms132-1903.41; 1903][https://egwwritings.org/read?panels=p9056.55]


\egwnogap{Let us put on every piece of the Christian armor and steadfastly resist the enemy. We shall have to meet fallen angels and the prince of the powers of darkness. Satan is by no means asleep; he is wide-awake and is playing the game of life for the souls of the people of God. He will come to them with flattery of all kinds, in the hope of leading them to swerve from their allegiance. \textbf{He desires to call their attention from the real issues to false theories}.}[Ms132-1903.42, 1903][https://egwwritings.org/read?panels=p9056.56]


\egwnogap{لنلبس كل قطعة من درع المسيحي ولنقاوم العدو بثبات. سنضطر لمواجهة الملائكة الساقطة وأمير قوات الظلام. الشيطان ليس نائماً بأي حال؛ إنه مستيقظ تماماً ويلعب لعبة الحياة من أجل نفوس شعب الله. سيأتي إليهم بإطراء من كل نوع، على أمل أن يقودهم للانحراف عن ولائهم. \textbf{إنه يرغب في صرف انتباههم من القضايا الحقيقية إلى نظريات كاذبة}.}[Ms132-1903.42, 1903][https://egwwritings.org/read?panels=p9056.56]


\egwnogap{Ministers and physicians, sound an alarm. Call upon the people of God to be true and faithful. Be on your guard. Remember that as you co-operate with God, you have as your helpers angels that excel in strength. \textbf{Accept not the theories advanced by those who are not standing on the \underline{true foundation}, those who are charmed with that [of] which they do not know the true meaning}.}[Ms132-1903.43; 1903][https://egwwritings.org/read?panels=p9056.57]


\egwnogap{أيها الخدام والأطباء، أطلقوا إنذاراً. ادعوا شعب الله ليكونوا أمناء ومخلصين. كونوا على حذر. تذكروا أنه عندما تتعاونون مع الله، يكون معكم كمساعدين ملائكة يتفوقون في القوة. \textbf{لا تقبلوا النظريات التي يطرحها أولئك الذين لا يقفون على \underline{الأساس الحقيقي}، أولئك المفتونون بما لا يعرفون معناه الحقيقي}.}[Ms132-1903.43; 1903][https://egwwritings.org/read?panels=p9056.57]


\egwnogap{\textbf{\underline{Wake up}, my brethren, wake up and lift the danger signal}. \textbf{\underline{Sound the warning}}. \textbf{Let no man persuade you to accept theories that are opposed to the truths of God’s Word}. The servants of God have a solemn message to bear to this fallen, sin-cursed world. They are to hold aloft the banner on which is inscribed the words, ‘\textbf{The commandments of God and the faith of Jesus}.’ [Revelation 14:12.] \textbf{Those who are working in harmony with God will be of one heart and one mind}. With unflagging zeal they will proclaim the message, ‘Prepare to meet thy God.’ [Amos 4:12.] \textbf{They will not unite with worldlings, but will take their stand firmly in defense of \underline{the principles of truth}}.}[Ms132-1903.44; 1903][https://egwwritings.org/read?panels=p9056.58]


\egwnogap{\textbf{\underline{استيقظوا}، يا إخوتي، استيقظوا وارفعوا إشارة الخطر}. \textbf{\underline{أطلقوا التحذير}}. \textbf{لا تدعوا أي إنسان يقنعكم بقبول نظريات تعارض حقائق كلمة الله}. لدى خدام الله رسالة جادة ليحملوها إلى هذا العالم الساقط الملعون بالخطيئة. عليهم أن يرفعوا عالياً الراية المكتوب عليها: ‘\textbf{وصايا الله وإيمان يسوع}.’ [رؤيا 14:12.] \textbf{أولئك الذين يعملون في انسجام مع الله سيكونون بقلب واحد وفكر واحد}. بحماس لا يفتر سيعلنون الرسالة، ‘استعد للقاء إلهك.’ [عاموس 4:12.] \textbf{لن يتحدوا مع أهل العالم، بل سيقفون بثبات دفاعاً عن \underline{مبادئ الحق}}.}[Ms132-1903.44; 1903][https://egwwritings.org/read?panels=p9056.58]


\egwnogap{Things must now be called by their right name. \textbf{Backsliding leaders are not to be honored as men who are holding their confidence firm}. God is in earnest with us. We are to sound the note of warning.}[Ms132-1903.45; 1903][https://egwwritings.org/read?panels=p9056.59]


\egwnogap{يجب الآن تسمية الأشياء بأسمائها الصحيحة. \textbf{القادة المرتدون لا ينبغي أن يُكرَّموا كرجال يحافظون على ثقتهم بثبات}. الله جاد معنا. علينا أن نطلق نغمة التحذير.}[Ms132-1903.45; 1903][https://egwwritings.org/read?panels=p9056.59]


\egwnogap{\textbf{Wake up, for Christ’s sake, wake up}. May God give success to those who are trying to arouse the sleepy watchmen. \textbf{Of many of those who profess to be shepherds of the flock, God says, ‘They are unfaithful}. \textbf{They have left their \underline{first love}}. \textbf{Unless they repent, I will come suddenly, and will remove their candlestick out of his place}.’ [See Revelation 2:4, 5.]}[Ms132-1903.46; 1903][https://egwwritings.org/read?panels=p9056.61]


\egwnogap{\textbf{استيقظوا، من أجل المسيح، استيقظوا}. ليمنح الله النجاح لأولئك الذين يحاولون إيقاظ الحراس النائمين. \textbf{عن كثيرين ممن يدّعون أنهم رعاة القطيع، يقول الله: ‘إنهم غير أمناء}. \textbf{لقد تركوا \underline{محبتهم الأولى}}. \textbf{ما لم يتوبوا، سآتي فجأة، وأزيح منارتهم من مكانها}.’ [انظر رؤيا 2:4، 5.]}[Ms132-1903.46; 1903][https://egwwritings.org/read?panels=p9056.61]


\egwnogap{Go to work now, without delay. How many judgment calls must the Lord make before His people cease to provoke Him to His face? Should He deal with them according to their backsliding, according to their worldliness and \textbf{to the way in which they have called darkness light and light darkness}, they would have no further calls to repentance, no more evidence or light to trifle with.}[Ms132-1903.47; 1903][https://egwwritings.org/read?panels=p9056.61]


\egwnogap{اذهبوا للعمل الآن، دون تأخير. كم من نداءات الدينونة يجب أن يطلقها الرب قبل أن يكف شعبه عن استفزازه في وجهه؟ لو تعامل معهم وفقاً لارتدادهم، ووفقاً لعالميتهم \textbf{وللطريقة التي بها دعوا الظلام نوراً والنور ظلاماً}، لما كانت هناك دعوات أخرى للتوبة، ولا مزيد من الأدلة أو النور ليستخفوا به.}[Ms132-1903.47; 1903][https://egwwritings.org/read?panels=p9056.61]


\egwnogap{God’s people provoke Him by their idolatry and by their union with worldlings. He says, ‘My Spirit shall not always strive with men. I will not always bear with the perversity of those who lead souls from the narrow way into paths of uncertainty and falsehood. Unless those who have been often reproved make a decided change, they will be left to follow their own way.’ \textbf{His blessings will be taken from those who choose darkness rather than light, those who choose false guides rather than true. To those who disregard the evidence given them, making no difference between truth and error, the light bestowed will become darkness, and how great will be that darkness}.}[Ms132-1903.48; 1903][https://egwwritings.org/read?panels=p9056.62]


\egwnogap{يستفز شعب الله غضبه بعبادتهم للأوثان وباتحادهم مع العالميين. يقول: ‘روحي لن يجاهد الإنسان إلى الأبد. لن أتحمل دائمًا انحراف أولئك الذين يقودون النفوس من الطريق الضيق إلى مسارات الشك والكذب. ما لم يُحدث أولئك الذين وُبخوا مرارًا تغييرًا حاسمًا، سيُتركون ليتبعوا طريقهم الخاص.’ \textbf{ستُؤخذ بركاته من أولئك الذين يختارون الظلمة بدلاً من النور، أولئك الذين يختارون المرشدين الكاذبين بدلاً من الحقيقيين. بالنسبة لأولئك الذين يتجاهلون الأدلة المقدمة لهم، ولا يفرقون بين الحق والخطأ، سيصبح النور الممنوح لهم ظلامًا، وما أعظم ذلك الظلام}.}[Ms132-1903.48; 1903][https://egwwritings.org/read?panels=p9056.62]


May the Lord help us to wake up to His callings. Let us choose light rather than darkness. Let us plant our feet upon the platform of eternal Truth. Let us believe and obey the truths we first embraced regarding the personality of the Father and of the Son, and the blessings will follow.


ليساعدنا الرب على الاستيقاظ لنداءاته. دعونا نختار النور بدلاً من الظلمة. دعونا نثبت أقدامنا على منصة الحق الأبدي. دعونا نؤمن ونطيع الحقائق التي اعتنقناها أولاً بخصوص شخصانية الآب والابن، وستتبعنا البركات.


\egw{A liar is one that \textbf{presents false theories and doctrines}. \textbf{He who \underline{denies the personality of God and of His Son Jesus Christ} is \underline{denying God and Christ}}. ‘If that which ye have heard \textbf{\underline{from the beginning}} shall \textbf{remain in you}, ye \textbf{also shall continue in the Son and in the Father}.’ \textbf{If you continue to believe and obey \underline{the truths you first embraced regarding the personality of the Father and the Son}, you will be joined together with them in love}. \textbf{There will be seen that union for which Christ prayed just before His trial and crucifixion}.}[Ms23-1906.20; 1906][https://egwwritings.org/read?panels=p7556.27]


\egw{الكاذب هو الذي \textbf{يقدم نظريات وتعاليم كاذبة}. \textbf{إن من \underline{ينكر شخصانية الله وابنه يسوع المسيح} هو \underline{ينكر الله والمسيح}}. ‘إن ثبت فيكم ما سمعتموه \textbf{\underline{من البدء}}، \textbf{تثبتون} أنتم أيضًا \textbf{في الابن وفي الآب}.’ \textbf{إذا استمررتم في الإيمان وطاعة \underline{الحقائق التي اعتنقتموها أولاً بخصوص شخصانية الآب والابن}، ستتحدون معهما في المحبة}. \textbf{سيُرى ذلك الاتحاد الذي صلى من أجله المسيح قبل محاكمته وصلبه مباشرة}.}[Ms23-1906.20; 1906][https://egwwritings.org/read?panels=p7556.27]


If we continue in the truth we first embraced regarding the personality of the Father and Son, we will be joined together with them in love. What a promise! We need this vital connection in our Christian life. Without exercising this union Christ prayed for, we are not able to fulfill the mission God has bestowed upon us as a prophetic movement.


إذا استمرينا في الحق الذي اعتنقناه أولاً بخصوص شخصانية الآب والابن، سنتحد معهما في المحبة. يا له من وعد! نحن بحاجة إلى هذا الاتصال الحيوي في حياتنا المسيحية. بدون ممارسة هذا الاتحاد الذي صلى المسيح من أجله، لن نتمكن من إتمام المهمة التي أوكلها الله إلينا كحركة نبوية.


\egw{\textbf{What possibilities are opened up to the youth who lay hold of the divine assurances of God’s Word}! \textbf{Scarcely can the human mind comprehend what is the breadth and depth and height of the spiritual attainments that can be reached by becoming partakers of the divine nature}. The human agent who <daily> yields obedience to God, who becomes a partaker of the divine nature, finds pleasure <daily> in keeping the commandments of God, \textbf{for he is one with God}. \textbf{<It is essential that> he \underline{holds as vital a relation with God as does the Son to the Father}}. \textbf{He understands the oneness that Christ prayed might exist between the Father and the Son}.}[Lt43-1895.18; 1895][https://egwwritings.org/read?panels=p5554.24]


\egw{\textbf{ما هي الإمكانيات التي تُفتح للشباب الذين يتمسكون بتأكيدات الله الإلهية في كلمته}! \textbf{بالكاد يستطيع العقل البشري أن يدرك ما هو عرض وعمق وارتفاع الإنجازات الروحية التي يمكن الوصول إليها من خلال المشاركة في الطبيعة الإلهية}. إن الإنسان الذي يطيع الله <يوميًا>، ويصبح شريكًا في الطبيعة الإلهية، يجد سرورًا <يوميًا> في حفظ وصايا الله، \textbf{لأنه واحد مع الله}. \textbf{<من الضروري أن> \underline{يحتفظ بعلاقة حيوية مع الله كما يفعل الابن مع الآب}}. \textbf{إنه يفهم الوحدة التي صلى المسيح من أجل أن توجد بين الآب والابن}.}[Lt43-1895.18; 1895][https://egwwritings.org/read?panels=p5554.24]


God has laid before us our church’s current state of affairs. The renewal of the old foundation will be accomplished. The \emcap{Fundamental Principles} will be restored and the three angels’ messages will be proclaimed in their true light and power. The only decision we are to make is a decision of faithfulness to Him, regardless of the cost. Do you and I want to be part of it? The blessings will follow as we hold fast unto the hand of the all-knowing God and His leadings. \bible{But those things, which God before had shewed by the mouth of all his prophets,} He will fulfill! \bible{Repent ye therefore, and be converted, that your sins may be blotted out, \textbf{when the times of refreshing shall come from the presence of the Lord; And he shall send Jesus Christ}, which before was preached unto you.}[Acts 3:18-20]. God will soon pour out the Spirit of Christ, which was preached unto us. He will send the Spirit of His Son into our hearts \bible{crying, Abba, Father}[Galatians 4:6].  Then, enabled with the power on High, His people will proclaim, in a loud cry, the three angels’ messages. \bible{And this gospel of the kingdom shall be preached in all the world for a witness unto all nations; and then shall the end come}[Matthew 24:14]. Christ will descend from High on the clouds of Heaven, escorted by thousands upon thousands of His Hosts, and He will take us into His Father’s presence, that we may be there where He is, and we may behold His glory, the glory which He had with God before the world was (\textit{John 17: 24, 5}). He will take us in the presence of His Father, and the glory will be revealed unto us, \egwinline{the unapproachable splendor of their converse together}[Lt232-1903.54; 1903][https://egwwritings.org/read?panels=p10197.63]. Then God shall wipe away all tears from our eyes; \bible{and there shall be no more death, neither sorrow, nor crying, neither shall there be any more pain: for the former things are passed away.}[Revelation 21:4]. \bible{And there shall be no more curse: but \textbf{the throne of God and of the Lamb} shall be in it; and his servants shall serve him.}[Revelation 22:3] And then from the core of our being, in gratitude toward God for His unspeakable Gift, all saved \bible{and every creature which is in heaven, and on the earth, and under the earth, and such as are in the sea, and all that are in them,} shall proclaim \bible{Blessing, and honour, and glory, and power, be unto \textbf{Him that sitteth upon the throne, and unto the Lamb} for ever and ever.}[Revelation 5:13]


لقد وضع الله أمامنا الوضع الحالي لكنيستنا. سيتم تجديد الأساس القديم. ستُستعاد \emcap{المبادئ الأساسية} وستُعلن رسائل الملائكة الثلاثة في نورها وقوتها الحقيقية. القرار الوحيد الذي علينا اتخاذه هو قرار الأمانة له، بغض النظر عن التكلفة. هل تريد أنت وأنا أن نكون جزءًا من ذلك؟ ستتبعنا البركات بينما نتمسك بيد الله العليم بكل شيء وقيادته. \bible{ولكن ما سبق الله وأنبأ به بأفواه جميع أنبيائه}، سيتممه! \bible{فتوبوا وارجعوا لتُمحى خطاياكم، \textbf{لكي تأتي أوقات الفرج من وجه الرب، ويرسل يسوع المسيح} المبشر به لكم من قبل.}[Acts 3:18-20]. سيسكب الله قريبًا روح المسيح، الذي بُشر به لنا. سيرسل روح ابنه إلى قلوبنا \bible{صارخًا: يا أبا الآب}[Galatians 4:6]. ثم، بتمكين من القوة العليا، سيعلن شعبه، بصرخة عالية، رسائل الملائكة الثلاثة. \bible{ويُكرز ببشارة الملكوت هذه في كل المسكونة شهادة لجميع الأمم. ثم يأتي المنتهى}[Matthew 24:14]. سينزل المسيح من العلى على سحاب السماء، يرافقه ألوف ألوف من جنوده، وسيأخذنا إلى حضرة أبيه، لنكون هناك حيث هو، ونرى مجده، المجد الذي كان له مع الله قبل كون العالم (\textit{يوحنا 17: 24، 5}). سيأخذنا إلى حضرة أبيه، وسيُكشف لنا المجد، \egwinline{البهاء الذي لا يُقترب منه لحديثهما معًا}[Lt232-1903.54; 1903][https://egwwritings.org/read?panels=p10197.63]. حينئذ يمسح الله كل دمعة من عيوننا؛ \bible{ولا يكون موت في ما بعد، ولا يكون حزن ولا صراخ ولا وجع في ما بعد، لأن الأمور الأولى قد مضت.}[Revelation 21:4]. \bible{ولا تكون لعنة ما في ما بعد. و\textbf{عرش الله والخروف} يكون فيها، وعبيده يخدمونه.}[Revelation 22:3] وحينئذ من صميم كياننا، امتنانًا لله على هبته التي لا يُعبر عنها، سيعلن جميع المخلصين \bible{وكل خليقة مما في السماء وعلى الأرض وتحت الأرض وما على البحر، كل ما فيها}، \bible{البركة والكرامة والمجد والسلطان \textbf{للجالس على العرش وللخروف} إلى أبد الآبدين.}[Revelation 5:13]


Amen!


آمين!


\input{lang/en/poems/chapter30}