\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter9}{The proper solution}


\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter9}{الحل المناسب}


God has given Sister White presentations regarding the deceptions that Satan is bringing into our church. She was instructed that we should read to and understand the works of our pioneers in order to combat these deceptions.


لقد أعطى الله الأخت وايت عروضًا تقديمية بخصوص الخداع الذي يُدخله الشيطان إلى كنيستنا. لقد تلقت تعليمات بأننا يجب أن نقرأ ونفهم أعمال روادنا من أجل مكافحة هذه الخداعات.


\egw{\textbf{I have had presentations regarding the deceptions that Satan is \underline{bringing in at this time}}. \textbf{I have been instructed} that \textbf{we should make prominent the testimony of some of the \underline{old workers} who are now dead}. \textbf{Let them continue to speak through their articles as found \underline{in the early numbers of our papers}}. \textbf{These articles should now be reprinted, that there may be \underline{a living voice from the Lord’s witnesses}}. \textbf{\underline{The history of the early experiences in the message will be a power to withstand the masterly ingenuity of Satan’s deceptions}.} This instruction has been repeated recently. I must present before the people \textbf{the testimonies of Bible truth and repeat the decided messages given years ago}...}[Lt99-1905.7; 1905][https://egwwritings.org/read?panels=p9997.13]


\egw{\textbf{لقد تلقيت عروضًا تقديمية بخصوص الخداع الذي يُدخله الشيطان \underline{في هذا الوقت}}. \textbf{لقد تلقيت تعليمات} بأننا \textbf{يجب أن نُبرز شهادة بعض \underline{العمال القدامى} الذين ماتوا الآن}. \textbf{دعوهم يستمرون في التحدث من خلال مقالاتهم كما وُجدت \underline{في الأعداد المبكرة من صحفنا}}. \textbf{يجب إعادة طباعة هذه المقالات الآن، حتى يكون هناك \underline{صوت حي من شهود الرب}}. \textbf{\underline{إن تاريخ التجارب المبكرة في الرسالة سيكون قوة لمقاومة براعة الشيطان الماهرة في الخداع}.} يجب أن أقدم للناس \textbf{شهادات حق الكتاب المقدس وأكرر الرسائل الحاسمة التي أُعطيت منذ سنوات}...}[Lt99-1905.7; 1905][https://egwwritings.org/read?panels=p9997.13]


\egw{\textbf{When men come in who would move one pin or pillar from the foundation} \textbf{which God has established by His Holy Spirit}, \textbf{let the} \textbf{\underline{aged men who were pioneers} in our work \underline{speak plainly}, and let those who are dead \underline{speak also by reprinting of their articles in our periodicals}}. Gather up the rays of divine light that God has given \textbf{as He has led His people on step by step in the way of truth}. \textbf{This truth will stand the test of time and trial}.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


\egw{\textbf{عندما يأتي رجال يريدون تحريك دبوس أو عمود واحد من الأساس} \textbf{الذي أسسه الله بروحه القدوس}، \textbf{دعوا} \textbf{\underline{الرجال المسنين الذين كانوا روادًا} في عملنا \underline{يتحدثون بوضوح}، ودعوا أولئك الذين ماتوا \underline{يتحدثون أيضًا من خلال إعادة طباعة مقالاتهم في دورياتنا}}. اجمعوا أشعة النور الإلهي التي أعطاها الله \textbf{عندما قاد شعبه خطوة بخطوة في طريق الحق}. \textbf{هذا الحق سيصمد أمام اختبار الزمن والتجربة}.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


We have seen that the Seventh-day Adventist Church has experienced the change in our foundation regarding the \emcap{personality of God}. We have been told that the first truth received by our pioneers will stand the test of time and trial. Has the truth on the \emcap{personality of God}, which we’ve received first, failed the test of time?


لقد رأينا أن كنيسة الأدفنتست السبتيين قد شهدت التغيير في أساسنا فيما يتعلق بـ \emcap{شخصانية الله}. لقد قيل لنا أن الحقيقة الأولى التي تلقاها روادنا ستصمد أمام اختبار الزمن والتجربة. هل فشلت الحقيقة حول \emcap{شخصانية الله}، التي تلقيناها أولاً، في اختبار الزمن؟


\egw{\textbf{When the power of God testifies to what is truth, the truth is to stand forever as the truth}. \textbf{No after suppositions, contrary to the light God has given are to be entertained}. Men will arise with interpretations of Scripture which are to them truth, but which are not truth. \textbf{The truth for this time, God has given us as \underline{a foundation for our faith}.} \textbf{He Himself has taught us what is truth.} One will arise and still another with new light which \textbf{contradicts }the light that God has given under the demonstration of His Holy Spirit. \textbf{A few are still alive who passed through the experience gained in the establishment of this truth}. God has graciously spared their lives\textbf{ to repeat and repeat} till the close of their lives, \textbf{the experience through which they passed} even as did John the apostle till the very close of his life. \textbf{And the standard bearers who have fallen in death, are to \underline{speak through the reprinting of their writings}. \underline{I am instructed} that thus [their] voices are to be heard}. \textbf{\underline{They are to bear their testimony as to what constitutes the truth for this time.}} \textbf{We are not to receive the words of those who come with a message \underline{that contradicts the special points of our faith}}. They gather together a mass of Scripture, and pile it as proof around their asserted theories. This has been done over and over again during the past fifty years. And while the Scriptures are God’s Word, and are to be respected, the application of them,\textbf{ if such application moves one pillar of the foundation that God has sustained these fifty years, is a great mistake}. He who makes such an application knows not the wonderful demonstration of the Holy Spirit that gave power and force to the past messages that have come to the people of God.}[PH020 14.2; 1905][https://egwwritings.org/read?panels=p276.89]


\egw{\textbf{عندما تشهد قوة الله على ما هو الحق، فإن الحق سيبقى إلى الأبد كحقيقة}. \textbf{لا ينبغي قبول أي افتراضات لاحقة مخالفة للنور الذي أعطاه الله}. سيظهر رجال بتفسيرات للكتاب المقدس هي بالنسبة لهم حق، لكنها ليست حقًا. \textbf{الحق لهذا الوقت، أعطانا الله إياه \underline{كأساس لإيماننا}.} \textbf{هو نفسه علمنا ما هو الحق.} سيظهر واحد وآخر بنور جديد \textbf{يتناقض مع }النور الذي أعطاه الله تحت إظهار روحه القدوس. \textbf{لا يزال هناك قليل على قيد الحياة ممن مروا بالتجربة المكتسبة في تأسيس هذا الحق}. لقد حفظ الله حياتهم برحمته \textbf{ليكرروا ويكرروا} حتى نهاية حياتهم، \textbf{التجربة التي مروا بها} تمامًا كما فعل يوحنا الرسول حتى نهاية حياته. \textbf{وحاملو الراية الذين سقطوا في الموت، يجب أن \underline{يتحدثوا من خلال إعادة طباعة كتاباتهم}. \underline{لقد تلقيت تعليمات} بأن أصواتهم يجب أن تُسمع بهذه الطريقة}. \textbf{\underline{عليهم أن يقدموا شهادتهم حول ما يشكل الحق لهذا الوقت.}} \textbf{يجب ألا نقبل كلمات أولئك الذين يأتون برسالة \underline{تتناقض مع النقاط الخاصة لإيماننا}}. إنهم يجمعون كمية من الكتاب المقدس، ويكدسونها كدليل حول نظرياتهم المؤكدة. لقد حدث هذا مرارًا وتكرارًا خلال الخمسين سنة الماضية. وبينما الكتب المقدسة هي كلمة الله، ويجب احترامها، فإن تطبيقها، \textbf{إذا كان هذا التطبيق يحرك عمودًا واحدًا من الأساس الذي دعمه الله هذه الخمسين سنة، هو خطأ كبير}. إن من يقوم بمثل هذا التطبيق لا يعرف الإظهار العجيب للروح القدس الذي أعطى قوة وفعالية للرسائل السابقة التي جاءت إلى شعب الله.}[PH020 14.2; 1905][https://egwwritings.org/read?panels=p276.89]


Let us become acquainted with the writing of our pioneers regarding the first point of the \emcap{Fundamental Principles}. Let’s become acquainted with the truth the pioneers wrote about the \emcap{personality of God} and where His presence is. Let us, \egwinline{gather up the rays of divine light that God has given as He has led His people on step by step in the way of truth.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


دعونا نتعرف على كتابات روادنا فيما يتعلق بالنقطة الأولى من \emcap{المبادئ الأساسية}. دعونا نتعرف على الحقيقة التي كتبها الرواد عن \emcap{شخصانية الله} وأين يوجد حضوره. دعونا، \egwinline{نجمع أشعة النور الإلهي التي أعطاها الله عندما قاد شعبه خطوة بخطوة في طريق الحق.}[Ms62-1905.20, 1905][https://egwwritings.org/read?panels=p10026.26]


\input{lang/en/poems/chapter9}