\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter29}{Remembering His teaching in our past history}


\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter29}{تذكر تعاليمه في تاريخنا الماضي}


We are bringing our journey in this study to its end. We started this journey with a strong proclamation of what constitutes the foundation of our faith. We became acquainted with the history and experience of our pioneers in establishing the Seventh-day Adventist Church. We have seen the mission and purpose God bestowed upon them in proclaiming the three angels’ messages to the whole world. These messages are interwoven with all important doctrines of the Bible. These doctrines are what our pioneers called the \emcap{fundamental principles}, or the pillars of our faith. These doctrines represent the foundation of our faith.


نحن نقترب من نهاية رحلتنا في هذه الدراسة. بدأنا هذه الرحلة بإعلان قوي لما يشكل أساس إيماننا. تعرفنا على تاريخ وتجربة روادنا في تأسيس كنيسة الأدفنتست السبتيين. رأينا المهمة والغرض الذي منحه الله لهم في إعلان رسائل الملائكة الثلاثة للعالم أجمع. هذه الرسائل متشابكة مع جميع العقائد المهمة في الكتاب المقدس. هذه العقائد هي ما أطلق عليها روادنا \emcap{المبادئ الجوهرية}، أو أعمدة إيماننا. هذه العقائد تمثل أساس إيماننا.


Once acquainted with the \emcap{Fundamental Principles} that we had in the beginning, we recognize their difference to our current Fundamental Beliefs, particularly to the question “\textit{who is God}”? Additionally, the present doctrine of God lacks understanding of His personality. In other words, it lacks the understanding of \textit{the quality or state} of God \textit{being a person}. This change in our doctrine becomes important in light of the first angel’s message, which pertains to the God we worship. Is the God we worship a triune God, or is He one God, the Father, the Ancient of Days, a personal and spiritual Being?


بعد التعرف على \emcap{المبادئ الجوهرية} التي كانت لدينا في البداية، ندرك اختلافها عن معتقداتنا الأساسية الحالية، خاصة فيما يتعلق بسؤال “\textit{من هو الله}”؟ بالإضافة إلى ذلك، تفتقر عقيدة الله الحالية إلى فهم شخصانيته. بعبارة أخرى، تفتقر إلى فهم \textit{الصفة أو الحالة} التي يكون بها الله \textit{شخصًا}. هذا التغيير في عقيدتنا يصبح مهمًا في ضوء رسالة الملاك الأول، التي تتعلق بالإله الذي نعبده. هل الإله الذي نعبده هو إله ثالوثي، أم هو إله واحد، الآب، القديم الأيام، كائن شخصي وروحي؟


The change in our understanding of who God is for Seventh-day Adventists was not instantaneous; it took years of controversy to arrive at our present standing. In these studies, we did not delve into history beyond the life of Ellen White. We saw change begin to take place in her time, when Dr. Kellogg insinuated the sentiment over the \emcap{personality of God}, which would \egwinline{lead astray the minds of those who are not thoroughly established on the foundation principles of present truth}[SpTB02 51.3; 1904][https://egwwritings.org/read?panels=p417.262]. He insinuated doubt in the clear revelations of the \emcap{personality of God} and of His Son, Jesus Christ. His sentiments were met by fierce rebukes from Sister White and strong warnings for the church, to avoid the path of doubt in the simple truth expressed in the \emcap{Fundamental Principles}—that one God is a personal, spiritual being, and Christ is His Son, \egwinline{begotten in the express image of the Father’s person}[ST May 30, 1895, par. 3; 1895][https://egwwritings.org/read?panels=p820.12891]. Thusly, Sister White firmly defended the first two points of the \emcap{Fundamental Principles}.


لم يكن التغيير في فهمنا لمن هو الله بالنسبة للأدفنتست السبتيين فوريًا؛ استغرق الأمر سنوات من الصراع للوصول إلى موقفنا الحالي. في هذه الدراسات، لم نتعمق في التاريخ بما يتجاوز حياة إلين وايت. رأينا بداية التغيير يحدث في زمنها، عندما أشار الدكتور كيلوغ إلى الرأي حول \emcap{شخصانية الله}، والذي من شأنه أن \egwinline{يضل عقول أولئك الذين ليسوا راسخين تمامًا على مبادئ الحق الحاضر الأساسية}[SpTB02 51.3; 1904][https://egwwritings.org/read?panels=p417.262]. لقد أثار الشك في الإعلانات الواضحة عن \emcap{شخصانية الله} وابنه يسوع المسيح. قوبلت آراؤه بتوبيخات شديدة من الأخت وايت وتحذيرات قوية للكنيسة، لتجنب طريق الشك في الحق البسيط المعبر عنه في \emcap{المبادئ الجوهرية}—أن الله الواحد هو كائن شخصي وروحي، والمسيح هو ابنه، \egwinline{المولود على صورة جوهره}[ST May 30, 1895, par. 3; 1895][https://egwwritings.org/read?panels=p820.12891]. وهكذا، دافعت الأخت وايت بقوة عن النقطتين الأوليين من \emcap{المبادئ الجوهرية}.


Just as in the time of Dr. Kellogg, when many brethren were departing from the simplicity of Christ’s teaching, so it is today. Sister White prophesied that this change over the understanding of the \emcap{personality of God} would occur in our church, and re-establishment of our old foundation of faith would be necessary. Will the \emcap{Fundamental Principles} be re-established in our midst? The outcome of it lies completely on every individual that makes up the body of the Seventh-day Adventist Church. Today, in this day and age, is when this re-establishment will take place. The warnings and rebukes uttered by the pen of the Spirit of Prophecy have never been more relevant than today. God has put the final outcome into your hands. If these warnings touch the core of your soul, God is calling you to stand firm on the platform of eternal Truth. He calls you to hold firmly to the \emcap{Fundamental Principles}, which are based on unquestionable authority.


كما كان الحال في زمن الدكتور كيلوغ، عندما كان العديد من الإخوة يبتعدون عن بساطة تعاليم المسيح، كذلك هو اليوم. تنبأت الأخت وايت بأن هذا التغيير في فهم \emcap{شخصانية الله} سيحدث في كنيستنا، وأن إعادة تأسيس أساس إيماننا القديم سيكون ضروريًا. هل ستتم إعادة تأسيس \emcap{المبادئ الجوهرية} بيننا؟ نتيجة ذلك تعتمد تمامًا على كل فرد يشكل جسد كنيسة الأدفنتست السبتيين. اليوم، في هذا العصر، هو الوقت الذي ستحدث فيه إعادة التأسيس هذه. لم تكن التحذيرات والتوبيخات التي صدرت عن قلم روح النبوة أكثر صلة من اليوم. لقد وضع الله النتيجة النهائية في يديك. إذا لمست هذه التحذيرات جوهر روحك، فإن الله يدعوك للوقوف بثبات على منصة الحق الأبدي. إنه يدعوك للتمسك بقوة بـ \emcap{المبادئ الجوهرية}، التي تستند إلى سلطة لا تقبل الجدل.


Below, we present a portion of the letter from Sister White to Dr. Kellogg, in which we find solemn warning for us today in re-establishing the foundation of our faith. When we are acquainted with the truth on the \emcap{personality of God} and its controversy over the Trinity doctrine, the following letter shines in a new light, with messages and principles that are valuable for us today, that we may know how to behave in our current state of affairs.


فيما يلي، نقدم جزءًا من رسالة الأخت وايت إلى الدكتور كيلوغ، والتي نجد فيها تحذيرًا جادًا لنا اليوم في إعادة تأسيس أساس إيماننا. عندما نكون على دراية بالحق حول \emcap{شخصانية الله} وصراعها مع عقيدة الثالوث، تشرق الرسالة التالية بنور جديد، مع رسائل ومبادئ قيمة لنا اليوم، حتى نعرف كيف نتصرف في حالتنا الراهنة.


\egw{\textbf{The Foundation of Our Faith}}


\egw{\textbf{أساس إيماننا}}


\egw{In regard to the book Living Temple, \textbf{I have been instructed by the heavenly messenger} that \textbf{some of the reasoning in this book is untrue, and that this reasoning would lead astray the minds of those who are not thoroughly established \underline{on the foundation principles} of present truth.} \textbf{It introduces that which is naught but speculation in regard to \underline{the personality of God} and \underline{where His presence is}}. No one on this earth has a right to speculate on this question. ‘The secret things belong unto the Lord our God, but those things which are revealed belong unto us, and to our children forever.’}[Lt232-1903.39; 1903][https://egwwritings.org/read?panels=p10197.48]


\egw{فيما يتعلق بكتاب ذا ليفينغ تمبل، \textbf{لقد تلقيت تعليمات من الرسول السماوي} بأن \textbf{بعض المنطق في هذا الكتاب غير صحيح، وأن هذا المنطق سيضل عقول أولئك الذين ليسوا راسخين تمامًا \underline{على المبادئ الأساسية} للحق الحاضر.} \textbf{إنه يقدم ما هو مجرد تكهنات فيما يتعلق \underline{بشخصانية الله} و\underline{أين يوجد حضوره}}. ليس لأحد على هذه الأرض الحق في التكهن بشأن هذا السؤال. ‘الأسرار للرب إلهنا، والمعلنات لنا ولأبنائنا إلى الأبد.’}[Lt232-1903.39; 1903][https://egwwritings.org/read?panels=p10197.48]


\egwnogap{\textbf{I am authorized by the Lord to say, The sentiments contained in Living Temple in regard to the personality of God \underline{are opposed to the truth that God has given us}}. \textbf{The truth for this time is now to be brought before the people.} Our brethren and sisters in every church and in every place are to guard carefully against allowing their minds to be engrossed with matters that draw them away from eternal things. The enemy will use some of the statements made in Living Temple to tempt some as he tempted Adam and Eve in Eden. \textbf{I warn our brethren not to enter into controversy over the presence and personality of God. \underline{The statements made in Living Temple in regard to this point are incorrect}.} The Scripture used to substantiate the doctrine there set forth is Scripture misapplied.}[Lt232-1903.40; 1903][https://egwwritings.org/read?panels=p10197.49]


\egwnogap{\textbf{أنا مفوضة من الرب لأقول، إن الآراء الواردة في ذا ليفينغ تمبل فيما يتعلق بشخصانية الله \underline{تعارض الحق الذي أعطانا الله إياه}}. \textbf{يجب الآن تقديم الحق لهذا الوقت للناس.} يجب على إخوتنا وأخواتنا في كل كنيسة وفي كل مكان أن يحرصوا على عدم السماح لعقولهم بالانشغال بأمور تبعدهم عن الأمور الأبدية. سيستخدم العدو بعض البيانات الواردة في ذا ليفينغ تمبل لإغواء البعض كما أغوى آدم وحواء في عدن. \textbf{أحذر إخوتنا من الدخول في جدل حول وجود وشخصانية الله. \underline{البيانات الواردة في ذا ليفينغ تمبل فيما يتعلق بهذه النقطة غير صحيحة}.} الكتاب المقدس المستخدم لدعم العقيدة المطروحة هناك هو كتاب مقدس مطبق بشكل خاطئ.}[Lt232-1903.40; 1903][https://egwwritings.org/read?panels=p10197.49]


\egwnogap{I was cautioned not to enter into controversy regarding the question that will come up over these things, \textbf{because controversy might lead men to resort to \underline{subterfuges}, and their minds would be led away from the truth of the Word of God \underline{to assumption and guesswork}. The more that fanciful theories are discussed, the less men will know of God and of the truth that sanctifies the soul}.}[Lt232-1903.41; 1903][https://egwwritings.org/read?panels=p10197.50]


\egwnogap{لقد حُذرت من الدخول في جدل بشأن المسألة التي ستظهر حول هذه الأمور، \textbf{لأن الجدل قد يؤدي بالناس إلى اللجوء إلى \underline{الحيل}، وستبتعد عقولهم عن حق كلمة الله \underline{إلى الافتراض والتخمين}. كلما نوقشت النظريات الخيالية أكثر، قل ما يعرفه الناس عن الله وعن الحق الذي يقدس النفس}.}[Lt232-1903.41; 1903][https://egwwritings.org/read?panels=p10197.50]


\egwnogap{We are God’s commandment-keeping people. \textbf{For the last fifty years every phase of heresy has been brought to bear upon us, to tear down \underline{the foundation principles of our faith}}. Messages of every order and kind have been urged upon Seventh-day Adventists to take the place of the truth which \textbf{point by point} has been testified to by the miracle-working power of the Lord. \textbf{But the waymarks which have made us what we are are to be preserved, and they \underline{will be preserved}, as God has signified through His Word and the testimony of His Spirit. From the great system of truth as it has been presented by God’s messengers, \underline{not a pin is to be removed}}.}[Lt232-1903.42; 1903][https://egwwritings.org/read?panels=p10197.51]


\egwnogap{نحن شعب الله الذي يحفظ وصاياه. \textbf{فخلال الخمسين سنة الماضية، تم توجيه كل نوع من أنواع الهرطقة علينا، لهدم \underline{المبادئ الأساسية لإيماننا}}. لقد تم حث رسائل من كل نظام ونوع على الأدفنتست السبتيين لتحل محل الحق الذي \textbf{نقطة بنقطة} تم الشهادة له بقوة الرب العاملة للمعجزات. \textbf{لكن المعالم التي جعلتنا ما نحن عليه يجب أن تُحفظ، وهي \underline{ستُحفظ}، كما أشار الله من خلال كلمته وشهادة روحه. من نظام الحق العظيم كما قدمه رسل الله، \underline{لا ينبغي إزالة دبوس واحد}}.}[Lt232-1903.42; 1903][https://egwwritings.org/read?panels=p10197.51]


\egwnogap{\textbf{I am called upon by God to stand in defense of the truth that has been given us as we have followed the leading of Him who is the way, the truth, and the life. Let every pioneer in the work adhere firmly to this truth. \underline{The peculiarities of our faith are to be held fast with the grip of faith}.}}[Lt232-1903.43; 1903][https://egwwritings.org/read?panels=p10197.52]


\egwnogap{\textbf{لقد دعاني الله للوقوف دفاعًا عن الحق الذي أُعطي لنا بينما اتبعنا قيادة ذاك الذي هو الطريق والحق والحياة. ليتمسك كل رائد في العمل بثبات بهذا الحق. \underline{يجب التمسك بخصوصيات إيماننا بقبضة الإيمان}}.}[Lt232-1903.43; 1903][https://egwwritings.org/read?panels=p10197.52]


\egwnogap{The fables that at the present time are being framed by some medical missionary workers are not to be regarded as truth. \textbf{\underline{Their true origin will ere long be revealed.}} \textbf{It will be seen that they were formed under the subtle power of the great apostate, who works as an angel of light, controlling minds by deceptions so concealed that he seeks by them to deceive if possible the very elect}.}[Lt232-1903.44; 1903][https://egwwritings.org/read?panels=p10197.53]


\egwnogap{الخرافات التي يتم صياغتها في الوقت الحاضر من قبل بعض العاملين في الإرسالية الطبية لا ينبغي اعتبارها حقيقة. \textbf{\underline{سيتم الكشف عن أصلها الحقيقي قريبًا.}} \textbf{سيُرى أنها تشكلت تحت القوة الخفية للمرتد العظيم، الذي يعمل كملاك نور، مسيطرًا على العقول بخداع مخفي لدرجة أنه يسعى من خلالها لخداع المختارين إن أمكن}.}[Lt232-1903.44; 1903][https://egwwritings.org/read?panels=p10197.53]


\egwnogap{What influence but that of the deceiver could lead men at \textbf{this stage of our history to work in an underhand, powerful way to tear down \underline{the foundations of our faith}—\underline{the foundations which were laid at the beginning of our work} by prayerful study of the Word and by revelation}. \textbf{\underline{Upon these foundations we have been building for the last fifty years}}. Shall a new foundation be built up by men to whom God has not granted the special experience He has granted to the men whom He ordained to establish the foundations of our faith? \textbf{The men who are striving to build up this false foundation may suppose that they have found a new way, and that they can lay a stronger foundation than that which has been laid. \underline{But this is a great deception}. \underline{Other foundation can no man lay than that which has been laid}.}}[Lt232-1903.45; 1903][https://egwwritings.org/read?panels=p10197.54]


\egwnogap{أي تأثير غير تأثير المخادع يمكن أن يقود الناس في \textbf{هذه المرحلة من تاريخنا للعمل بطريقة خفية وقوية لهدم \underline{أسس إيماننا}—\underline{الأسس التي وُضعت في بداية عملنا} من خلال الدراسة المصلية للكلمة ومن خلال الوحي}. \textbf{\underline{على هذه الأسس كنا نبني طوال الخمسين سنة الماضية}}. هل سيتم بناء أساس جديد من قبل أناس لم يمنحهم الله الخبرة الخاصة التي منحها للرجال الذين عينهم لتأسيس أسس إيماننا؟ \textbf{قد يظن الرجال الذين يسعون لبناء هذا الأساس الزائف أنهم وجدوا طريقًا جديدًا، وأنهم يستطيعون وضع أساس أقوى من ذلك الذي تم وضعه. \underline{لكن هذا خداع عظيم}. \underline{لا يستطيع أحد أن يضع أساسًا آخر غير الذي وُضع}}.}[Lt232-1903.45; 1903][https://egwwritings.org/read?panels=p10197.54]


\egwnogap{I am instructed to say to our people that in the past many have undertaken the building of a new faith, the establishment of new principles. But how long did their building stand? It soon fell to pieces; \textbf{for it was not founded upon the Rock}.}[Lt232-1903.46; 1903][https://egwwritings.org/read?panels=p10197.55]


\egwnogap{لقد أُرشدت أن أقول لشعبنا أنه في الماضي تعهد كثيرون ببناء إيمان جديد، وتأسيس مبادئ جديدة. لكن كم دام بناؤهم؟ سرعان ما تفكك إلى أجزاء؛ \textbf{لأنه لم يكن مؤسسًا على الصخرة}.}[Lt232-1903.46; 1903][https://egwwritings.org/read?panels=p10197.55]


\egwnogap{Did not the first disciples have to meet the sayings of men? Did they not have to listen to false theories and then stand firm, having done all, to stand, saying, ‘Other foundation can no man lay than that which is laid’? One class after another arose with false doctrines, because men were so little acquainted with God.}[Lt232-1903.47; 1903][https://egwwritings.org/read?panels=p10197.56]


\egwnogap{ألم يكن على التلاميذ الأوائل أن يواجهوا أقوال الناس؟ ألم يكن عليهم الاستماع إلى نظريات كاذبة ثم الوقوف بثبات، بعد أن فعلوا كل شيء، للوقوف، قائلين: “لا يستطيع أحد أن يضع أساسًا آخر غير الذي وُضع”؟ ظهرت فئة تلو الأخرى بتعاليم كاذبة، لأن الناس كانوا على معرفة قليلة جدًا بالله.}[Lt232-1903.47; 1903][https://egwwritings.org/read?panels=p10197.56]


\egwnogap{\textbf{My brethren and sisters, study the thirteenth, fourteenth, fifteenth, sixteenth, and seventeenth chapters of John. The words of these chapters explain themselves. ‘This is life eternal,’ Christ declared, ‘that they might know Thee the only true God, and Jesus Christ, whom Thou hast sent.’ \underline{In these words the personality of God and of His Son is clearly spoken of.} \underline{The personality of the one does not do away with the necessity for the personality of the other}.}}[Lt232-1903.48; 1903][https://egwwritings.org/read?panels=p10197.57]


\egwnogap{\textbf{إخوتي وأخواتي، ادرسوا الإصحاحات الثالث عشر والرابع عشر والخامس عشر والسادس عشر والسابع عشر من إنجيل يوحنا. كلمات هذه الإصحاحات تشرح نفسها. “وهذه هي الحياة الأبدية،“ أعلن المسيح، “أن يعرفوك أنت الإله الحقيقي وحدك ويسوع المسيح الذي أرسلته.” \underline{في هذه الكلمات يتم التحدث بوضوح عن شخصانية الله وابنه.} \underline{شخصانية أحدهما لا تلغي الحاجة إلى شخصانية الآخر}.}}[Lt232-1903.48; 1903][https://egwwritings.org/read?panels=p10197.57]


\egwnogap{God is never to be understood by any human being. His ways and His works are past finding out. In regard to the revelations \textbf{that He has made of Himself in His Word, we may talk}. \textbf{But when it comes to talking or writing of God’s person and presence, let us say, ‘Thou art God, and Thy ways are past finding out.’}}[Lt232-1903.49; 1903][https://egwwritings.org/read?panels=p10197.58]


\egwnogap{لا يمكن لأي إنسان أن يفهم الله أبدًا. طرقه وأعماله تفوق الإدراك. فيما يتعلق بالإعلانات \textbf{التي قدمها عن نفسه في كلمته، يمكننا أن نتحدث}. \textbf{لكن عندما يتعلق الأمر بالتحدث أو الكتابة عن شخص الله وحضوره، دعنا نقول: “أنت الله، وطرقك تفوق الإدراك.”}}[Lt232-1903.49; 1903][https://egwwritings.org/read?panels=p10197.58]


\egwnogap{It is sacrilegious to put into the minds of young or old the \textbf{seeds of \underline{speculation} regarding this subject}. Such seeds, planted and left to grow, will spring up and \textbf{bring forth a harvest of \underline{infidel sentiments}}. I give this warning to all. \textbf{We want no such sophistry as that presented in Living Temple}. There are excellent things in the book. But there are also tares among the wheat. The book contains many correct ideas, but it contains also statements that will do harm. Those who accept the chaff for the wheat will find themselves losing their sense of God’s greatness and bringing Him into cheap commonness. This is the work of the great deceiver. \textbf{Our brethren are not to be called from their work to study the question of where God is and what He is. We are not to dare to engage in this discussion, lest we be destroyed.} When the ark of God was being taken from the land of the Philistines to the camp of Israel, curiosity led the men of Bethshemesh to look into it. God was displeased, and many were smitten with death.}[Lt232-1903.50; 1903][https://egwwritings.org/read?panels=p10197.59]


\egwnogap{إنه من التجديف أن نضع في عقول الشباب أو الكبار \textbf{بذور \underline{التكهنات} حول هذا الموضوع}. مثل هذه البذور، إذا زُرعت وتُركت لتنمو، ستنبت و\textbf{تأتي بحصاد من \underline{الآراء الإلحادية}}. أعطي هذا التحذير للجميع. \textbf{لا نريد مثل هذه السفسطة المقدمة في كتاب ذا ليفينغ تمبل}. هناك أشياء ممتازة في الكتاب. لكن هناك أيضًا زوان بين القمح. أولئك الذين يقبلون التبن بدلاً من القمح سيجدون أنفسهم يفقدون إحساسهم بعظمة الله ويجعلونه في ابتذال رخيص. هذا هو عمل المخادع العظيم. \textbf{لا ينبغي استدعاء إخوتنا من عملهم لدراسة مسألة أين الله وما هو. لا يجب أن نجرؤ على الانخراط في هذه المناقشة، لئلا نُدمر.} عندما كان تابوت الله يُنقل من أرض الفلسطينيين إلى معسكر إسرائيل، قاد الفضول رجال بيت شمس للنظر إليه. غضب الله، وضُرب كثيرون بالموت.}[Lt232-1903.50; 1903][https://egwwritings.org/read?panels=p10197.59]


\egwnogap{\textbf{Let us talk of Christ, His preexistence, His humble ministry, His mighty power, His prospective personal glory in the heavenly courts. The Son of God restores to life whom He will. ‘All that the Father hath is Mine,’ He says. ‘I and My Father are one.’ He has greatness, present and prospective, that baffles human conception. He encircles the race with His long human arm, while with His divine arm He grasps the throne of the Infinite}.}[Lt232-1903.51; 1903][https://egwwritings.org/read?panels=p10197.60]


\egwnogap{\textbf{دعونا نتحدث عن المسيح، وجوده السابق، خدمته المتواضعة، قوته العظيمة، مجده الشخصي المرتقب في المحاكم السماوية. ابن الله يعيد إلى الحياة من يشاء. “كل ما للآب هو لي،“ يقول. “أنا والآب واحد.” لديه عظمة، حاضرة ومستقبلية، تتحدى الفهم البشري. إنه يحيط بالجنس البشري بذراعه البشرية الطويلة، بينما بذراعه الإلهية يمسك بعرش اللانهائي}.}[Lt232-1903.51; 1903][https://egwwritings.org/read?panels=p10197.60]


\egwnogap{\textbf{There is a knowledge of God and of Christ which all who are saved must have. ‘\underline{This is life eternal},’ Christ says, ‘\underline{that they might know Thee, the only true God, and Jesus Christ whom Thou hast sent}.’ And He says again, ‘If any man will come after Me, let him deny himself, and take up his cross, and follow Me.’ To all who receive Him as their Redeemer, He gives power to become the sons of God. Every one who truly believes in Him will be inspired by faith and raised by the arm of Omnipotence.}}[Lt232-1903.52; 1903][https://egwwritings.org/read?panels=p10197.61]


\egwnogap{\textbf{هناك معرفة بالله والمسيح يجب أن يمتلكها كل من يخلصون. ‘\underline{هذه هي الحياة الأبدية}،‘ يقول المسيح، ‘\underline{أن يعرفوك أنت الإله الحقيقي وحدك ويسوع المسيح الذي أرسلته}.’ ويقول أيضًا، ‘إن أراد أحد أن يأتي ورائي فلينكر نفسه ويحمل صليبه ويتبعني.’ لكل من يقبلونه كفاديهم، يعطيهم سلطانًا أن يصيروا أبناء الله. كل من يؤمن به حقًا سيلهم بالإيمان ويُرفع بذراع القدير.}}[Lt232-1903.52; 1903][https://egwwritings.org/read?panels=p10197.61]


\egwnogap{\textbf{Those who do not receive in faith God’s plan for redeeming the race} do despite to the Spirit of grace, and at the last great day their sentence will be, ‘Depart from Me.’ They have hated righteousness and fostered iniquity, and they must be banished forever from the presence of God, exiled from happiness to death—eternal death.}[Lt232-1903.53; 1903][https://egwwritings.org/read?panels=p10197.62]


\egwnogap{\textbf{أولئك الذين لا يقبلون بالإيمان خطة الله لفداء الجنس البشري} يهينون روح النعمة، وفي يوم الدينونة العظيم سيكون حكمهم، ‘اذهبوا عني.’ لقد أبغضوا البر وغذوا الإثم، ويجب أن يُنفوا إلى الأبد من حضرة الله، منفيين من السعادة إلى الموت - الموت الأبدي.}[Lt232-1903.53; 1903][https://egwwritings.org/read?panels=p10197.62]


\egwnogap{Those who in this life love God and cherish the thought of Him will employ their faculties and their talents as faithful stewards, making the very best use of them, but not claiming any reward as their due. \textbf{As they deny self and follow Jesus, lifting the cross, they will find that the cross is light, and that it is a pledge, as they bear it, that they will one day be given a crown of everlasting life}. What will be the glory and the gain and the enjoyment of that eternal life that is to be given to those only for whom it has been prepared? The great joy of the overcomer will be that he is in the presence of Christ. ‘Where I am, there shall also My servant be,’ He declared. And He prayed, ‘Father, I will that they also whom Thou hast given Me be with Me where I am; that they may behold My glory.’ Christ is speaking of the glory of His Father’s presence and His Father’s house. \textbf{The glory that is to be revealed to all who are saved is the glory which Christ had with His Father before the world was—\underline{the unapproachable splendor of their converse together}}. \textbf{The angels were not admitted to the interviews between the Father and the Son when the plan of salvation was laid.} Those human beings who seek to intrude into the secrets of the Most High, who inhabiteth eternity, show their ignorance of spiritual and eternal things. Far better might they, while mercy’s voice is still heard, humble themselves in the dust and plead with God to teach them His ways.}[Lt232-1903.54; 1903][https://egwwritings.org/read?panels=p10197.63]


\egwnogap{أولئك الذين يحبون الله في هذه الحياة ويعتزون بالتفكير فيه سيستخدمون قدراتهم ومواهبهم كوكلاء أمناء، مستفيدين منها أفضل استفادة، ولكن دون المطالبة بأي مكافأة كحق لهم. \textbf{وإذ ينكرون ذواتهم ويتبعون يسوع، حاملين الصليب، سيجدون أن الصليب خفيف، وأنه تعهد، عندما يحملونه، بأنهم سيُعطون يومًا ما إكليل الحياة الأبدية}. ما هو المجد والربح والمتعة في تلك الحياة الأبدية التي ستُعطى فقط لأولئك الذين أُعدت لهم؟ الفرح العظيم للغالب سيكون أنه في حضرة المسيح. “حيث أكون أنا هناك أيضًا يكون خادمي،“ أعلن. وصلى قائلاً، “أيها الآب أريد أن هؤلاء الذين أعطيتني يكونون معي حيث أكون أنا، لينظروا مجدي.” المسيح يتحدث عن مجد حضور أبيه وبيت أبيه. \textbf{المجد الذي سيُكشف لكل من يخلصون هو المجد الذي كان للمسيح مع أبيه قبل تأسيس العالم - \underline{البهاء الذي لا يُقترب منه في حديثهما معًا}}. \textbf{لم يُسمح للملائكة بالمشاركة في المقابلات بين الآب والابن عندما وُضعت خطة الخلاص.} أولئك البشر الذين يسعون للتدخل في أسرار العلي، الساكن في الأبدية، يُظهرون جهلهم بالأمور الروحية والأبدية. من الأفضل لهم بكثير، بينما لا يزال صوت الرحمة مسموعًا، أن يتواضعوا في التراب ويتضرعوا إلى الله ليعلمهم طرقه.}[Lt232-1903.54; 1903][https://egwwritings.org/read?panels=p10197.63]


\egw{A Timely Warning}


\egw{تحذير في وقته}


\egw{There are those who have been seeking to carry out their own selfish designs, without regard to the influence that this would have upon the cause and work of God. It is time that such ones felt the inward work of grace upon their hearts, that the medical missionary work may not be grossly misrepresented. Let not our medical missionary workers become so like the world in habit and practice that worldlings will turn away from them with scorn, saying, ‘I am just as good as they are.’ There are instances where the medical missionary work has been so conducted that the name ‘medical missionary’ might better be dropped; for it has been badly misrepresented, and God has been dishonored.}[Lt232-1903.55; 1903][https://egwwritings.org/read?panels=p10197.65]


\egw{هناك أولئك الذين كانوا يسعون لتنفيذ تصاميمهم الأنانية، دون اعتبار للتأثير الذي سيكون لهذا على قضية الله وعمله. حان الوقت لأن يشعر هؤلاء بعمل النعمة الداخلي في قلوبهم، حتى لا يُساء تمثيل العمل الطبي التبشيري بشكل فظيع. لا تدعوا عاملينا الطبيين التبشيريين يصبحون مثل العالم في العادات والممارسات لدرجة أن العالميين ينصرفون عنهم باحتقار، قائلين، ‘أنا جيد تمامًا مثلهم.’ هناك حالات حيث تم إجراء العمل الطبي التبشيري بطريقة تجعل من الأفضل إسقاط اسم ‘طبي تبشيري’؛ لأنه قد أُسيء تمثيله بشدة، وقد أُهين الله.}[Lt232-1903.55; 1903][https://egwwritings.org/read?panels=p10197.65]


\egwnogap{\textbf{We are living amidst the perils of the last days. Our people must now arouse to the work before them. We are to lift up the standard and proclaim the last message of warning to a perishing world. Those who have a knowledge of the truth for this time are now to hold firmly aloft the banner bearing the inscription, ‘The commandments of God and the faith of Jesus.’ }[Revelation 14:12.]}[Lt232-1903.56; 1903][https://egwwritings.org/read?panels=p10197.66]


\egwnogap{\textbf{نحن نعيش وسط مخاطر الأيام الأخيرة. يجب على شعبنا الآن أن ينهض للعمل الذي أمامهم. علينا أن نرفع المعيار ونعلن رسالة التحذير الأخيرة لعالم هالك. أولئك الذين لديهم معرفة بالحق لهذا الوقت يجب أن يرفعوا الآن بثبات الراية التي تحمل النقش، ‘وصايا الله وإيمان يسوع.’ }[رؤيا 14:12.]}[Lt232-1903.56; 1903][https://egwwritings.org/read?panels=p10197.66]


\egwnogap{I ask my ministering brethren to examine themselves, to see whether they are in the faith or not. \textbf{If they accept the spiritualistic representations made in Living Temple, their feet will soon be treading in forbidden paths. These representations are the Alpha of doctrines that would lead far away from the truth as we have received it from the Word of God}. \textbf{\underline{The acceptation of these sentiments will result in a weak, wavering faith}}. \textbf{If this is the teaching that is to be given in the medical missionary work, it will not be long before we have no foundation upon which to plant our feet}. \textbf{I am bidden to say that these erroneous sentiments are the sentiments of the wily foe} and should not be presented to any of our youth who are seeking to gain an education in medical missionary lines. For the sake of these youth, I speak decidedly.}[Lt232-1903.57; 1903][https://egwwritings.org/read?panels=p10197.67]


\egwnogap{أطلب من إخوتي في الخدمة أن يفحصوا أنفسهم، ليروا ما إذا كانوا في الإيمان أم لا. \textbf{إذا قبلوا التمثيلات الروحانية الموجودة في ذا ليفينغ تمبل، فستسير أقدامهم قريبًا في مسارات محظورة. هذه التمثيلات هي ألفا العقائد التي ستقود بعيدًا عن الحق كما تلقيناه من كلمة الله}. \textbf{\underline{قبول هذه الآراء سيؤدي إلى إيمان ضعيف متذبذب}}. \textbf{إذا كان هذا هو التعليم الذي سيُقدم في العمل الطبي التبشيري، فلن يمر وقت طويل قبل أن لا يكون لدينا أساس نقف عليه}. \textbf{لقد أُمرت أن أقول إن هذه الآراء الخاطئة هي آراء العدو الماكر ويجب ألا تُقدم لأي من شبابنا الذين يسعون للحصول على تعليم في مجالات العمل الطبي التبشيري}. من أجل هؤلاء الشباب، أتحدث بحزم.}[Lt232-1903.57; 1903][https://egwwritings.org/read?panels=p10197.67]


\egwnogap{\textbf{The expiring faith of the people of God \underline{must have a resurrection}}. \textbf{The exaltation of human reason has begun its work among us and has gone altogether too far}. Human reason is placed where divine, sanctifying truth should be. \textbf{Christ is waiting to kindle faith and love in the hearts of His people}. \textbf{Let not erroneous theories receive countenance from the people who ought to be standing firm on the platform of eternal truth.} \textbf{God calls upon us to hold firmly to \underline{the fundamental principles that are based upon unquestionable authority}}. He calls upon us to study the words and works of Christ, the greatest missionary that this world has ever known.}[Lt232-1903.58; 1903][https://egwwritings.org/read?panels=p10197.68]


\egwnogap{\textbf{إيمان شعب الله المحتضر \underline{يجب أن يقوم من جديد}}. \textbf{لقد بدأ تمجيد العقل البشري عمله بيننا وذهب بعيدًا جدًا}. العقل البشري وُضع حيث يجب أن يكون الحق الإلهي المقدس. \textbf{المسيح ينتظر ليشعل الإيمان والمحبة في قلوب شعبه}. \textbf{لا تدعوا النظريات الخاطئة تلقى تشجيعًا من الناس الذين ينبغي أن يكونوا واقفين بثبات على منصة الحق الأبدي.} \textbf{الله يدعونا للتمسك بقوة \underline{بالمبادئ الأساسية التي تستند إلى سلطة لا تقبل الشك}}. إنه يدعونا لدراسة كلمات وأعمال المسيح، أعظم مبشر عرفه هذا العالم.}[Lt232-1903.58; 1903][https://egwwritings.org/read?panels=p10197.68]


\egwnogap{\textbf{When the mind of a teacher of truth becomes in any way divorced from plain, self-denying gospel truth, he is prepared to receive fanciful sentiments called truth. Arrayed in the garments of light, these sentiments are presented to others, and too often they find favor. I am instructed to say to the members of our churches, Keep away from spiritualistic ideas}. \textbf{We are not dealing in fables}. God forbid that fables in the disguise of truth shall be presented to our people. \textbf{God forbid that any among us shall build upon the sand}.}[Lt232-1903.59; 1903][https://egwwritings.org/read?panels=p10197.69]


\egwnogap{\textbf{عندما ينفصل عقل معلم الحق بأي شكل من الأشكال عن حق الإنجيل البسيط المنكر للذات، يكون مستعدًا لتلقي آراء خيالية تسمى حقًا. مرتدية ثياب النور، تُقدم هذه الآراء للآخرين، وغالبًا ما تلقى قبولًا. لقد أُمرت أن أقول لأعضاء كنائسنا، ابتعدوا عن الأفكار الروحانية}. \textbf{نحن لا نتعامل مع الخرافات}. الله لا يسمح بأن تُقدم الخرافات المتنكرة بالحق لشعبنا. \textbf{الله لا يسمح بأن يبني أي منا على الرمل.}}[Lt232-1903.59; 1903][https://egwwritings.org/read?panels=p10197.69]


\egwnogap{\textbf{The Lord has given us a clear, distinct message of truth for this time}. Let us proclaim this message. \textbf{Let us study the teaching of Christ}, and present what He has commanded us to present. \textbf{He who launches out in his own wisdom to preach strange things, which God has not given him, finds minds ready to be leavened with the new ideas that he has to present}. \textbf{Satan follows up the work that he does, and the efforts of the true servants of God are made much harder}. \textbf{The advancement of His cause is hindered, and His Spirit is grieved}.}[Lt232-1903.60; 1903][https://egwwritings.org/read?panels=p10197.70]


\egwnogap{\textbf{لقد أعطانا الرب رسالة حق واضحة ومميزة لهذا الوقت}. دعونا نعلن هذه الرسالة. \textbf{دعونا ندرس تعاليم المسيح}، ونقدم ما أمرنا بتقديمه. \textbf{من ينطلق بحكمته الخاصة ليعظ بأمور غريبة، لم يعطها الله له، يجد عقولًا مستعدة لأن تُخمر بالأفكار الجديدة التي لديه ليقدمها}. \textbf{الشيطان يتابع العمل الذي يقوم به، وتصبح جهود خدام الله الحقيقيين أكثر صعوبة}. \textbf{يُعاق تقدم قضيته ويحزن روحه.}}[Lt232-1903.60; 1903][https://egwwritings.org/read?panels=p10197.70]


We pray that God will speak decisively clear into everyone's heart reading these warnings, to keep from stepping off the foundation of our faith. God calls upon us to hold firmly to the \emcap{Fundamental Principles} that are based upon unquestionable authority.


نصلي أن يتكلم الله بوضوح حاسم في قلب كل من يقرأ هذه التحذيرات، ليحفظهم من الخروج عن أساس إيماننا. يدعونا الله للتمسك بثبات بـ\emcap{المبادئ الجوهرية} التي تستند إلى سلطة لا تقبل الجدل.


\input{lang/en/poems/chapter29}