% Setting up the wrong Fundamental Principles

\begin{titledpoem}
    
    \stanza{
        In the silence of our own church’s walls, \\
        We’ve strayed from the light that first on us falls. \\
        An ancient query resounds with age: \\
        Why from sacred paths do we disengage?
    }

    \stanza{
        Ellen spoke of priests and rabbis of old, \\
        Men of cloth, with hearts yet so cold. \\
        They spurned the truth when it brightly shone, \\
        Proud hearts rejecting what was divinely shown.
    }

    \stanza{
        Behold the fishermen with unlearned tongues, \\
        Who dared to challenge the learned ones. \\
        "The Holy Spirit guides," they boldly claimed, \\
        While leaders scorned and fiercely blamed.
    }

    \stanza{
        Do not the humble hearts perceive \\
        The quiet whispers they should believe? \\
        While proud hearts preach their own command, \\
        True faith slips like fine sand from hand.
    }

    \stanza{
        Even now, echoes of the past, \\
        Warn us of shadows that leaders cast. \\
        God’s own voice, through years does span, \\
        Yet resisted by the inventions of man.
    }

    \stanza{
        To the foundations we must hold fast, \\
        Not swayed by shadows that leaders cast. \\
        For in the Scripture’s unerring light, \\
        Lies the path that is just and right.
    }

    \stanza{
        Beware of new doctrines, thinly veiled, \\
        On old, firm rocks they have not sailed. \\
        Let not man’s wisdom lead astray, \\
        But God’s own Spirit show the way.
    }

    \stanza{
        In faith, let us each day commence, \\
        With Bible as shield, no human pretense. \\
        For truth in Christ alone is found, \\
        And on this rock, our faith is sound.
    }
\end{titledpoem}