\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter17}{Reply to Kellogg’s trinitarian sentiments}


\qrchapter{https://forgottenpillar.com/rsc/en-fp-chapter17}{Réponse au raisonnement trinitaire de Kellogg}


If we look at the Kellogg crisis through the perspective of the \emcap{personality of God} and the \emcap{Fundamental Principles}, Sister White’s quotations inevitably shine in a new light. In this light we see the conflict between the truth we have received in the beginning, on the \emcap{personality of God}, and the Trinity doctrine. In order to avoid discrepancy, in the interest of defending the Trinity doctrine, scholars always overemphasize the pantheistic side of the problem.


Si nous regardons la crise de Kellogg à travers la perspective de la \emcap{personnalité de Dieu} et des \emcap{Principes Fondamentaux}, les citations de Sœur White brillent inévitablement sous un jour nouveau. Dans cette lumière, nous voyons le conflit entre la vérité que nous avons reçue au commencement, sur la \emcap{personnalité de Dieu}, et la doctrine de la Trinité. Afin d'éviter toute divergence, dans l'intérêt de défendre la doctrine de la Trinité, les érudits mettent toujours trop l'accent sur le côté panthéiste du problème.


We would like to challenge this tendency to overemphasize the pantheistic side of Kellogg’s controversy. Sister White generally wrote proactive truth; she approached the error by uplifting the truth. This is why she wrote so much about the \emcap{personality of God}. In most of her quotations on this subject, we see her dispelling the Trinitarian error, rather than pantheistic error. We read one such example where she establishes the truth on the \emcap{personality of God} referencing the seventeenth chapter of John.


Nous aimerions contester cette tendance à trop mettre l'accent sur le côté panthéiste de la controverse de Kellogg. Sœur White écrivait généralement la vérité de manière proactive ; elle approchait l'erreur en élevant la vérité. C'est pourquoi elle a tant écrit sur la \emcap{personnalité de Dieu}. Dans la plupart de ses citations sur ce sujet, nous la voyons dissiper l'erreur trinitaire, plutôt que l'erreur panthéiste. Nous lisons un tel exemple où elle établit la vérité sur la \emcap{personnalité de Dieu} en faisant référence au dix-septième chapitre de Jean.


\egw{\textbf{The personality of the Father and the Son, also the unity that exists between Them, are presented in the seventeenth chapter of John}, in the prayer of Christ for His disciples:}[MH 421.7; 1905][https://egwwritings.org/read?panels=p135.2173]


\egw{\textbf{La personnalité du Père et du Fils, ainsi que l'unité qui existe entre Eux, sont présentées dans le dix-septième chapitre de Jean}, dans la prière du Christ pour Ses disciples :}[MH 421.7; 1905][https://egwwritings.org/read?panels=p135.2173]


There are many cases where Sister White quotes John 17 in regard to Kellogg’s crisis. Those who assert that Kellogg’s crisis was solely about pantheism should inquire how John 17 addresses God in nature. And it is not only John 17, but also chapters 13-16. In her letter to Kellogg, she wrote:


Il y a de nombreux cas où Sœur White cite Jean 17 en ce qui concerne la crise de Kellogg. Ceux qui affirment que la crise de Kellogg concernait uniquement le panthéisme devraient se demander comment Jean 17 aborde Dieu dans la nature. Et ce n'est pas seulement Jean 17, mais aussi les chapitres 13-16. Dans sa lettre à Kellogg, elle a écrit :


\egw{\textbf{\underline{…study the thirteenth, fourteenth, fifteenth, sixteenth, and seventeenth chapters of John}. The words of these chapters explain themselves. ‘This is life eternal,’ Christ declared, ‘that they might know \underline{Thee the only true God}, and Jesus Christ, whom Thou hast sent.’ \underline{In these words the personality of God and of His Son is clearly spoken of.} \underline{The personality of the one does not do away with the necessity for the personality of the other}.}}[Lt232-1903.48, 1903][https://egwwritings.org/read?panels=p10197.57]


\egw{\textbf{\underline{…étudiez les treizième, quatorzième, quinzième, seizième et dix-septième chapitres de Jean}. Les paroles de ces chapitres s'expliquent d'elles-mêmes. « C'est ici la vie éternelle », a déclaré Christ, « qu'ils Te connaissent, \underline{Toi, le seul vrai Dieu}, et Jésus-Christ, que Tu as envoyé. » \underline{Dans ces paroles, la personnalité de Dieu et de Son Fils est clairement exprimée.} \underline{La personnalité de l'un ne supprime pas la nécessité de la personnalité de l'autre}.}}[Lt232-1903.48, 1903][https://egwwritings.org/read?panels=p10197.57]


In the aforementioned chapters of John, John did not reference anything pertaining to God in nature. The content of those chapters covers who is the only true God, how the Father and the Son are one, their true relation, and how Jesus can be everywhere present yet will ascend to the Father.


Dans les chapitres susmentionnés de Jean, Jean n'a fait référence à rien concernant Dieu dans la nature. Le contenu de ces chapitres couvre qui est le seul vrai Dieu, comment le Père et le Fils sont un, leur vraie relation, et comment Jésus peut être partout présent tout en montant vers le Père.


\egw{Jesus said to the Jews: ‘My Father worketh hitherto, and I work.... The Son can do nothing of Himself, but what He seeth the Father do: for what things soever He doeth, these also doeth the Son likewise. For the Father loveth the Son, and showeth Him all things that Himself doeth.’ John 5:17-20.}[8T 268.4, 1904][https://egwwritings.org/read?panels=p112.1557]


\egw{Jésus dit aux Juifs : « Mon Père travaille jusqu'à présent, et je travaille aussi.... Le Fils ne peut rien faire de lui-même, mais seulement ce qu'il voit faire au Père ; car tout ce que le Père fait, le Fils aussi le fait pareillement. Car le Père aime le Fils, et lui montre tout ce qu'il fait lui-même. » Jean 5:17-20.}[8T 268.4, 1904][https://egwwritings.org/read?panels=p112.1557]


\egwnogap{\textbf{Here again is brought to view the \underline{personality of the Father and the Son}, showing the unity that exists between them}.}[8T 269.1; 1904][https://egwwritings.org/read?panels=p112.1560]


\egwnogap{\textbf{Ici encore est mise en évidence la \underline{personnalité du Père et du Fils}, montrant l'unité qui existe entre eux}.}[8T 269.1; 1904][https://egwwritings.org/read?panels=p112.1560]


\egwnogap{\textbf{This unity is expressed also in \underline{the seventeenth chapter of John}}, in the prayer of Christ for His disciples:}[8T 269.2; 1904][https://egwwritings.org/read?panels=p112.1561]


\egwnogap{\textbf{Cette unité est également exprimée dans \underline{le dix-septième chapitre de Jean}}, dans la prière du Christ pour Ses disciples :}[8T 269.2; 1904][https://egwwritings.org/read?panels=p112.1561]


\egwnogap{‘Neither pray I for these alone, but for them also which shall believe on Me through their word; that they all may be one; \textbf{as Thou, Father, art in Me, and I in Thee, that they also may be one in Us}: that the world may believe that Thou hast sent Me. And \textbf{the glory which Thou gavest Me} I have given them; \textbf{that they may be one, even as We are one: I in them, and Thou in Me, that they may be made perfect in one}; and that the world may know that Thou hast sent Me, and hast loved them, as Thou hast loved Me.’ John 17:20-23.}[8T 269.3; 1904][https://egwwritings.org/read?panels=p112.1562]


\egwnogap{‘Je ne prie pas pour eux seulement, mais aussi pour ceux qui croiront en Moi par leur parole, afin que tous soient un, \textbf{comme Toi, Père, Tu es en Moi, et comme Je suis en Toi, afin qu'eux aussi soient un en Nous}, pour que le monde croie que Tu M'as envoyé. \textbf{Je leur ai donné la gloire que Tu M'as donnée}, \textbf{afin qu'ils soient un comme Nous sommes un, Moi en eux, et Toi en Moi, afin qu'ils soient parfaitement un}, et que le monde connaisse que Tu M'as envoyé et que Tu les as aimés comme Tu M'as aimé.’ Jean 17:20-23.}[8T 269.3; 1904][https://egwwritings.org/read?panels=p112.1562]


\egwnogap{Wonderful statement! \textbf{The unity that exists between Christ and His disciples \underline{does not destroy the personality of either}. They are one in purpose, in mind, in character, but \underline{not in person}. It is thus that God and Christ are one}.}[8T 269.4; 1904][https://egwwritings.org/read?panels=p112.1563]


\egwnogap{Déclaration merveilleuse ! \textbf{L'unité qui existe entre Christ et Ses disciples \underline{ne détruit la personnalité d'aucun d'eux}. Ils sont un dans le but, dans l'esprit, dans le caractère, mais \underline{non en personne}. C'est ainsi que Dieu et Christ sont un}.}[8T 269.4; 1904][https://egwwritings.org/read?panels=p112.1563]


\egwnogap{\textbf{The relation between the Father and the Son, and the personality of both, are made plain in this scripture also}:}[8T 269.5; 1904][https://egwwritings.org/read?panels=p112.1564]


\egwnogap{\textbf{La relation entre le Père et le Fils, et la personnalité des deux, sont rendues claires dans cette écriture également} :}[8T 269.5; 1904][https://egwwritings.org/read?panels=p112.1564]


\egwnogap{Thus speaketh \textbf{Jehovah of hosts}, saying,} \\
\egw{Behold, \textbf{the man} whose name is\textbf{ the Branch}:} \\
\egw{And He shall grow up out of His place;} \\
\egw{\textbf{And He shall build the temple of Jehovah;... }} \\
\egw{\textbf{And He shall bear the glory,}} \\
\egw{\textbf{And shall sit and rule upon His throne;}} \\
\egw{\textbf{And He shall be a priest upon His throne;}} \\
\egw{\textbf{And \underline{the counsel of peace shall be between Them both}}.’}[8T 269.6; 1904][https://egwwritings.org/read?panels=p112.1565]


\egwnogap{Ainsi parle \textbf{l'Éternel des armées}, disant,} \\
\egw{Voici, \textbf{l'homme} dont le nom est \textbf{le Germe} :} \\
\egw{Et Il germera de Son lieu ;} \\
\egw{\textbf{Et Il bâtira le temple de l'Éternel ;... }} \\
\egw{\textbf{Et Il portera la gloire,}} \\
\egw{\textbf{Et s'assiéra et dominera sur Son trône ;}} \\
\egw{\textbf{Et Il sera sacrificateur sur Son trône ;}} \\
\egw{\textbf{Et \underline{le conseil de paix sera entre Eux deux}}.’}[8T 269.6; 1904][https://egwwritings.org/read?panels=p112.1565]


The aforementioned chapters of the Gospel of John deal with the \emcap{personality of God}, which had been expressed in the first two points of the \emcap{Fundamental Principles}. What error did Sister White combat when she referenced verses on how the Father was the only true God, and how the Father and the Son are not one in person? Pantheism? Certainly not; but most probably the trinitarian sentiments, or belief in a one-in-three, or three-in-one God.


Les chapitres susmentionnés de l'Évangile de Jean traitent de la \emcap{personnalité de Dieu}, qui avait été exprimée dans les deux premiers points des \emcap{Principes Fondamentaux}. Quelle erreur Sœur White combattait-elle lorsqu'elle faisait référence aux versets sur comment le Père était le seul vrai Dieu, et comment le Père et le Fils ne sont pas un en personne ? Le panthéisme ? Certainement pas ; mais très probablement le raisonnement trinitaire, ou la croyance en un Dieu un-en-trois, ou trois-en-un.


Brother J. N. Loughborough, one of the first brethren who wrote on the \emcap{personality of God}, wrote the following comment on John chapter 17:


Frère J. N. Loughborough, l'un des premiers frères qui a écrit sur la \emcap{personnalité de Dieu}, a écrit le commentaire suivant sur le chapitre 17 de Jean :


\others{\textbf{\underline{The seventeenth chapter of John is alone sufficient to refute the doctrine of the Trinity}}. \textbf{...\underline{Read the seventeenth chapter of John, and see if it does not completely upset the doctrine of the Trinity}}.}[John N. Loughborough, The Adventist Review, and Sabbath Herald, November 5, 1861, p. 184.10][https://egwwritings.org/read?panels=p1685.6615]


\others{\textbf{\underline{Le dix-septième chapitre de Jean est à lui seul suffisant pour réfuter la doctrine de la Trinité}}. \textbf{...\underline{Lisez le dix-septième chapitre de Jean, et voyez s'il ne renverse pas complètement la doctrine de la Trinité}}.}[John N. Loughborough, The Adventist Review, and Sabbath Herald, November 5, 1861, p. 184.10][https://egwwritings.org/read?panels=p1685.6615]


Sister White’s proactive writing in support of the truth on the \emcap{personality of God} and His presence is the same as other Adventist pioneers. If Adventist pioneers were debunking the Trinity doctrine by exalting the truth on the \emcap{personality of God} and God’s presence, what makes us think Ellen White was not doing the same, when the theological side of the question of the Trinity was raised? By stating this, we do not deny the pantheistic side of Kellogg’s controversy, but by overemphasizing it, it falls short of accurately describing its real issue. The correct understanding of the Kellogg controversy can only be accomplished by focusing primarily on the truth Sister White uplifted, rather than focusing on error, whether pantheism or Trinity. This truth that Sister White uplifted was the truth on the \emcap{personality of God} and where His presence is. This is expressed in the first point of the \emcap{Fundamental Principles}, which were the official synopsis and representation of Seventh-day Adventist beliefs in the time of Ellen White; the truth which we, as a church, \egwinline{have received and heard and advocated}[Ms124-1905.12; 1905][https://egwwritings.org/read?panels=p9099.18] in the beginning.


L'écriture proactive de Sœur White en soutien de la vérité sur la \emcap{personnalité de Dieu} et Sa présence est la même que celle des autres pionniers adventistes. Si les pionniers adventistes démystifiaient la doctrine de la Trinité en exaltant la vérité sur la \emcap{personnalité de Dieu} et la présence de Dieu, qu'est-ce qui nous fait penser qu'Ellen White ne faisait pas de même, lorsque le côté théologique de la question de la Trinité était soulevé ? En déclarant cela, nous ne nions pas le côté panthéiste de la controverse de Kellogg, mais en le suraccentuant, on ne parvient pas à décrire avec précision son véritable problème. La compréhension correcte de la controverse de Kellogg ne peut être accomplie qu'en se concentrant principalement sur la vérité que Sœur White a élevée, plutôt qu'en se concentrant sur l'erreur, qu'il s'agisse du panthéisme ou de la Trinité. Cette vérité que Sœur White a élevée était la vérité sur la \emcap{personnalité de Dieu} et où se trouve Sa présence. Ceci est exprimé dans le premier point des \emcap{Principes Fondamentaux}, qui étaient le synopsis officiel et la représentation des croyances adventistes du septième jour au temps d'Ellen White ; la vérité que nous, en tant qu'église, \egwinline{avons reçue et entendue et défendue}[Ms124-1905.12; 1905][https://egwwritings.org/read?panels=p9099.18] au commencement.


\egw{\textbf{I entreat every one to be clear and firm regarding the certain truths that we have received and heard and advocated. The statements of God’s Word are plain. Plant your feet firmly on \underline{the platform of eternal truth}. \underline{Reject every phase of error}, even \underline{though it be covered with a semblance of reality, which denies the personality of God or of Christ}}.}[Ms124-1905.12; 1905][https://egwwritings.org/read?panels=p9099.18]


\egw{\textbf{Je supplie chacun d'être clair et ferme concernant les vérités certaines que nous avons reçues et entendues et défendues. Les déclarations de la Parole de Dieu sont claires. Plantez fermement vos pieds sur \underline{la plateforme de la vérité éternelle}. \underline{Rejetez toute phase d'erreur}, même \underline{si elle est couverte d'un semblant de réalité, qui nie la personnalité de Dieu ou de Christ}}.}[Ms124-1905.12; 1905][https://egwwritings.org/read?panels=p9099.18]


The warning from the previous quotations did not lessen in the course of time. Today it is even more relevant. We should \egwinline{reject every phase of error, even though it be covered with a semblance of reality, which denies the personality of God or of Christ}. In the following chapter we want to point out to the specific phase of error that is covered with a semblance of reality, which denies the personality of God and of Christ—three living persons of \textit{one} God, as opposed to \egwinline{three living persons of the heavenly trio.}[Ms21-1906.11; 1906][https://egwwritings.org/read?panels=p9754.18]


L'avertissement des citations précédentes n'a pas diminué au fil du temps. Aujourd'hui, il est encore plus pertinent. Nous devrions \egwinline{rejeter toute phase d'erreur, même si elle est couverte d'un semblant de réalité, qui nie la personnalité de Dieu ou de Christ}. Dans le chapitre suivant, nous voulons souligner la phase spécifique d'erreur qui est couverte d'un semblant de réalité, qui nie la personnalité de Dieu et de Christ — trois personnes vivantes d’\textit{un seul} Dieu, par opposition à \egwinline{trois personnes vivantes du trio céleste.}[Ms21-1906.11; 1906][https://egwwritings.org/read?panels=p9754.18]


\input{lang/en/poems/chapter17}