% Remembering the beginning

\begin{titledpoem}
    
    \stanza{
        In faith's first light, they sought His face, \\
        A vision pure, of divine grace. \\
        The pioneers, with vision clear, \\
        In 1844, held God so dear.
    }

    \stanza{    
        "The hour of judgment has come," they cried, \\
        To a world, both far and wide. \\
        The Ancient of Days, they did proclaim, \\
        Not a trinity, but a singular name.
    }

    \stanza{    
        Ellen White, with pen in hand, \\
        Spoke of a sanctuary, not of earthly land. \\
        A message of heaven, pure and bright, \\
        Guiding the faithful through the night.
    }

    \stanza{    
        The first angel’s message, a call to revere, \\
        God the Father, whom we should fear. \\
        "Who is the God we are to adore?" \\
        Not a trinity—this, they implored.
    }

    \stanza{    
        The Trinity, a concept unembraced, \\
        By pioneers, who in God's word traced. \\
        The Father, the Ancient, they did declare, \\
        His judgment and mercy, beyond compare.
    }

    \stanza{
        Yet, whispers now, through time have spread, \\ 
        A trinity's shadow, causing dread. \\
        If this be the God they were to declare, \\
        Their mission failed, caught in despair.
    }

    \stanza{
        But this is a falsehood, bold and cold, \\
        A narrative modern, but wrongly told. \\
        The Father they worshiped, with fervent zeal, \\
        Was the true God, their mission real.
    }

    \stanza{
        In unity, may we seek His face, \\
        Embracing truth, with grace and grace. \\
        The pioneers' vision, let us not lose, \\
        For in their footsteps, we must choose.
    }

    \stanza{
        To worship the God, of days of old, \\
        The Ancient of Days, as was foretold. \\
        In Revelation’s message, clear and bright, \\
        Guiding us still, through darkest night.
    }
    
\end{titledpoem}