% Adventist pioneers and the Trinity doctrine

\begin{titledpoem}
    \stanza{
        The pioneers stood firm against the Trinity's sway, \\
        Their testimony clear as the light of day. \\
        They saw how this doctrine did subtly disguise, \\
        What Scripture reveals to discerning eyes.
    }

    \stanza{
        James White declared it among "popular fables" found, \\
        A teaching that makes God's personality unsound. \\
        For Father and Son as distinct beings stand, \\
        Not merged into one as the Trinity planned.
    }

    \stanza{
        Two separate persons with purpose aligned, \\
        In spirit and action, united in mind. \\
        Just as disciples in Christ become one, \\
        Yet remain individual under God's Son.
    }

    \stanza{
        The Holy Spirit, God's presence divine, \\
        Not a separate being of the Godhead's design. \\
        But truly God's Spirit sent forth from above, \\
        The Father's representative, agent of love.
    }

    \stanza{
        No Scripture supports what tradition declared, \\
        When at Nicea this doctrine was shared. \\
        John seventeen shatters the trinitarian view, \\
        Revealing the Father and Son as beings true.
    }

    \stanza{
        Christ's divinity rests not in mysterious three, \\
        But in His begotten Sonship we see. \\
        The express image of His Father's face, \\
        Inheriting fullness of divine grace.
    }

    \stanza{
        The pioneers knew what the Bible made clear, \\
        That God is the Father, a Being most dear. \\
        A personal, spiritual presence on high, \\
        Omnipresent through Spirit, yet dwelling on high.
    }

    \stanza{
        So stands the truth against error's long night, \\
        Preserved by the pioneers and Sister White. \\
        Not through confusion or mystical thought, \\
        But through plain Scripture the truth has been taught.
    }
\end{titledpoem}