% Setting up the wrong Fundamental Principles

\begin{titledpoem}
    
    \stanza{
        Sadly, within our own church walls \\
        From our own pulpits, error falls \\
        Members want smooth words for their ears \\
        Don’t step on toes, Allay their fears.
    }

    \stanza{
        Pastors and elders do preside \\
        While sins remain, untouched inside \\
        Laodicean comfort zone \\
        But they will reap what they have sown.
    }

    \stanza{
        Ask for the old paths, walk therein \\
        From the old truth, don’t move a pin. \\
        They spurned the truth which brightly shone, \\
        The Spirit’s pow’r, to them unknown.
    }

    \stanza{
        Do not the humble hearts perceive \\
        Whispers of truth they should believe? \\
        Meanwhile the stories ease concern. \\
        What God would tell them, they won’t learn.
    }

    \stanza{
        Beware of error, thinly veiled, \\
        God’s Word is true, but men have failed. \\
        Beware of shadows leaders cast. \\
        To the foundations true, hold fast,
    }

    \stanza{
        Let not man’s wisdom lead astray, \\
        Let God’s own Spirit show the way. \\
        For in the Scripture’s glowing light, \\
        We find the path of safety bright.
    }

    \stanza{
        Let us, in faith, each day commence, \\
        God’s Word our shield, not man’s pretense. \\
        For truth in Christ alone is found, \\
        And on this rock, our faith is sound.
    }
    
\end{titledpoem}