% Steps to Omega

\begin{titledpoem}
    
    \stanza{
        Upon the pillars of old, shadows now cast— \\
        A truth forsaken, as whispers of the past. \\
        In steps they chart the course of silent drift, \\
        Five marks of change, through faith's sacred rift.
    }

    \stanza{    
        1. Denial blooms where once truth firmly stood, \\
        Foundations lost, veiled under falsehood’s hood. \\
        The Fundamental Principles, once held so dear, \\
        Now obscured, as new doctrines appear.
    }

    \stanza{
        2. Warnings ignored, the prophetic voice dimmed, \\
        New structures rise, with the old hymns trimmed. \\
        Ellen's caution, once a bright, guiding flare, \\
        Fades in the mist of doctrinal air.
    }

    \stanza{
        3. The personality of God, a cornerstone cast aside, \\
        In scholarly debates, His essence they deride. \\
        Once a pillar strong, now questioned in the dark, \\
        Leaving faith's vessel absent of its holy mark.
    }

    \stanza{
        4. Altered meanings twist the sacred term, \\
        God's "personality" redefined, scholars affirm. \\
        No longer seen as the Father's clear face, \\
        But lost in abstract, a confusing embrace.
    }

    \stanza{
        5. Focus shifted, from Divine to pantheistic lore, \\
        Kellogg's crisis viewed in light never seen before. \\
        Neglecting God’s persona, they divert the gaze, \\
        To lesser quarrels, setting minds ablaze.
    }

    \stanza{
        Thus stands the omega, in shadows deep and wide, \\
        A truth once vivid, now struggling to abide. \\
        Remember, O faithful, the anchors of your creed, \\
        For only in true foundations, will your spirits be freed.
    }
\end{titledpoem}