% Dr. Kellogg and Ellen White writings

\begin{titledpoem}
    \stanza{
        In faith's foundation, once so clear, \\
        Ellen White's words, we ought to revere. \\
        Kellogg claimed harmony, yet in deceit, \\
        Misrepresented truths, in his conceit.
    }

    \stanza{
        "The enemy seeks," she solemnly warned, \\
        To twist our beliefs, until they're scorned. \\
        Spiritualistic theories, cleverly dressed, \\
        In Scripture's garb, they falsely impressed.
    }

    \stanza{
        "False!" she declared, against the tide, \\
        Her writings and the Trinity, wrongly tied. \\
        Her defense was strong, her vision, broad, \\
        The Fundamental Principles, by God, lauded.
    }

    \stanza{    
        Not a word changed, nor principle dimmed, \\
        Since 1844, when light on faith brimmed. \\
        The pillars of truth, so firmly believed, \\
        By misinterpretations, were not deceived.
    }

    \stanza{
        Kellogg's stance, a misrepresentation, \\
        Of the faith's core, and its foundation. \\
        The personality of God, so crucial, so dear, \\
        By his theories, was muddled, we all fear.
    }

    \stanza{   
        Yet, Ellen White stood, unyielding, firm, \\
        Against falsehoods, her teachings confirm. \\
        Against the tide of Trinity's doctrine, she stood, \\
        Uplifting the truth of God and His Son, as she should.
    }

    \stanza{    
        Today, as then, let's hold the line, \\
        To the original faith, divine. \\
        Ellen White's legacy, let's rightly claim, \\
        In truth and spirit, always the same.
    }

    \stanza{    
        In battles of faith, let's not forget, \\
        Her defense of principles, firmly set. \\
        Against the currents of change and doubt, \\
        Her writings guide, without and within, throughout.
    }
\end{titledpoem}