% The Heavenly Trio

\begin{titledpoem}
    
    \stanza{
        In heaven’s realm, where truths unfold, \\
        A message clear, so brave and bold. \\
        God spoke through Ellen, clear and bright, \\
        Revealing depths of heavenly light.
    }

    \stanza{
        Misunderstood by some who read, \\
        Her words of God that all must heed. \\
        Not as triune, but trio three \\
        Distinct as persons, heavenly.
    }

    \stanza{
        The Father, not a formless feel, \\
        Invisible to us, yet real. \\
        He is the fullness, all complete, \\
        The Godhead, bodily, concrete.
    }

    \stanza{
        The Son, God’s fullness, manifest \\
        In Him, divinity does rest. \\
        God’s character, seen in His face, \\
        In Christ, we see His Father’s grace.
    }

    \stanza{
        The Spirit, in all fullness dwells, \\
        A mystery nature, Ellen tells. \\
        With forms, the Father and His Son \\
        With Them, in Spirit, we are one.
    }

    \stanza{
        Distinct and clear, Their roles unfold, \\
        The Father, Son, in form behold. \\
        Yet present everywhere we find, \\
        Their Spirit shows Their heart and mind.
    }

    \stanza{
        God’s message true, from up above. \\
        Reveals to us the Father’s love. \\
        To know this truth about our God— \\
        It lights the path that we must trod.
    }

    \stanza{
        Dear Ellen’s words, in context found, \\
        Reveal a truth that’s so profound \\
        Not trinity did she embrace, \\
        But trio persons in their place.
    }

    \stanza{
        The pillar stands, our platform firm, \\
        God’s personality we learn. \\
        The trio that is heavenly, \\
        Exposes falsehood—trinity.
    }
    
\end{titledpoem}