% Dr. Kellogg and the Trinity doctrine

\begin{titledpoem}
    \stanza{
        In Kellogg's quest, a question posed, \\
        "Is the Holy Ghost, indeed, a person enclosed?" \\
        A controversy stirred, the essence debated, \\
        The Trinity's mystery, intricately related.
    }

    \stanza{
        Kellogg questioned beyond the tangible, the seen, \\
        Where the Holy Spirit's form has never been. \\
        A spiritual essence, not bound by space, \\
        Challenging the Father's physical embrace.
    }

    \stanza{
        The Holy Ghost, a person, yet not in form, \\
        In actions, emotions, a divine norm. \\
        Witnessing, teaching, decisions so clear, \\
        A presence felt, though not seen near.
    }

    \stanza{
        Butler contrasted, in physical sense, \\
        The Father and Christ, their presence immense. \\
        Yet, the Spirit's persona, distinct in its role, \\
        A spiritual manifestation, completing the whole.
    }

    \stanza{
        Ellen asked of Jesus, if His Father's form was like His own, \\
        "In my Father's image," He replied, in a tone so gently sown. \\
        Yet, the Spirit, in essence, a guiding light, \\
        Not seen, but felt, in the believers' sight.
    }

    \stanza{
        A doctrine emerges, the Trinity's core, \\
        Equal in personalities, yet so much more. \\
        Kellogg's perspective, once seen as stray, \\
        Raises the question, in a theological way.
    }

    \stanza{
        Yet, Scripture guides, through mystery's veil, \\
        Revealing God's form, where human views fail. \\
        The Father, embodied, a truth we embrace, \\
        While the Spirit's presence, a formless grace.
    }

    \stanza{
        In divine revelation, the answers found, \\
        God's personality, in Scripture, profound. \\
        The Father, in form; the Spirit, without, \\
        In this, the Bible resolves all doubt.
    }
\end{titledpoem}