% The future of the Fundamental Principles
% TODO
\begin{titledpoem}
    
    \stanza{
        In the early whisper of prophecy’s sound, \\
        Ellen White warned where dangers abound. \\
        "The enemy plots," she sternly declared, \\
        "To dismantle truths our forebears shared."
    }

    \stanza{
        Fundamental Principles, strong and sure, \\
        Once the bedrock, pure and unpure. \\
        Now the sands shift beneath our creed, \\
        Where new doctrines grow like wayward weed.
    }

    \stanza{
        Reformation masked as light so bright, \\
        Undermines the pillars, eroding right. \\
        Books rewritten, philosophies anew, \\
        Skepticism veils what once we knew.
    }

    \stanza{
        Shall the Sabbath lose its sacred glow? \\
        Shall we forget the God that we owe? \\
        "The foundation crumbles," so it seems, \\
        As truth is lost to intellectual dreams.
    }

    \stanza{
        Look back to the days, to the Spirit-led start, \\
        Where divine truths were etched in heart. \\
        Satan’s strategies, cunning and keen— \\
        Eroding what once was clearly seen.
    }

    \stanza{
        So, brethren, now to the past return, \\
        To the roots of faith, let our hearts yearn. \\
        For the warnings spoken, the visions seen, \\
        Call us to defend what they truly mean.
    }

    \stanza{
        Stand firm in the storm as the tempests roar, \\
        Reclaim the truths worth fighting for. \\
        Ellen’s voice echoes, stark and clear: \\
        "Repudiate the false, hold the righteous near."
    }

    \stanza{
        Make your choice, as the battle lines draw, \\
        On the side of the timeless, divine law. \\
        To the Fundamental Principles, fiercely hold, \\
        The original faith, courageous and bold.
    }
    
\end{titledpoem}